Evidence suggesting creatine as a new central neurotransmitter: presence in synaptic vesicles, release upon stimulation, effects on cortical neurons and uptake into synaptosomes and synaptic vesicles

Version 1

Abstract

It has never been easy to discover a new neurotransmitter, especially one in the central nervous system (CNS). We have been searching for new neurotransmitters for 12 years. We detected creatine (Cr) in synaptic vesicles (SVs), at a level lower than glutamate (Glu) and gamma-aminobutyric acid (GABA) but higher than acetylcholine (ACh) and 5-hydroxytryptamine (5-HT). SV Cr was reduced in mice lacking either arginine:glycine amidinotransferase (AGAT, a Cr synthetase) or SLC6A8, a Cr transporter with mutations among the most common causes of intellectual disability (ID) in men. Calcium-dependent release of Cr was detected after stimulation in brain slices. Cr release was reduced in SLC6A8 and AGAT mutants. Cr inhibited neocortical pyramidal neurons. SLC6A8 was necessary for Cr uptake into synaptosomes. Cr was found by us to be taken up into SVs in an ATP dependent manner. Thus, our biochemical, chemical, genetic and electrophysiological results suggest Cr as a neurotransmitter, illustrate a novel approach to discover neurotransmitters and provide a new basis for ID pathogenesis.

Neural signaling depends on chemical transmission between neurons and their target cells 1–6. Neurotransmission depends on chemicals such as neurotransmitters, neuromodulators and neuropeptides. Decades of work, sometimes with convoluted paths, were involved before a molecule was established as a classic neurotransmitter 7. Initial hints about cholinergic signaling were obtained in the 1800s 8–11. Choline 12,13 and acetylcholine (ACh) 14 were discovered decades before their pharmacological effects were found around the turn of 20th century 15–17. Henry Dale and colleagues realized similarities of ACh and parasympathetic stimulation 5,18 but it was not until 1929 when ACh was detected in the body 19 and 1934 when ACh was proven a neurotransmitter in the peripheral nervous system (PNS) 20–22. It took nearly 100 years from the finding of the effects of supradrenal gland damage 23 or removal 24, the observation of an activity in the supradrenal gland 25, the isolation of an inactive derivative 26–29, and the successful isolation of adrenaline 30–32, the notice of similarities between adrenaline and sympathetic stimulation 3,4, 33–35, to the mid-1940s when Ulf von Euler proved that noradrenaline (NA) was the neurotransmitter of the sympathetic nerves 36–38. While it is not easy to establish a molecule as a neurotransmitter in the PNS, it is even harder to establish a central nervous system (CNS) neurotransmitter. Three decades elapsed between the time when ACh was proven to be a PNS neurotransmitter and the time when it was established as a CNS neurotransmitter 39–41 and two decades between NA as a peripheral transmitter and a central transmitter 42,43.

If a neurotransmitter acts only in the CNS, but not in the PNS, it is much more difficult to discover or to prove. Most neurotransmitters were discovered for their effects on peripheral tissues, with muscle contraction or relaxation as a major readout. Glutamate (Glu) 44–47 and gamma-aminobutyric acid (GABA) 46,48–51 were discovered partly because of their peripheral effects and partly because of their effects on spinal neurons. There is no reason for central neurotransmitters to also act peripherally, but relatively little efforts have been reported to find small molecule neurotransmitters acting only on CNS neurons with no peripheral bioassays available. Premature assumptions and technical difficulties are among the major reasons why the hunt for neurotransmitters has not been a highly active area of research over the last three decades.

Are there more neurotransmitters and how can they be discovered? Classic neurotransmitters are stored in synaptic vesicles (SVs) 52–60. They are released upon electric stimulation before being degraded enzymatically or taken up into the presynaptic terminal by cytoplasmic transporters and into SVs by vesicular transporters61–66. Most of the major textbooks list either 3 67–69 or 4 70–72 criteria of a neurotransmitter: presence in presynaptic neurons, release upon stimulation, action on postsynaptic neurons, mechanism of removal. Some molecules commonly accepted as neurotransmitters still do not meet all the criteria listed in different textbooks, but they nonetheless play important functional roles in the CNS and their defects cause human diseases. Over time, different small molecules have been proposed to function as neurotransmitters (e.g., 46,73), but none satisfies all the criteria. Robust and reliable detection of the candidate molecule in SVs is often, though not always, the problem (cf. 74).

Beginning in 2011, we have been actively searching for new neurotransmitters in the mammalian brain. We have tried different approaches, including searching for neuroactive substances in the cerebral spinal fluid (CSF) and following transporters potentially localized in the SVs. One approach that we have now taken to fruition is the purification of the SVs from mouse brains coupled with chemical analysis of their contents. We have found known transmitters such as Glu, GABA, ACh, and 5-hydroxytryptamine (5-HT). But more importantly, we have reproducibly detected creatine (Cr) in SVs.

Cr was discovered in 1832 by Michel-Eugène Chevreul 75,76 and has long been considered as an energy buffer in the muscle and the brain 77–79. Half of Cr in a mammalian animal is thought to come from diet and the rest from endogenous synthesis 80. Most of the Cr is present in the muscle but it is also present in the brain. Although most of the endogenous Cr is synthesized in the kidney, the pancreas and the liver 77,81, Cr is also synthesized in the brain 80,82,83.

Solute carriers (SLC) contribute to both cytoplasmic and vesicular transporters. With 19 members in humans, family 6 (SLC6) are secondary active transporters relying on electrochemical Na+ or H+ gradients 84–89. SLC6 is also known as the neurotransmitter transporter (NTT) family because some members transport neurotransmitters such as GABA (by SLC6A1 or GABA transporter 1, GAT1; SLC6A13 or GAT2; SLC6A11 or GAT3) 90–93, NA (by SLC6A2 or NA transporter, NET) 65, dopamine (by SLC6A3 or DAT) 94–96, 5-HT (by SLC6A4 or serotonin transporter, SERT) 66,97, and glycine (by SLC6A9, or GlyT1; SLC6A5 or GlyT2) 90,98,99. Cr is transported by SLC6A8 (also known as CrT, CT1 or CRCT) 100–106. In addition to peripheral organs and tissues, SLC6A8 is also expressed in the nervous system where it is mainly in neurons 82,104, 107–109. SLC6A8 protein could be found on the plasma membrane of neurons 109–111.

The functional significance of SLC6A8 in the brain is supported by symptoms of humans defective in SLC6A8. Mutations in SLC6A8 were found in human patients with intellectual disability (ID), delayed language development, epileptic seizures and autistic-like behaviors 112,113. They are collectedly known as Cr transporter deficiency (CTD), with ID as the hallmark. Particular vulnerability of language development has been observed in some SLC6A8 mutations which had mild ID but severe language delay 114. CTD contributes to approximately 1 to 2.1% of X-linked mental retardation 115–123. While CTD is highly prevalent in ID males, it is also present in females, with an estimated carrier frequency of 0.024% 124.

SLC6A8 knockout mice 125–127 showed typical symptoms of human CTD patients with early and progressive impairment in learning and memory. Mice with brain- and neuronal-specific knockout of SLC6A8 showed deficits in learning and memory without changes in locomotion caused by peripheral involvement of SLC6A8 128,129. Deletion of SLC6A8 from dopaminergic neurons in the brain caused hyperactivity 130. These results demonstrate that SLC6A8 is functionally important in neurons.

Cr deficiency syndromes (CDS) are inborn errors of Cr metabolism, which can result from defects in one of the three genes: guanidinoacetate methyltransferase (GAMT) 131, arginine-glycine amidinotransferase (AGAT) 132,133, and SLC6A8 112. That they all show brain disorders indicates the functional importance of Cr in the brain134–137.

Here we first biochemically purified SVs from the mouse brain and discovered the presence of Cr, as well as classic neurotransmitters Glu and GABA, ACh and 5-HT, in SVs. We then detected calcium (Ca2+) dependent releases of Cr, Glu and GABA but not ACh and 5-HT when neurons were depolarized by increased extracellular concentrations of potassium (K+). Both the level of Cr in SVs and that of Cr released upon stimulation were decreased significantly when either the gene for SLC6A8 or the gene for AGAT were eliminated genetically. When Cr was applied to slices from the neocortex, the activities of pyramidal neurons were inhibited. Furthermore, we confirmed that Cr was taken up by synaptosomes and found that Cr uptake was significantly reduced when the SLC6A8 gene was deleted. Finally, we found that Cr was transported into SVs. Thus, multidisciplinary approaches with biochemistry, genetics and electrophysiology have obtained results consistent with the possibility of Cr as a neurotransmitter.

Detection of Cr in SVs from the mouse brain

To search for new neurotransmitters, we tried several approaches. For example, we used Ca2+ imaging to detect neuroactive substances in the cerebrospinal fluid (CSF), but it was difficult to rule out existing neurotransmitters and select responses from potentially new neurotransmitters. We also transfected cDNAs for all human SLCs into dissociated cultures of primary neurons from the mouse brain and found that more than 50 out of all SLCs could be localized in SVs. However, when we used CRISPR-CAS9 to tag some of the candidate SLCs in mice, some of them were found to be expressed outside the CNS, indicating that, while ectopic expression of these candidate SLCs could be localized on SVs, the endogenous counterparts were not localized on SVs.

Here we report our approach using the purification of SVs as the first step (Supplementary Figure 1A) 138–145. Synaptophysin (Syp) is a specific marker for SVs146–149 and an anti-Syp antibody was used to immunoisolate SVs 143,144, 150–152. Visualization by electronic microscopy (EM) (Supplementary Figure 1B, left panel) showed that the purified vesicles were homogenous, with an average diameter of 40.44 ± 0.26 nm (n=596 particles) (Supplementary Figure 1B, right panel), consistent with previous descriptions of SVs 153,154.

Immunoblot analysis with twenty markers of subcellular organelles of neurons and one marker for glia (Supplementary Figure 1C) indicate that our purifications were highly effective, with SV markers detected after purification with the anti-Syp antibody, but not that with the control immunoglobulin G (IgG). SV proteins included Syp 146–149, synaptotagmin (Syt1) 155, synatobrevin2 (Syb2) 156,157, SV2A 158, H+-ATPase 159, and vesicular neurotransmitter transporters for glutamate (VGLUT1, VGLUT2) 160 and GABA (VGAT) 161. Immunoisolation by the anti-Syp antibody did not bring down markers for the synaptic membrane (with SNAP23 as a marker) 162–164, postsynaptic components (with PSD95 and GluN1 as markers) 165,166, the Golgi apparatus (with GM130 and Golgin 97 as markers) 167,168, early endosome (with early endosome-associated 1, EEA1, as a marker) 169, the lysosome (with LC3B and cathepsinB as markers), the cytoplasma (with glyceraldehyde-3-phophate dehydrogenase or GAPDH as a marker), mitochondria (with voltage-dependent anion channel or VDAC, as a marker), cytoplasmic membrane (with calcium voltage-gated channel subunit alpha 1 or CACNA1A as a marker), axonal membrane (with glucose transporter type 4 or GluT4 as a marker) and glia membrane (with myelin basic protein or MBP as a marker). These results indicated that the SVs we obtained were of high integrity and purity.

To detect and quantify small molecules as candidate transmitters present in the purified SVs, capillary electrophoresis-mass spectrometry (CE-MS) was optimized and utilized (Figure 1A, Supplementary Figure 1A) 151,170. We found that the levels of classical neurotransmitters such as Glu, GABA, ACh and 5-HT were significantly higher in SVs pulled down by the anti-Syp antibody than those in lysates pulled down by the control IgG (Figure 1 A-E). Consistent with previous reports 144,151, significant enrichment of neurotransmitters was observed only from SVs immunoisolated at near 0 °C, but not at the room temperature (RT) (Figure1 A-E). By contrast, another small molecule, alanine (Figure 1G), was not elevated in SVs as compared to the control.

Presence of Cr in SVs from the mouse brain.

(A) Representative raw traces from CE-MS of indicated molecules from samples immunoisolated by the control IgG (blue) at 0-2 °C, the monoclonal anti-Syp antibody at 0-2 °C (red) and the anti-Syp antibody at room temperature (RT, green). Q1/Q3 for identifying targets were indicated. (B-G) Quantification of the amounts of indicated molecules. The amount of a molecule was divided by the amount of the anti-Syp antibody bound to magnetic beads. Note Glu (B), GABA(C), ACh (D), 5-HT (E), Cr (F), but not alanine (G) was higher in SVs pulled down by the anti-Syp antibody at 0-2 °C than those pulled down by the IgG control or those pulled down at the RT. n=10 (B-E, and G) or 14(F) samples per group, ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

The amount of Glu was 171.1 ± 5.4 pmol/μg anti-Syp antibody (n=10, Figure 1B), approximately 10 times that of GABA (n=10,17.81 ± 1.47 pmol/μg anti-Syp antibody, Figure 1C). The amount of ACh was 1.29 ± 0.10 pmol/μg anti-Syp antibody (n=10, Figure 1D), approximately 0.072 that of GABA. The amount of 5-HT was 0.096 ± 0.017 pmol/μg anti-Syp antibody (n=10, Figure 1E). Thus, our purification and detection methods were highly reliable and sensitive enough to detect established neurotransmitters.

Under the same conditions, we also detected Cr in SVs (n=14, Figure 1A and 1F). Amount of Cr in the SVs was found to be 3.43 ± 0.40 pmol/μg anti-Syp antibody (Figure 1F), which was approximately 2% of Glu, 19% of GABA, 266% of ACh and 3573% of 5-HT. It is unlikely that these could be attributable to different levels for different neurotransmitters in each SV, but more likely attributable to the relative abundance of SVs containing different neurotransmitters. 85% to 90% neurons in the mouse brain were glutamatergic while 10%-15% were GABAergic 171–173, which can explain our detection of Glu as approximately 10 times that of GABA (Figure 1 B and C). Similarly, cholinergic neurons (5.67×105) 174 represented 0.81% of total number of neurons (approximately 70 million) in the mouse brain 175, serotonergic neurons (approximately 26,000) for 0.037% of total neurons 175,176. Assuming that the content of each neurotransmitter in a single SV is similar, extrapolation from the above data would suggest that approximately 1.3% to 2.15% of neurons in the mouse brain are creatinergic.

To distinguish whether small molecules co-purified with SVs were in the SVs 143,144, or that they were just associated with the outside of SVs154, we tested the dependence of the presence of these molecules in the SVs on temperature and on the electrochemical gradient of H+. Cr was significantly reduced in SVs purified at the room temperature as compared to that immunoisolated at near 0 °C (Figure 1F), supporting the presence of Cr inside, instead of outside, SVs.

Classical neurotransmitters are stored in SVs with an acidic environment inside (pH of 5.6 ∼ 6.4) 177–179. To further verify the storage of Cr in SVs and examine the role of H+ electrochemical gradient, we applied pharmacological inhibitors during purification 74,180. The proton ionophore FCCP (carbonyl cyanide-4-(tri-fluoromethoxy) phenylhydrazone) was used to dissipate H+ electrochemical gradient 180,181. FCCP significantly reduced the amount of Cr as well as classical neurotransmitters in SVs (Supplementary Figure 2A-E). The extent of FCCP induced reduction was correlated with the value of pKa or PI (isoelectric point) for different molecule: 5-HT (with pKa predicted to 10 and 9.31, Supplementary Figure 2E) > Cr (PI of ∼7.94, Supplementary Figure 2A) > GABA (PI of 7.33, Supplementary Figure 2C) > Glu (PI of 3.22, Supplementary Figure 2B). Nigericin, a K+/H+ exchanger which dissipates ΔpH180,181, also reduced the amount of Cr and classical neurotransmitters in SVs (Supplementary Figure 2A-E). Furthermore, in the presence of FCCP or nigerin, SV Cr was reduced to a level comparable to that pulled down by control IgG (Supplementary Figure 2A), demonstrating the storage of Cr in SVs was dependent on H+ gradient. As a control, the non-neurotransmitter molecule alanine in SVs was not changed by the inhibitors (Supplementary Figure 2F).

Reduction of SV Cr in mouse mutants lacking slc6a8

SLC6A8, located on the X chromosome, encodes a transporter for Cr and its loss of function (LOF) mutations caused behavioral deficits in humans 112,113 and mice 125–130. To investigate whether SLC6A8 affects Cr in SVs, we generated slc6a8 knockout (KO) mice. Exon 1 of the SLC6A8 gene was partially replaced with CreERT2-WPRE-polyA by CRISPR/Cas9 (Figure 2A)). Examination by reverse polymerase chain reaction (RT-PCR) (Supplementary Figure 3A and 3B) and quantitative real-time reverse PCR (qPCR, Supplementary Figure 3C-D) showed that SLC6A8 mRNA was not detected in either male or female mutants, and significantly reduced in female heterozygous (Slc6a8+/-). Consistent with previous reports, the body weights of SLC6A8 KO mice were reduced (Supplementary Figure 4 B and D) 125,182,183. Brain weight was not significantly different between SLC6A8 KO mice and WT mice (Supplementary Figure 4A and C).

SLC6A8 and Cr in SVs.

(A) A schematic illustration of the strategy for generating SLC6A8 knockouts using CRISPR/Cas9. An iCreERT2-WPRE-pA cassette (∼3.5 kb) was inserted immediately downstream of ATG in the SLC6A8 gene, substituting bp 4 to bp 51 in exon 1 (E1). (B) Representative raw traces of Creatine immunoisolated by control IgG from WT mice (blue), the anti-Syp antibody from WT mice (red), IgG from SLC6A8 KO mice (blue) and the anti-Syp antibody from SLC6A8 KO mice (red). (C-H) Quantification of indicated molecules. Note the selective reduction of Cr in SVs from SLC6A8 KO mice. n=14 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

Reduction of SV Cr in AGAT knockout mice.

(A) Representative raw traces of Creatine, pulled down by IgG or anti-Syp from AGAT+/+, AGAT+/- and AGAT-/- mice. (B-G) Quantification of indicated molecules. Cr was significantly decreased in AGAT-/- mice as compared to Cr in AGAT+/+ or AGAT+/- mice (B). GABA was significantly decreased in SVs from AGAT-/- mice as compared to AGAT+/+ mice (D) but the difference was smaller than that of Cr. Glu (C), ACh (E), 5-HT (F) and alanine was not different among AGAT+/+, AGAT+/- and AGAT-/- mice. n=10 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

When we examined the contents of SVs isolated by the anti-Syp antibody vs the control IgG, significant reduction was only observed for Cr, but not classical neurotransmitters (Figure 2 B-H, Supplementary Figure 5 A-E). While Cr pulled down by IgG was not significantly different between SLC6A8-/Y and SLC6A8+/Y mice, SV Cr purified by the anti-Syp antibody from SLC6A8-/Y was reduced to approximately 1/3 that of the WT (SLC6A8+/Y) littermates (n =14, Figure 2 B-C). Compared to the IgG control, Cr in SVs was enriched in WT mice, but not in SLC6A8-/Y mice (Figure 2 B and C). In both SLC6A8-/Y and SLC6A8+/Y mice, classical neurotransmitters in SVs were all enriched as compared to IgG controls (Figure 2 C-G, Supplementary Figure 5 A-D). The amounts of Glu (Figure 2D and Supplementary Figure 5A), GABA (Figure 2E and Supplementary Figure 5B), ACh (Figure 2F and Supplementary Figure 5C) and 5-HT (Figure 2G and Supplementary Figure 5D) in SVs were not different between SLC6A8-/Y and SLC6A8+/Y mice. Molecules not enriched in SVs from WT mice, such as alanine, was also unaffected by SLC6A8 KO (Figure 2H and Supplementary Figure 5E).

It is unlikely that the specific reduction of Cr in SVs from SLC6A8 KO mice was due to technical artifacts. First, the possibility of less SVs obtained from SLC6A8 KO mice were precluded by immunoblot analysis, as assessed by SV markers Syp, Syt and H+-ATPase (Supplementary Figure 6A). Second, data collected by high resolution MS (Q Exactive HF-X, Thermoscientific) also revealed selective decrease of SV Cr (m/z=132.0576) from SLC6A8 KO mice (n=8, Supplementary Figure 6 B-G), as quantified by the peak area. Peak areas for Glu (n=8, Supplementary Figure 6B, m/z=148.0604), GABA (n=8, Supplementary Figure 6C, m/z=104.0712), ACh (n=8, Supplementary Figure 6D, m/z = 146.1178) and alanine (n=8, Ala, Supplementary Figure 6E, m/z=90.055) were not significantly different between SVs immmunoisolated with the anti-Syp antibody and control IgG from WT and Slc6a8 KO mice. However, peak areas (Supplementary Figure 5G) and amplitude of Cr (n=8, Supplementary Figure 5F) signal were significantly increased in SVs from WT mice (anti-Syp antibody vs IgG), but not that from SLC6A8 KO mice.

Reduction of SV Cr in mouse mutants lacking AGAT

AGAT is the enzyme catalyzing the first step in Cr synthesis 82,184 and its absence also led to Cr deficiency in the human brain and mental retardation 132,133.

To investigate requirement of AGAT for SV Cr, we utilized AGAT ‘knockout-first’ mice (Fig. 6a) 185. The targeting cassette containing Frt (Flip recombination sites)-flanked EnS2A, an IRES::lacZ trapping cassette and a floxed neo cassette were inserted downstream of exon 2 to interfere with normal splicing of AGAT pre-mRNA. Examination by RT-PCR (Supplementary Figure 7A) and quantitative rtPCR (Supplementary Figure 7B) showed reduction of AGAT mRNA in AGAT+/- mice and absence of AGAT mRNA in AGAT-/- mice. Body weight (Supplementary Figure 8B), but not brain weight (Supplementary Figure 8A), of AGAT-/- mice were lower than both AGAT+/+ and AGAT+/- mice, which were similar to SLC6A8 KO mice.

Immunoblot analysis showed SVs purified from the brains were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice (Supplementary Figure 9 A-D), as supported by quantitative analysis of Syp, Syt and H+-ATPase (n=20, with 2 repeats for 10 samples).

We analyzed small molecules present in SVs from AGAT+/+, AGAT+/- and AGAT-/- mice. Cr was significantly enriched in SVs from all three genotypes as compared to the IgG control (Figure 3A and 3B). However, the level of Cr from AGAT-/- mice was significantly lower than those from AGAT+/+ and AGAT+/- mice (n=10, Figure 3A and B).

Glu (Figure 3C), ACh (Figure 3E) and 5-HT (Figure 3F) were all enriched in SVs (as compared to IgG controls), and not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice. GABA in SVs from AGAT-/- mice was also decreased from AGAT+/+ mice by 30%, to an extent less than that of Cr (78.4%). Alanine was not different among three genotypes of mice (n=6, Figure 3G). Thus, Cr and GABA, but not other neurotransmitters, in SVs were reduced in AGAT KO mice.

Pattern of SLC6A8 expression indicated by knockin mice

We generated SLC6A8-HA knockin mice by CRISPR/Cas9. Three repeats of the hemagglutinin (HA) tag 186, the T2A sequence 187–189 and CreERT2 190–193 were inserted in-frame at the C terminus of the SLC6A8 protein (Figure 4A).

Expression pattern of SLC6a8.

(A) A diagram illustrating the generation of SLC6A8-HA KI mice. 3*HA, T2A and CreERT2 were inserted in-frame, before the stop codon, to the C terminus of SLC6A8 protein. (B-K) Low magnification photomicrographs of coronal sections immunohistochemically labelled with the anti-HA antibody in SLC6A8-HA mice. (I and M) immunostaining with the anti-HA antibody in control WT mice. Abbreviations: Pir, piriform cortex; M, motor cortex; LS: lateral septum; Hp: hippocampus; Hb, habenular nucleus; Vp: ventral posterior nucleus of thalamus; Au auditory cortex; Amyg: amygdala; Ip, interpeduncular nucleus; Pn; pontine nucleus; Cb, cerebellum; Pr, prepositus; SC, Superior colliculus; MVe, medial vestibular nucleus. Scale bar: 500 μm.

To examine the expression pattern of SLC6A8, we performed immunocytochemistry with an antibody against the HA epitope in SLC6A8-HA and WT mice. SLC6A8-HA mice howed positive signals in the olfactory bulb (Figure 4B), the piriform cortex (Figure 4 C-F), the somatosensory cortex (Figure 4 F and G), the ventral posterior thalamus (Figure 4H), the interpeduncular nucleus (Figure 4I) and the pontine nuclei (Figure 4J). In addition, moderate levels of immunoreactivity were observed in the motor cortex (Figure 4 D-H), the medial habenular nucleus (Figure 4H), the hippocampus (Figure 4H) and the cerebellum (Figure 4K). These results were consistent with previous reports 109,111. WT mice were negative for anti-HA antibody staining (Figure 4 L and M).

Ca2+ dependent release of Cr upon stimulation

Classical neurotransmitters are released from the SVs into the synaptic cleft in a Ca2+ dependent manner after stimulation. For example, high extracellular potassium (K+) stimulated Ca2+ dependent release of Glu, GABA and other neurotransmitters in brain slices 194–198.

300 μm thick coronal slices of the mouse brain within 1∼2 millimeters (mm) posterior to the bregma were used, because the cortex, the thalamus, the habenular nucleus and the hippocampus were positive for SLC6A8 (cf., Figure 4H). We monitored the effect of K+ stimulation by recording neurons in the slices. Immediately after K+ stimulation, pyramidal neurons in the CA1 region of the hippocampus were depolarized, firing a train of action potentials and reaching a large depolarization plateau in less than 1 minute (min) (Figure 5A). K+-induced depolarization persisted for several mins before returning to the baseline and being washed within 10 mins. Thus, superfusates in 1-min fraction at the time points of 1.5 min before (control) and after K+ stimulation, and 10 mins after the wash were collected (Figure 5A), and the metabolites in the superfusates were analyzed by CE-MS.

Cr release in brain slices from WT and SLC6a8 knockout mice.

(A) Neuronal depolarization induced by 70 mM K+ and time points for collecting the release sample. “Control” samples were collected 1.5 min to 0.5 min before K+ stimulation, “70 mM K+” ACSF samples were collected during 70 mM K+ stimulation, and “wash” samples were collected 10 mins after washout with ACSF. Efflux of Glu (B), GABA (C) or Cr (D) from WT or SLC6A8 KO male mice (n=7 samples per group). Note that a small amount of Cr released in SLC6A8 KO mice did not return to the baseline after 10 mins of washing. (E-G) Ca2+ dependent release of Glu, GABA and Cr in WT and SLC6A8 KO mice (n=5 samples per group). ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction. #, p<0.05, ##, p<0.01, paired t test.

In the presence of Ca2+, depolarization with elevated extracellular K+ led to robust release of Glu and GABA in slices from WT (SLC6A8+/Y) mice (n=7 per group, Figure 5 B and C). After 10 mins wash, levels of Glu and GABA returned to the baseline (Figure 5 B and C). In the presence of Ca2+, depolarization with elevated K+ led to robust release of Cr. Extracellular Cr returned to the baseline after 10 min wash (Figure 5D). For quantification, the stimulated releases of metabolites were calculated by subtracting the basal levels from the total releases in response to K+ stimulation. In the presence of Ca2+, K+ stimulation induced the efflux of Glu, GABA and Cr at 0.46, 0.33 and 0.086 nmol/min, respectively (n=7 per group) (Figure 5 B, C, D). From the detection limits of ACh and 5-HT in our system, we inferred that the efflux rate for ACh was lower than 0.001 nmol/min and that for 5-HT lower than 0.003 nmol/min. The efflux rate for Cr in brain slices is lower than those of Glu and GABA, but higher than those for ACh and 5-HT.

Ca2+ dependence of transmitter release was examined by comparing responses to ACSF without Ca2+ or elevated K+ (supplemented with 1 mM EGTA), elevated extracullular K+ in the absence of Ca2+ (supplemented with 1 mM EGTA), or K+ in presence of 2.5 mM Ca2+ (Figure 5 E-G, n=5 per group). In the absence of Ca2+, elevated K+ stimulated the release of a small but significant amount of Glu and GABA, with efflux rates at 0.056 nmol/min and 0.066 respectively (Figure 5 E,F). In the presence of 2.5 mM Ca2+, elevated K+ further augmented the release of Glu and GABA by 5 to 6 times, confirming previously reported Ca2+ dependent release of neurotransmitters in response to depolarization 194,196.

Cr was also released both in a Ca2+ dependent and a Ca2+ independent manner (Figure 5G). More Cr was released in response to K+ stimulation in the presence of 2.5 mM Ca2+ than that in the absence of Ca2+. These results demonstrate Ca2+ dependent release of Cr upon stimulation.

Cr release reduced in SLC6A8 and AGAT mutant mice

We examined whether SLC6A8 KO affected K+-induced release of Cr. While Glu and GABA were released in slices from SLC6A8 KO (SLC6A8-/Y) mice at levels not significantly different from those of WT mice (Figure 5 B,C), release of Cr in response to K+ stimulation was significantly reduced in SLC6A8-/Y mice as compared to SLC6A8+/Y mice (Figure 5D). The basal level of Cr in SLC6A8 KO mice was lower than that of WT mice. In addition, K+ stimulation induced release of Cr persisted to some extent even after 10 mins of washout (Figure 5D), possibly due to the inability of presynaptic terminals in SLC6A8 KO mice to reuptake Cr in the synaptic cleft (Figure 8).

Experiments with slices from brains of SLC6A8 KO (SLC6A8-/Y) mice showed that Ca2+ dependent release of either Glu or GABA was not affected by the genotype of SLC6A8 (Figure 5 E and F). By contrast, Ca2+ dependent release of Cr was abolished in SLC6A8-/Y slices. Interestingly, Ca2+ independent release of Cr was significantly reduced, but not abolished in SLC6A8-/Y slices. In the absence of Ca2+, the basal level of Cr was not changed in SLC6A8 KO mice. Taken together, these results indicate that there is Ca2+ dependent release of Cr upon stimulation and that SLC6A8 is required specifically for Ca2+ dependent release of Cr, but not for Ca2+ dependent release of other neurotransmitters such as Glu and GABA, or for Ca2+ independent release of Cr.

Knockout of AGAT (Figure 6A) selectively reduced K+ evoked release of Cr, but not those of Glu or GABA (n=5 per group, Figure 6 B, C, D). Although K+ stimulation still elicited Cr release from brain slices of AGAT+/- mice, the efflux rate in AGAT-/- mice was reduced to less than 10% that in AGAT+/+ mice and 20% that in AGAT+/- mice. Thus, there is a strong AGAT dependent component in K+ induced release of Cr.

Cr release in WT and AGAT knockout mice.

(A) A schematic diagram illustrating the strategy of AGAT knockout first. With the AGAT gene (also known as GATM) shown in the upper part, and the gene targeting strategy in the lower part. The homologous arm is approximately 10 kb. A targeting cassette, containing Frt-flanked lacZ and neomycin, was inserted downstream of exon 2. At the same time, exon 3 of AGAT was flanked by two loxP sites. K+ induced release of Glu (C) and GABA (D) were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice, whereas that of Cr (B) was significantly lower in AGAT-/- mice than those in AGAT+/+ and AGAT+/- mice.

Cr inhibition of neocortical neurons

Our own data (Figure 4h) and previous reports 109,111 have shown SLC6A8 in the neocortex, with dense SLA6A8-HA immunoreactive fibers in layer 4 (Supplementary Figure 10A) 199,200. Layer 5 neurons in the somatosensory cortex have been reported to express SLC6A8 previously 109,111. To investigate electrophysiological effects of Cr, we performed whole cell patch clamp recordings from the pyramidal neurons in layer 4/5 of the somatosensory cortex (Figure 7 A and B,, Supplementary Figure 10).

Inhibitory effects of Cr on cortical neurons.

(A) A photograph showing recording at layer 4 in the somatosensory cortex. Scale bar: 10 μm. (B) Patch-clamp recording of a pyramidal neuron. Scale bar: 10 μm. (C) Ratios of Cr - responsive and -unresponsive neurons in the region. (D) Representative raw electrophysiological traces showing inhibition of evoked firing by Cr, with the lower panel showing the stimulus protocol. (E) Evoked spike numbers in response to different current injections from (D). (F) Relationship between evoked spike numbers and different current injections to neurons that were inhibited by Cr (n=16). (G) The same for Cr-unresponsive neurons (n=35). (H) Rheobase for Cr responsive neurons. (I) Evoked spike number when these neurons were injected with current of rheobase + 50 pA. (J) input resistance. *, p<0.05; **, p<0.01; ***, p<0.001, paired t test.

Medium-sized pyramidal neurons (Figure 7B) with a membrane capacitance (Cm) of 114.96 ± 3.92 pF (n=51, Supplementary Figure 11A) were recorded. These neurons exhibited regular firing patterns 201,202 in response to depolarization current injection (Figure 7D and Supplementary Figure 11A) with moderate maximal evoked spiking frequencies of 10-30 spikes per 500 milliseconds (ms) (Figure 7 E-G), increasing of inter-spike intervals during depolarizing steps (Figure 7D and Supplementary Figure 11A), high action potential amplitude (81.64 ± 1.06 mV, Supplementary Figure 12E) and large spike half-width (1.12 ± 0.031 ms, Supplementary Figure 12F).

Cr was bath applied only after the evoked firing pattern reached a steady state. Of the 51 neurons, 16 were inhibited by 100 μM creatine (Figure 7 B-F). Fewer spikes were evoked in Cr responsive neurons in response to depolarizing current injections during Cr application (25 pA step, 500 ms) (Figure 7 D-F). The inhibitory effect of Cr was reversible (Figure 7 D-F), typically observed within 2-3 min following Cr application (with maximal effect from 2 to 8 mins) and disappeared after 10 to 25 min washout. This could be repeated by a second application of Cr. The rheobase, defined as the minimal electrical current necessary to elicit an action potential, was increased during bath application of Cr (Figure 7 H). The inhibitory effect was most obvious at near spike threshold. When a neuron was depolarized with a current of 50 pA above rheobase, the number of evoked spikes was decreased dramatically during Cr application (Figure 7I). Cr also mildly inhibited the input resistance (Figure 7J), slightly hyperpolarized resting membrane potential (Supplementary Figure 12C) or reduced amplitude of afterhyperploarization (AHP) followed by the first evoked action potential (Supplementary Figure 12G). The spike threshold (Supplementary Figure 12D), amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F) were not changed by Cr.

The remaining 35 neurons were not responsive to Cr (Figure 7C and G, Supplementary Figure 11). Cr did not change electrical parameters tested, including evoked firing rates (Figure 7G, Supplementary Figure 11 A, B and D), rheobase (Supplementary Figure 11C), resting membrane potential (Supplementary Figure 12C), spike threshold amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F). In addition, electrical properties of responsive neurons and unresponsive neurons were not significantly different. With the limited number of neurons recorded, the ratio of responsive neurons appeared higher in layer 4 or border of layer4/5, than the deeper layer in layer 5 (Supplementary Figure 10).

SLC6A8 dependent uptake of Cr into the synaptosomes

Along with enzymatic degradation, re-uptake by transporters serves as an important way to remove neurotransmitters released into the synaptic cleft. As synaptosomes contain the apparatus for neurotransmission, they are often used for studying uptake of neurotransmitters 203.

To investigate whether Cr uptake into synaptosomes required SLC6A8, we first examined whether SLC6A8 was present in synaptosomes. Using Slc6a8-HA knockin mice and an anti-HA antibody, we found that SLC6A8-HA was present and enriched in crude synaptosomal fraction (P2 fraction in Figure 8A, enrichment score: P2/H = 1.76 ± 0.15, n = 4) and synaptosomal fraction prepared using a discontinuous Ficoll gradient (Sy and 4-Sy fractions in Figure 8A, enrichment score: Sy/H = 2.02 ±0.14, n = 4). The integrity of synaptosomes was confirmed by multiple markers of the synaptosomes 203, including the presynaptic membrane marker SNAP25 204 and the SV marker Syp (synaptophysin Syp) 146–149 in presynaptic terminals, the postsynaptic density marker PSD95 165,166 and the postsynaptic membrane protein GluN1 205, the synaptic membrane protein SNAP23 162–164, the plasma membrane marker Na+-K+-ATPase 206–208, and the mitochondria marker VDAC 203. These were all enriched in our synaptosomal preparations. The cytosol marker GAPDH was also present in synaptosomes, whereas the oligodendrocyte marker MBP was nearly absent, suggesting myelin pollution was largely avoided 203.

We have also used EM to confirm the quality of our synaptosome preparations. As reported previously 203,209, synaptosomes were composed of membrane bounded structures (Sy in Figure 8B) filled with synaptic vesicles (SV in Figure 8B), sometimes with a segment of postsynaptic membrane along with the postsynaptic density (PSD in Figure 8B) and mitochondria (Mt in Figure 8B). The sizes of synaptosomes from WT mice and Slc6a8 knockout mice were similar, with areas of 0.245 ± 0.01 μ (n=302 particles) and 0.247 ± 0.01 μm (n=317 particles), respectively (Figure 8B).

Cr uptake into synaptosomes and SVs.

(A) Markers for subcellular organelles detected in synaptosomes prepared from WT mice or mice with Slc6a8 gene fused in frame to the HA epitope. SLC6A8-HA was enriched in synaptosomes (Sy or 4-Sy) or crude synaptosomes (P2) (see methods), as were other markers for the subcellular organelles in synaptosomes but not the marker for myelin (MBP). (B) Representative electron micrographs and histograms of size distribution in synaptosomes isolated from WT (Slc6a8+/Y) and Slc6a8 KO (Slc6a8-/Y) mice by Ficoll density-gradient centrifugation. Sy, synaptosome; SV, synaptic vesicles; Mt, mitochondria; PSD, post-synaptic density. Bar, 20 m. (C) Cr uptake into synaptosomes (n=5 per group). The two left columns were results from WT mice and the two right columns from Slc6a8 knockout mice. The control baseline was [14C]-Cr uptake at 0 °C at 10 mins (205). Cr uptake into synaptosomes at 37 °C measured at 10 mins was observed in WT synaptosomes. Uptake into Slc6a8 knockout synaptosomes was significantly reduced when compared to the WT synaptosomes. ***, p<0.001, one-way ANOVA with Tukey’s correction. (D) uptake of [13C]-Cr into immunoisolated SVs in the presence or absence of ATP. (n=3 samples per group)***, p<0.001, t test.

We then examined whether SLC6a8 participated in Cr uptake into the synaptosomes. A mixture of 18 μ C]-Cr (with a total radioactivity of 0.4 μCi) and 5 μM Cr was used, and uptake at 0 °C measured at 10 mins was the baseline. Cr uptake into synaptosomes from WT mice was stimulated approximately sevenfold at 37 °C (Uptake, Figure 8C) as compared with 0 °C (Ctrl, Figure 8C). Cr uptake into synaptosomes from SLC6a8 knockout mice was less than 3 times as compared to its control, and was decreased to approximately 1/3 of that of WT mice (Figure 8C). Thus, SLC6A8 is necessary for uptake of Cr into the synaptosomes.

Cr uptake into SVs

Classical neurotransmitters were taken up in SVs in an ATP-dependent manner144,159,181,210,211. We examined whether Cr could be transported into SVs.

We used 10 μg anti-Syp antibody to purify SVs from mouse brains. Purified SVs were mixed with 1 mM [13C]-Cr in the presence or absence of 4 mM ATP and placed at 25 for 10 mins to allow adequate uptake. The SV content of [C]-Cr was then examined by high-performance liquid chromatography-mass spectrometry (HPLC-MS). Significantly more [13C]-Cr were taken up by SVs in the presence of ATP, with about 3.3 pmol [13C]-Cr transported into SVs (0.33 pmol/μg α-Syp or transportation rate of 0.033 pmol/min, n=3, Figure 8D). In the absence of ATP, less than 0.5 pmol [13C]-Cr was absorbed by SVs, as inferred from the detection limits of HPLC/MS.

In summary, Cr could be transported into SVs in ATP-dependent manner. At this point, we do not know what is the transporter(s) on the SVs for Cr uptake. SLC6A8 is only found in plasma membrane, not on SVs, and is not a candidate for Cr uptake into SVs.

Discussion

At various times and by different researchers, taurine 46,212, proline 73, D-aspartic acid213, hydrogen sulfide 214, agmatine 215, DOPA 216, estradiol 217, β-alanine 218, protons 219 have been suspected as neurotransmitters, but they do not meet all the criteria. Some of the suspected molecules can be released upon stimulation or removed by transporters. Often, they have not been reproducibly found in SVs 74.

Our discovery of Cr in SVs significantly raised the priority of testing the candidacy of Cr, and our further investigations have led to more evidence suggesting Cr as a neurotransmitter: 1) Cr is stored in SVs; 2) Ca++ dependent release of Cr upon stimulation has been observed; 3) both Cr storage in SVs and Cr release are reduced when either the gene for SLC6A8 or the gene for AGAT was deficient; and 4) Cr inhibits activities of pyramidal neurons in the neocortex; 5) Cr uptake into synaptosomes requires SLC6A8; 6) Cr uptake into SVs was ATP-dependent.

Of the above results, 1, 3, 4 and 6 are reported for the first time in this paper. Furthermore, we have demonstrated that detection of Cr in SVs was lower than those for Glu and GABA, but higher than those for ACh and 5-HT, placing Cr at a level in the middle of known central transmitters (Figures 1-3). The storage of Cr in SVs is dependent on preserved H+ gradient (Supplementary Figure 2) and Cr can be transported into SVs (Figure 8D).

There was a single previous report of Ca++ dependent release of [3H]Cr and endogenous Cr in response to electrical stimulation 220. We now provide evidence that Cr was released in response to extracellular K+ stimulation (within 1-2 min) (Figure 5-6). Furthermore, Cr release was reduced when either the Slc6A8 or AGAT gene was removed (Figure 5-6). Although the Ca2+ dependent component of K+ evoked Cr release was smaller as compared to those of Glu and GABA, it nevertheless existed and was totally abolished by Slc6A8 knockout (Figure 5). The reported electrically evoked Cr release showed more Ca2+ dependence 220. Taken together, our data and previous report 220 supported a role of Cr as a neurotransmitter. Our observation of extremely low efflux rates of 5-HT or ACh may have arisen from very limited numbers of cholinergic174 or serotoninergic neurons176 in the sliced sections and rapid enzymatic degradation of these neurotransmitters.

Cr uptake from the extracellular space into the cells was reported twice previously, once with brain slices showing sodium dependent uptake of [3H]Cr 220 and once with synaptosomes 221. Our new results have not only replicated the synaptosome Cr uptake experiment but also shown the requirement of SLC6A8, a membrane transporter expressed in synaptosomes (Figure 8A), for Cr uptake into synaptosomes. Transportation of Cr into synaptosomes by Slc6A8 may function for both the clearance of Cr from the synaptic cleft and recycling of Cr into SVs residing in neurons (Figure 8B, Figure 6D).

In summary, in addition to confirming and extending previous results which have stood alone for more than a decade without replication or follow-up, we have obtained entirely new results suggesting the candidacy of Cr as a neurotransmitter. We discuss below the criteria for a neurotransmitter, Cr as a neurotransmitter, and the implications of Cr as a neurotransmitter.

Criteria of a neurotransmitter

The criteria for establishing a non-peptide small molecule as a neurotransmitter have varied from time to time and from author to author.

Some textbooks simply state that a neurotransmitter is stored presynaptically, released upon stimulation and active on postsynaptic neurons. The details of these 3 criteria can vary. For example, one textbook stipulates that “the substance must be present within the presynaptic neuron; the substance must be released in response to presynaptic depolarization, and the release must be Ca2+ dependent; specific receptors for the substance be present on the postsynaptic cell” 67,68. Another states that “the molecule must be synthesized and stored in the presynaptic neuron; the molecule must be released by the presynaptic axon terminal upon stimulation; the molecule, when experimentally applied, must produce a response in the postsynaptic cell that mimics the response produced by the release of neurotransmitter from the presynaptic neuron69.

The neuroscience textbook most widely used internationally for the last 4 decades lists 4 criteria for a neurotransmitter 70,71: it is synthesized in the presynaptic neuron; it is present within vesicles and is released in amounts sufficient to exert a defined action on the postsynaptic neuron or effector organ; when administered exogenously in reasonable concentrations it mimics the action of the endogenous transmitter; a specific mechanism usually exists for removing the substance from the synaptic cleft. These are similar, but not identical, to the classic textbook on neurotransmitters: a neurotransmitter “should be synthesized and released presynaptically; it must mimic the action of the endogenous compound that is release on nerve stimulation; and where possible, a pharmacological identity is required where drugs that either potentiate or block postsynaptic responses to the endogenously released agent also act identically to the suspected neurotransmitter that is administered” 222. The pharmacological criterion is listed in another textbook 223.

Some authors note difficulties of establishing a CNS neurotransmitter. For example, a specialized neurotransmitter book states that “the candidate neurotransmitter should be present in the presynaptic terminal, be released when the presynaptic terminal is active, and when applied experimentally, induce faithful responses in the postsynaptic neuron. In practice, since central nervous system neurons continuously integrate diverse excitations and inhibitions, the last criterion is relaxed to demonstrating merely changes in such activity” 224.

Solomon Snyder, a leading scientist of classic neurotransmitters, neuropeptides and their receptors, wrote that “designating a molecule as a transmitter depends on the criteria employed, the most common of which are that the substance is synthesized in neurons, released by their terminals, mimics the effects of physiologic neurotransmission and possess a mechanism for inactivation. However, with each new candidate the rules have been modified and broadened” 225.

Evidence supporting Cr as a neurotransmitter

16 small molecules have been listed as neurotransmitters in the classic textbook 70,71. Among them, adenosine, arachidonic acid, nitric oxide and carbon monoxide do not meet all four criteria, at the present. Cr appears to be better than these in meeting the criteria for a central neurotransmitter.

The results obtained by us in this paper have satisfied the criteria of Robinson 224 for Cr to be a CNS neurotransmitter.

Our results have also satisfied the criteria of Purves et al. 67,68, because the presence of postsynaptic receptors can be inferred by postsynaptic responses.

The four criteria of Snyder and colleagues 225 have been mostly met but the physiological neurotransmission would require more research because a specific synapse(s) would have to be defined and studied for putative creatinergic neurotransmission. This can take much longer in the CNS than the PNS.Some commonly accepted neurotransmitters have never satisfied this criterion in a strict sense. The mechanism of Cr removal criterion is met not only by the Cr uptake in brain slices 220 and in synaptosomes 221, but also by our demonstration that SLC6A8 is required for synaptosome uptake of Cr.

The four criteria of Kandel et al.70,71 are mostly satisfied with some details requiring further research. The synthesis requirement is usually not strict because there are transmitters synthesized in some cells and transported into others where they function as transmitters. Our discovery of Cr in SVs can replace the synthesis requirement because the presence in neuronal SVs provide sufficient evidence that Cr is located in the right location to function as a neurotransmitter. The level of Cr in SVs is higher than those of ACh and 5-HT (Figs. 1 and 2). The amount of released Cr is in the same order of magnitude as those of Glu and GABA (Figs. 5 and 6). The criterion of a specific mechanism of removal was met by Cr uptake experiments in slices 220 and in synaptosomes 221, and further strengthened by our finding of SLC6A8 involvement in synaptosome uptake of Cr (Figure 8).

Here we report that Cr, at a concentration comparable to classical neurotransmitters, inhibits pyramidal neurons in specific regions of the mouse brain, with approximate 1/3 of pyramidal neurons responding to 100 μM Cr (Figure 7). In previous reports, 100 M to 2 mM of GABA226–228, 50 μM to 2 mM of Glu 229,230, 1 to 100μM of DA229, 0.1-100 μM of 5-HT231–236 were bath applied to investigate the physiological functions of neurotransmitters. Our results revealed that, when bath applied, Cr could inhibit cortical neurons at 100 μM within several minutes, with a time course similar to that of 5-HT237 234 236 and DA 229, but significantly slower than that of Glu230, GABA48,227 and 5-HT233. We speculate that Cr may act on G-protein coupled receptors (GPCRs), rather than the fast acting ligand-gated ion channels, such as AMPA or NMDA receptors for Glu and GABAA receptor for GABA. There have been previous reports of Cr effects on neurons, including Cr as a partial agonist for GABAA receptors 238–241. These effects require very high concentrations of Cr (in the ten millimolar range). There was also a report of the opposite effect: that Cr (at a concentration above 500 μM) increased neuronal excitability through NMDA receptors after incubation for 60 mins, with a time course significantly slower than those of classic neurotransmitters 242.

With much longer history of research, ACh and 5-HT now have more evidence in other aspects than Cr as a central transmitter, especially because there are many agonists and antagonists for ACh and 5-HT to prove an additional criterion that is required in some 222,223, but not the majority of, textbooks for a neurotransmitter. The pharmacology criterion will take some time and efforts, because, so far, no efforts have ever been made to find agonists or antagonists for Cr.

Implications of SLC6A8 and Cr

It is notable that SLC6A8 belongs to the neurotransmitter transporter family, with multiple members already shown to transport neurotransmitters 65,66, 90–99.

The uptake experiments by others and us indicate that SLC6A8 transports Cr into neurons within the brain. AGAT is also expressed in the brain, but in cells not expressing SLC6A8 243,244. Cr and its precursor were thought to be transported between different cells in the nervous system. When SLC6A8 was completely missing, such as in homozygous SLC6A8 deficient patients, Cr treatment was not effective. But, if SLC6A8 was partially active, Cr was effective 245. Intractable epilepsy in a female with heterozygous SLC6A8 mutation was completely treated by Cr 246. Our data of inhibitory effect of creatine on cortical neurons might provide a new mechanism to its anti-epileptic activity 247.

The absence of SLC6A8 expression in astrocytes whose endfeet lining microcapillary endothelial cells (MCEC) form the blood-brain barrier (BBB) indicates that Cr in the brain does not rely on import from the periphery and is instead mainly synthesized in the brain 80,82,248. SLC6A8 functions within the brain to transport Cr and its precursors, not as a major contributor of Cr transport across the BBB. It is thought to mediate Cr uptake into the presynaptic terminal, based on studies of synaptosomes (Figure 8) 221.

Cr is known to have effects other than an energy source, and Cr supplement has been thought to be beneficial for children, pregnant and lactating women, and old people 79,249. Cr has been reported to improve human mental performance 250–255. Cr has been used as potential treatment in animal models of neurodegenerative diseases 256,257.

Our work will stimulate further research to distinguish which of the previously suspected effects of Cr is not attributed to its role as an energy storage, but can be attributed to its role as a neurotransmitter.

Search for new neurotransmitters

Our work may stimulate the search for more neurotransmitters. Our discovery indicates that the hunt for neurotransmitters stopped decades ago because of technical difficulties, not due to absence of more neurotransmitters. The fact that most of the known small molecule neurotransmitters have been found because of their peripheral effects also argues that what is missing is the concerted efforts to uncover central neurotransmitters with no peripheral effects. New neurotransmitters may be discovered from candidates which have been long suspected, and from previously unsuspected molecules or even previously unknown molecules.

Innovative approaches should be taken to uncover molecules with no previous suspicions or hints. Highly purified SVs, SVs from different regions of the brain, SVs with specific SLCs offer some of the starting points for future research.

Supporting information

eLife Assessment

This study presents valuable observations on a potential role of creatine (Cr) as a novel neurotransmitter. While the data provide compelling evidence that Cr is present in synaptic vesicles, the evidence that Cr is synaptically released and binds to a post-synaptic receptor is incomplete, as no receptor is described. With further evidence, these results will be of wide interest to the field of neuroscience.

Reviewer #1 (Public Review):

This is an interesting and somewhat unusual paper supporting the idea that creatine is a neurotransmitter in the central nervous system of vertebrates. The idea is not entirely new, and the authors carefully weigh the evidence, both past and newly acquired, to make their case. The strength of the paper lies in the importance of the potential discovery - as the authors point out, creatine ticks more boxes on criteria of neurotransmitters than some of the ones listed in textbooks - and the list of known transmitters (currently 16) certainly is textbook material. A further strength of the manuscript is the careful consideration of a list of criteria for transmitters and newly acquired evidence for four of these criteria: 1. evidence that creatine is stored in synaptic vesicles, 2. mutants for creatine synthesis and a vesicular transporter show reduced storage and release of creatine, 3. functional measurement that creatine release has an excitatory or inhibitory (here inhibitory) effect in vivo, and 4. ATP-dependence. The key weakness of the paper is that there is no single clear 'smoking gun', like a postsynaptic creatine receptor, that would really demonstrate the function as a transmitter. Instead, the evidence is of a cumulative nature, and not all bits of evidence are equally strong. On balance, I found the path to discovery and the evidence assembled in this manuscript to establish a clear possibility, positive evidence, and to provide a foundation for further work in this direction.

Reviewer #2 (Public Review):

Summary:

Bian et al studied creatine (Cr) in the context of central nervous system (CNS) function. They detected Cr in synaptic vesicles purified from mouse brains with anti-Synaptophysin using capillary electrophoresis-mass spectrometry. Cr levels in the synaptic vesicle fraction were reduced in mice lacking the Cr synthetase AGAT, or the Cr transporter SLC6A8. They provide evidence for Cr release within several minutes after treating brain slices with KCl. This KCl-induced Cr release was partially calcium-dependent and was attenuated in slices obtained from AGAT and SLC6A8 mutant mice. Cr application also decreased the excitability of cortical pyramidal cells in one third of the cells tested. Finally, they provide evidence for SLC6A8-dependent Cr uptake into synaptosomes, and ATP-dependent Cr loading into synaptic vesicles. Based on these data, the authors propose that Cr may act as a neurotransmitter in the CNS.

Strengths:

1. A major strength of the paper is the broad spectrum of tools used to investigate Cr.

2. The study provides strong evidence that Cr is present in/loaded into synaptic vesicles.

Weaknesses:

(in sequential order)

1. Are Cr levels indeed reduced in Agat-/-? The decrease in Cr IgG in Agat-/- (and Agat+/-) is similar to the corresponding decrease in Syp (Fig. 3B). What is the explanation for this? Is the decrease in Cr in Agat-/- significant when considering the drop in IgG? The data should be normalized to the respective IgG control.

2. The data supporting that depolarization-induced Cr release is SLC6A8 dependent is not convincing because the relative increase in KCl-induced Cr release is similar between SLC6A8-/Y and SLC6A8+/Y (Fig. 5D). The data should be also normalized to the respective controls.

3. The majority (almost 3/4) of depolarization-induced Cr release is Ca2+ independent (Fig. 5G). Furthermore, KCl-induced, Ca2+-independent release persists in SLC6A8-/Y (Fig. 5G). What is the model for Ca2+-independent Cr release? Why is there Ca2+-independent Cr release from SLC6A8 KO neurons?

How does this relate to the prominent decrease in Ca2+-dependent Cr release in SLC6A8-/Y (Fig. 5G)? They show a prominent decrease in Cr control levels in SLC6A8-/Y in Fig. 5D. Were the data shown in Fig. 5D obtained in the presence or absence of Ca2+? Could the decrease in Ca2+-dependent Cr release in SLC6A8-/Y (Fig. 5G) be due to decreased Cr baseline levels in the presence of Ca2+ (Fig. 5D)?

4. Cr levels are strongly reduced in Agat-/- (Fig. 6B). However, KCl-induced Cr release persists after loss of AGAT (Fig. 6B). These data do not support that Cr release is Agat dependent.

5. The authors show that Cr application decreases excitability in ~1/3 of the tested neurons (Fig. 7). How were responders and non-responders defined? What justifies this classification? The data for all Cr-treated cells should be pooled. Are there indeed two distributions (responders/non-responders)? Running statistics on pre-selected groups (Fig. 7H-J) is meaningless. Given that the effects could be seen 2-8 minutes after Cr application - at what time points were the data shown in Fig. 7E-J collected? Is the Cr group shown in Fig. 7F significantly different from the control group/wash?

6. Indirect effects: The phenotypes could be partially caused by indirect effects of perturbing the Cr/PCr/CK system, which is known to play essential roles in ATP regeneration, Ca2+ homeostasis, neurotransmission, intracellular signaling systems, axonal and dendritic transport... Similarly, high GAMT levels were reported for astrocytes (e.g., Schmidt et al. 2004; doi: 10.1093/hmg/ddh112), and changes in astrocytic Cr may underlie the phenotypes. Cr has been also reported to be an osmolyte: a hyperosmotic shock of astrocytes induced an increase in Cr uptake, suggesting that Cr can work as a compensatory osmolyte (Alfieri et al. 2006; doi: 10.1113/jphysiol.2006.115006). Potential indirect effects are also consistent with a trend towards decreased KCl-induced GABA (and Glutamate) release in SLC6A8-/Y (Fig. 5C). These indirect effects may in part explain the phenotypes seen after perturbing Agat, SLC6A8, and should be thoroughly discussed.

7. As stated by the authors, there is some evidence that Cr may act as a co-transmitter for GABAA receptors (although only at high concentrations). Would a GABAA blocker decrease the fraction of cells with decreased excitability after Cr exposure?

8. The statement "Our results have also satisfied the criteria of Purves et al. 67,68, because the presence of postsynaptic receptors can be inferred by postsynaptic responses." (l.568) is not supported by the data and should be removed.

Reviewer #3 (Public Review):

SUMMARY:

The manuscript by Bian et al. promotes the idea that creatine is a new neurotransmitter. The authors conduct an impressive combination of mass spectrometry (Fig. 1), genetics (Figs. 2, 3, 6), biochemistry (Figs. 2, 3, 8), immunostaining (Fig. 4), electrophysiology (Figs. 5, 6, 7), and EM (Fig. 8) in order to offer support for the hypothesis that creatine is a CNS neurotransmitter.

STRENGTHS:

There are many strengths to this study.

• The combinatorial approach is a strength. There is no shortage of data in this study.

• The careful consideration of specific criteria that creatine would need to meet in order to be considered a neurotransmitter is a strength.

• The comparison studies that the authors have done in parallel with classical neurotransmitters are helpful.

• Demonstration that creatine has inhibitory effects is another strength.

• The new genetic mutations for Slc6a8 and AGAT are strengths and potentially incredibly helpful for downstream work.

WEAKNESSES:

• Some data are indirect. Even though Slc6a8 and AGAT are helpful sentinels for the presence of creatine, they are not creatine themselves. Therefore, the conclusions that are drawn should be circumspect.

• Regarding Slc6a8, it seems to work only as a reuptake transporter - not as a transporter into SVs. Therefore, we do not know what the transporter is.

• Puzzlingly, Slc6a8 and AGAT are in different cells, setting up the complicated model that creatine is created in one cell type and then processed as a neurotransmitter in another.

• No candidate receptor for creatine has been identified postsynaptically.

• Because no candidate receptor has been identified, is it possible that creatine is exerting its effects indirectly through other inhibitory receptors (e.g., GABAergic Rs)?

• More broadly, what are the other possibilities for roles of creatine that would explain these observations other than it being a neurotransmitter? Could it simply be a modifier that exists in the SVs (lots of molecules exist in SVs)?

• The biochemical studies are helpful in terms of comparing relevant molecules (e.g., Figs. 8 and S1), but the images of the westerns are all so fuzzy that there are questions about processing and the accuracy of the quantification.

APPRAISAL OF WHETHER THE AUTHORS ACHIEVED THEIR AIMS AND WHETHER THE RESULTS SUPPORT THE CONCLUSIONS:

There are several criteria that define a neurotransmitter. The authors nicely delineated many criteria in their discussion, but it is worth it for readers to do the same with their own understanding of the data.

By this reviewer's understanding (and the Purves' textbook definition) a neurotransmitter: 1) must be present within the presynaptic neuron and stored in vesicles; 2) must be released by depolarization of the presynaptic terminal; 3) must require Ca2+ influx upon depolarization prior to release; 4) must bind specific receptors present on the postsynaptic cell; 5) exogenous transmitter can mimic presynaptic release; 6) there exists a mechanism of removal of the neurotransmitter from the synaptic cleft.

For a paper to claim that the work has identified a new neurotransmitter, several of these criteria would be met - and the paper would acknowledge in the discussion which ones have not been met. For this particular paper, this reviewer finds that condition 1 is clearly met.

Conditions 2 and 3 seem to be met by electrophysiology, but there are caveats here. High KCl stimulation is a blunt instrument that will depolarize absolutely everything in the prep all at once and could result in any number of non-specific biological reactions as a result of K+ rushing into all neurons in the prep. Moreover, the results in 0 Ca2+ are puzzling. For creatine (and for the other neurotransmitters), why is there such a massive uptick in release, even when the extracellular saline is devoid of calcium?

Condition 4 is not discussed in detail at all. In the discussion, the authors elide the criterion of receptors specified by Purves by inferring that the existence of postsynaptic responses implies the existence of receptors. True, but does it specifically imply the existence of creatinergic receptors? This reviewer does not think that is necessarily the case. The authors should be appropriately circumspect and consider other modes of inhibition that are induced by activation or potentiation of other receptors (e.g., GABAergic or glycinergic).

Condition 5 may be met, because the authors applied exogenous creatine and observed inhibition (Fig. 7). However, this is tough to know without understanding the effects of endogenous release of creatine. if they were to test if the absence of creatine caused excess excitation (at putative creatinergic synapses), then that would be supportive of the same.

For condition 6, the authors made a great effort with Slc6a8. This is a very tough criterion to understand for many synapses and neurotransmitters.

DISCUSSION OF THE LIKELY IMPACT OF THE WORK:

In terms of fundamental neuroscience, the story would be impactful if proven correct. There are certainly more neurotransmitters out there than currently identified.

The impact as framed by the authors in the abstract and introduction for intellectual disability is uncertain (forming a "new basis for ID pathogenesis") and it seems quite speculative beyond the data in this paper.

Author Response

The following is the authors’ response to the original reviews.

Reviewer #1 (Public Review):

This is an interesting and somewhat unusual paper supporting the idea that creatine is a neurotransmitter in the central nervous system of vertebrates. The idea is not entirely new, and the authors carefully weigh the evidence, both past and newly acquired, to make their case. The strength of the paper lies in the importance of the potential discovery - as the authors point out, creatine ticks more boxes on criteria of neurotransmitters than some of the ones listed in textbooks - and the list of known transmitters (currently 16) certainly is textbook material. A further strength of the manuscript is the careful consideration of a list of criteria for transmitters and newly acquired evidence for four of these criteria: 1. evidence that creatine is stored in synaptic vesicles, 2. mutants for creatine synthesis and a vesicular transporter show reduced storage and release of creatine, 3. functional measurement that creatine release has an excitatory or inhibitory (here inhibitory) effect in vivo, and 4. ATP-dependence. The key weakness of the paper is that there is no single clear 'smoking gun', like a postsynaptic creatine receptor, that would really demonstrate the function as a transmitter. Instead, the evidence is of a cumulative nature, and not all bits of evidence are equally strong. On balance, I found the path to discovery and the evidence assembled in this manuscript to establish a clear possibility, positive evidence, and to provide a foundation for further work in this direction.

it is notable that, historically, no neurotransmitter has ever been established in a single paper. While creatine will not be an exception, data presented in this paper are more than any previous paper in demonstrating the possibility of a new neurotransmitter. However, we added an entire paragraph in the Discussion part about differences between Cr and classic neurotransmitters such as Glu, beginning with the absence of a molecularly defined receptor at this point and the Ca2+ independent component of Cr release induced by extracellular K+.

We appreciate the reviewer for noting that evidence obtained by us now support that creatine satisfies all 4 criteria of transmitters.

We respectively disagree the point about a smoking gun: any of these four is a smoking gun, while the satisfication of all 4 is quite strong, more than a smoking gun.

We find it disagreeable that a receptor “would really demonstrate the function of a transmitter”. Textbook criteria for a transmitter usually require postsynaptic responses, not a molecularly defined receptor. A molecularly defined receptor for many of the known transmitters required many years of work, while they were accepted as transmitters before their receptors were finally molecularly defined. As long as there is a postsynaptic response, there is of course a receptor, though its molecular properties should be further studied. For examples, responses to choline were discovered in 1900 (Hunt, Am J Physiol 3, xviii-xix, 1900), those to acetylcholine in 1906 (Hunt and Taveau, Br Med J 2:1788-1789, 1906), those to supradrenal glands before 1894 (Oliver and Schäfer, J Physiol 18:230-276 1895). Henry Dale was awarded a Nobel prize in 1936 partly for his work on acetylcholine. Receptors for acetylcholine and noradrenaline were not molecularly defined until the 1970s and 1980s. Before then, they were only known by mediating responses to natural transmitters and synthesized chemicals.

There were two previous reports that creatine could be taken into brain slices (Almeida et al., 2006) or synaptosomes (Peral, Vázquez-Carretero and Ilundain, 2010). These were used by the reviewer to argue that the idea of creatine as a neurotransmitter “is not entirely new”. However, no one has followed up these studies for 10 years, thus they would not be considered as good smoking guns. While we have reproduced the synaptosome uptake result (together with our new finding that this uptake was dependent on SLC6A8), it should be noted that uptake of molecules into synaptosomes is not absolutely required for a neurotransmitter because degradation of a transmitter is equally valid. Furthermore, molecules required synaptically but not as a transmitter can also be transported into the synaptic terminal.

Our detection of Cr in the synaptic vesicles provides much stronger evidence supporting its importance. If a smoking gun is important, the detection of creatine in the SVs is the best smoking gun, whose discovery in fact was the reason leading us to study its release, postsynaptic responses as well as repeating the uptake experiment with genetic mutants.

Reviewer #2 (Public Review):

Summary:

Bian et al studied creatine (Cr) in the context of central nervous system (CNS) function. They detected Cr in synaptic vesicles purified from mouse brains with anti-Synaptophysin using capillary electrophoresis-mass spectrometry. Cr levels in the synaptic vesicle fraction were reduced in mice lacking the Cr synthetase AGAT, or the Cr transporter SLC6A8. They provide evidence for Cr release within several minutes after treating brain slices with KCl. This KCl-induced Cr release was partially calcium-dependent and was attenuated in slices obtained from AGAT and SLC6A8 mutant mice. Cr application also decreased the excitability of cortical pyramidal cells in one third of the cells tested. Finally, they provide evidence for SLC6A8-dependent Cr uptake into synaptosomes, and ATP-dependent Cr loading into synaptic vesicles. Based on these data, the authors propose that Cr may act as a neurotransmitter in the CNS.

Strengths:

1. A major strength of the paper is the broad spectrum of tools used to investigate Cr.

1. The study provides strong evidence that Cr is present in/loaded into synaptic vesicles.

Weaknesses:

(in sequential order)

1. Are Cr levels indeed reduced in Agat-/-? The decrease in Cr IgG in Agat-/- (and Agat+/-) is similar to the corresponding decrease in Syp (Fig. 3B). What is the explanation for this? Is the decrease in Cr in Agat-/- significant when considering the drop in IgG? The data should be normalized to the respective IgG control.

We measured the Cr concentration in the whole brain lysates using Creatine Assay Kit (Sigma, MAK079). Cr levels in the brain were reduced in Agat-/- mice. The Cr concentration in AGAT-/- mice was reduced to about 1/10 of AGAT+/+ and AGAT+/- mice (Author response image 1).

Cr concentration in brain from AGAT+/+, AGAT+/- and AGAT-/- mice (n=5 male mice for each group).

*, p<0.05, ***, p<0.001, one-way ANOVA with Tukey’s correction.

As pointed by the reviewer, the decrease in Cr IgG in Agat-/- seems similar to the corresponding decrease in Syp (Fig. 3B in the paper). Cr pulled down by IgG was 0.46 ± 0.04, 0.37 ± 0.06 and 0.17 ±0.03 pmol/μg anti-syp antibody for Agat+/+, Agat+/-, and Agat-/- mice respectively. There was a trend of reduction Cr IgG in Agat-/-, however, there were no statistically significant differences between Agat-/- and Agat+/+, or between Agat-/- and Agat+/-, as determined by one-way ANOVA (Fig. 3B in the paper). Due to the fact that Agat-/- reduced Cr concentration in the brain, we speculate that the apparent drop in Cr pulled down by IgG may have partially resulted from the overall reduction of Cr content in the brain.

The absolute content of Cr pulled down by Syp in Agat-/- mice was reduced to 21.6% of Agat+/+ mice and 23.6% of Agat+/- mice (Fig. 3B in the paper). As suggested by the reviewer, we normalized the Cr pulled down by Syp to the respective IgG control (Author response image 2). The normalized Cr content in AGAT-/- mice has a tendency to decrease, but not statistically significant, as compared to Agat+/+ and Agat+/- mice (n=10 for each group, one-way ANOVA).

Normalized Cr content in brain from AGAT+/+, AGAT+/- and AGAT-/- mice (n=10 for each group).

Cr pulled down by anti-Syp antibody was normalized to that of IgG.

1. The data supporting that depolarization-induced Cr release is SLC6A8 dependent is not convincing because the relative increase in KCl-induced Cr release is similar between SLC6A8-/Y and SLC6A8+/Y (Fig. 5D). The data should be also normalized to the respective controls.

As suggested by the reviewer, we normalized the Cr release during KCl stimulation to the baseline (Author response image 3). The ratio of Cr release evoked by high KCl stimulation to the baseline was similar in WT and Slc6a8 knockouts. This suggests that Cr is not released through SLC6A8 transporter.

Normalized Cr release from slices from Slc6a8+/Y and Slc6a8-/Y mice (n=7 slices for each group).

Cr released evoked by high KCl stimulation was normalized to baseline.

However, without Slc6a8, KCl-induced release of Cr was significantly reduced (Figure 5D in the paper). This is because Slc6a8 is a transporter to Cr uptake into synaptic terminals (Figure 5D and 8C in the paper). Therefore, Cr content in SVs (Figure 2C in the paper) indirectly reduced Cr release.

1. The majority (almost 3/4) of depolarization-induced Cr release is Ca2+ independent (Fig. 5G). Furthermore, KCl-induced, Ca2+-independent release persists in SLC6A8-/Y (Fig. 5G). What is the model for Ca2+-independent Cr release? Why is there Ca2+-independent Cr release from SLC6A8 KO neurons? How does this relate to the prominent decrease in Ca2+-dependent Cr release in SLC6A8-/Y (Fig. 5G)? They show a prominent decrease in Cr control levels in SLC6A8-/Y in Fig. 5D. Were the data shown in Fig. 5D obtained in the presence or absence of Ca2+? Could the decrease in Ca2+-dependent Cr release in SLC6A8-/Y (Fig. 5G) be due to decreased Cr baseline levels in the presence of Ca2+ (Fig. 5D)?

These are interesting questions that, at this point, could only be answered by references to literature. For example, one possibility was that Ca2+-independent Cr release might occurs in glia, since as pointed by the reviewer in Point 6, high GAMT levels were reported for astrocytes and oligodendrites (Schmidt et al. 2004; Rosko et al. 2023). As reported, other neuromodulators such as taurine can be released from astrocytes (Philibert, Rogers, and Dutton 1989) or slices (Saransaari and Oja 2006) in Ca2+ independent manner. In addition, in the absence of potassium stimulation, Ca2+ depletion lead to increased release of taurine in cultured astrocytes (Takuma et al. 1996) or in striatum in vivo (Molchanova, Oja, and Saransaari 2005). Similarly, in SLC6A8 KO slices, Ca2+ depletion (Figure 5G) also increased creatine baseline levels as compared to that in normal ACSF (Figure 5D). Another possibility was that Ca2+-independent Cr release might occurs in neurons lacking SLC6a8 expression.

As mentioned in the paper, data shown in Figure 5D was obtained in the presence Ca2+. Reduction of Ca2+-dependent Cr release evoked by potassium in SLC6A8-/Y (Figure 5G) may be due to decreased Cr baseline levels in the presence of Ca2+ and reduced Cr in synaptic vesicles (Figure 5D).

1. Cr levels are strongly reduced in Agat-/- (Figure 6B). However, KCl-induced Cr release persists after loss of AGAT (Figure 6B). These data do not support that Cr release is Agat dependent.

Although KCl-induced Cr release persisted in AGAT-/- mutants, it was dropped to 11.6% of WT mice (Figure 6B). AGAT is not directly involved in the release, but required for providing sufficient Cr.

1. The authors show that Cr application decreases excitability in ~1/3 of the tested neurons (Figure 7). How were responders and non-responders defined? What justifies this classification? The data for all Cr-treated cells should be pooled. Are there indeed two distributions (responders/non-responders)? Running statistics on pre-selected groups (Figure 7H-J) is meaningless. Given that the effects could be seen 2-8 minutes after Cr application - at what time points were the data shown in Figure 7E-J collected? Is the Cr group shown in Figure 7F significantly different from the control group/wash?

The responders were defined by three criteria: (1) When Cr was applied, the rheobase was increased as compared to both control and wash conditions. (2) The number of total evoked spikes was decreased during Cr application than both control and wash. (3) The number of total evoked spikes was decreased at least by 10% than control or wash.

For all the individual responders, when Cr was applied, the rheobase was increased (Figure 7E and 7F). While in individual non-responders, the rheobase was either identical to both control and wash (n=19/35), identical to either control or wash (n=11/35), between control and wash (n=2/35) or smaller than both control and wash (n=3/35) following Cr application. Thus, the responders and non-responders were separatable. When the rheobase data were pulled together, many points were overlapped, so we did not pull the data here.

As suggested, we pulled the data of the ratio of spike changes in response to 100 μM Cr application for all neurons together (Author response image 4). Evoked spikes of non-responders were typically (34/35) changed in the range of -10% to 10%.

Relative changes of total evoked spikes in response to 100 μM Cr.

Responders are represented by red dots and non-responders by black dots. Dashed black line indicates 10%. Relative change = (Cr-(Control +wash)/2)/((Control +wash)/2)*100%.

In Figure 7E-J, we collected data at time points when the maximal response was reached. The Cr group shown in Figure 7F was indeed significantly different from the control group/wash (p<0.05, paired t test, for data points collected under 75-500 pA current injection).

1. Indirect effects: The phenotypes could be partially caused by indirect effects of perturbing the Cr/PCr/CK system, which is known to play essential roles in ATP regeneration, Ca2+ homeostasis, neurotransmission, intracellular signaling systems, axonal and dendritic transport... Similarly, high GAMT levels were reported for astrocytes (e.g., Schmidt et al. 2004; doi: 10.1093/hmg/ddh112), and changes in astrocytic Cr may underlie the phenotypes. Cr has been also reported to be an osmolyte: a hyperosmotic shock of astrocytes induced an increase in Cr uptake, suggesting that Cr can work as a compensatory osmolyte (Alfieri et al. 2006; doi: 10.1113/jphysiol.2006.115006). Potential indirect effects are also consistent with a trend towards decreased KCl-induced GABA (and Glutamate) release in SLC6A8-/Y (Figure 5C). These indirect effects may in part explain the phenotypes seen after perturbing Agat, SLC6A8, and should be thoroughly discussed.

We discussed the possibility of creatine/phosphocreatine as non-transmitters in discussion part. We added the possibility of astrocytic Cr in discussion part. KCl-induced GABA (and Glutamate) release in SLC6A8-/Y (Figure 5C) was not significant.

1. As stated by the authors, there is some evidence that Cr may act as a co-transmitter for GABAA receptors (although only at high concentrations). Would a GABAA blocker decrease the fraction of cells with decreased excitability after Cr exposure?

We performed another experiment in CA1 pyramidal neurons in hippocampus showing that Cr at 100 μM did not change GABAergic neurotransmission (n=8, Author response image 5). Inhibitory postsynaptic currents (IPSCs) recorded in the presence of glutamate receptor blockers (10 μM APV and 10 μM CNQX) were not changed by 100 μM creatine in hippocampal CA1 pyramidal neurons (Bgroup data of IPSC frequency (B) and amplitude (C) averaged in 1 min duration). These did not support Cr activation of GABAA receptors.

IPSCs recorded in in hippocampal CA1 pyramidal neurons.

(A) representative raw traces before (Control), during (Creatine) and after (Wash) the application of 100 μM creatine. (B&C) group data of IPSC frequency (B) and amplitude (C) averaged in 1 min duration.

1. The statement "Our results have also satisfied the criteria of Purves et al. 67,68, because the presence of postsynaptic receptors can be inferred by postsynaptic responses." (l.568) is not supported by the data and should be removed.

We have deleted this sentence, though what could mediate postsynaptic responses other than receptors?

Reviewer #3 (Public Review):

SUMMARY:

The manuscript by Bian et al. promotes the idea that creatine is a new neurotransmitter. The authors conduct an impressive combination of mass spectrometry (Fig. 1), genetics (Figs. 2, 3, 6), biochemistry (Figs. 2, 3, 8), immunostaining (Fig. 4), electrophysiology (Figs. 5, 6, 7), and EM (Fig. 8) in order to offer support for the hypothesis that creatine is a CNS neurotransmitter.

We thank the reviewer for the summary.

STRENGTHS:

There are many strengths to this study.

• The combinatorial approach is a strength. There is no shortage of data in this study.

• The careful consideration of specific criteria that creatine would need to meet in order to be considered a neurotransmitter is a strength.

• The comparison studies that the authors have done in parallel with classical neurotransmitters are helpful.

• Demonstration that creatine has inhibitory effects is another strength.

• The new genetic mutations for Slc6a8 and AGAT are strengths and potentially incredibly helpful for downstream work.

WEAKNESSES:

• Some data are indirect. Even though Slc6a8 and AGAT are helpful sentinels for the presence of creatine, they are not creatine themselves. Therefore, the conclusions that are drawn should be circumspect.

SLC6A8 and AGAT mutants are not essential for Cr’s role as a neurotransmitter.

• Regarding Slc6a8, it seems to work only as a reuptake transporter - not as a transporter into SVs. Therefore, we do not know what the transporter is.

Indeed, SLC6A8 is only a transporter on the cytoplasmic membrane, not a transporter on synaptic vesicles. We have shown biochemistry here, and we have unpublished data that showed other SLCs on SVs, which did not include SLC6A8.

• Puzzlingly, Slc6a8 and AGAT are in different cells, setting up the complicated model that creatine is created in one cell type and then processed as a neurotransmitter in another.

• No candidate receptor for creatine has been identified postsynaptically.

• Because no candidate receptor has been identified, is it possible that creatine is exerting its effects indirectly through other inhibitory receptors (e.g., GABAergic Rs)?

As shown in our response to Question 7 of Reviewer 2, Cr did not exert its effects through inhibitory GABAA receptors.

• More broadly, what are the other possibilities for roles of creatine that would explain these observations other than it being a neurotransmitter? Could it simply be a modifier that exists in the SVs (lots of molecules exist in SVs)?

We discussed the possibility of a non-transmitter role for creatine/phosphocreatine in discussion part.

• The biochemical studies are helpful in terms of comparing relevant molecules (e.g., Figs. 8 and S1), but the images of the westerns are all so fuzzy that there are questions about processing and the accuracy of the quantification.

Multiple members (>4) have carried out SV purifications repeatedly over the last decade in our group, we are highly confident of SV purifications presented in Figs. 8 and S1.

There are several criteria that define a neurotransmitter. The authors nicely delineated many criteria in their discussion, but it is worth it for readers to do the same with their own understanding of the data.

By this reviewer's understanding (and the Purves' textbook definition) a neurotransmitter: 1) must be present within the presynaptic neuron and stored in vesicles; 2) must be released by depolarization of the presynaptic terminal; 3) must require Ca2+ influx upon depolarization prior to release; 4) must bind specific receptors present on the postsynaptic cell; 5) exogenous transmitter can mimic presynaptic release; 6) there exists a mechanism of removal of the neurotransmitter from the synaptic cleft.

6 criteria seem to be only required by the reviewer. As discussed in our Discussion part, Purves’ textbook did not list 6 criteria but only three criteria, “the substance must be present within the presynaptic neuron; the substance must be released in response to presynaptic depolarization, and the release must be Ca2+ dependent; specific receptors for the substance be present on the postsynaptic cell” (Purves et al., 2001, 2016).

Kandel et al. (2013, 2021) listed 4 criteria for a neurotransmitter: “it is synthesized in the presynaptic neuron; it is present within vesicles and is released in amounts sufficient to exert a defined action on the postsynaptic neuron or effector organ; when administered exogenously in reasonable concentrations it mimics the action of the endogenous transmitter; a specific mechanism usually exists for removing the substance from the synaptic cleft”.

While we agree that any neuroscientist can have his/her own criteria, it is more reasonable to accept the textbooks that have been widely read for decades.

For a paper to claim that the work has identified a new neurotransmitter, several of these criteria would be met - and the paper would acknowledge in the discussion which ones have not been met. For this particular paper, this reviewer finds that condition 1 is clearly met.

Conditions 2 and 3 seem to be met by electrophysiology, but there are caveats here. High KCl stimulation is a blunt instrument that will depolarize absolutely everything in the prep all at once and could result in any number of non-specific biological reactions as a result of K+ rushing into all neurons in the prep. Moreover, the results in 0 Ca2+ are puzzling. For creatine (and for the other neurotransmitters), why is there such a massive uptick in release, even when the extracellular saline is devoid of calcium?

To avoid the disadvantage of high KCl stimulation, we performed optogenetic experiments recently, with encouraging preliminary data. We do not know the source of Ca2+-independent release of Cr and neurotransmitters, though astrocytes are a possibility.

Condition 4 is not discussed in detail at all. In the discussion, the authors elide the criterion of receptors specified by Purves by inferring that the existence of postsynaptic responses implies the existence of receptors. True, but does it specifically imply the existence of creatinergic receptors? This reviewer does not think that is necessarily the case. The authors should be appropriately circumspect and consider other modes of inhibition that are induced by activation or potentiation of other receptors (e.g., GABAergic or glycinergic).

Our results did not support Cr stimulation of inhibitory GABAA receptors (see our answer to Point 7 in of Reviewer 2).

Condition 5 may be met, because the authors applied exogenous creatine and observed inhibition (Fig. 7). However, this is tough to know without understanding the effects of endogenous release of creatine. if they were to test if the absence of creatine caused excess excitation (at putative creatinergic synapses), then that would be supportive of the same.

After the submission of our manuscript, we found a recent paper showing that slc6a8 knockout led to increased excitation in pyramidal neurons in the prefrontal cortex (PFC), with increased firing frequency (Ghirardini et al., 2023). Because we have shown that slc6a8 knockout would cause decrease of Cr in SVs (Figure 2 in our paper), this result provide the evidence described as Condition 5 of this reviewer: that decrease of Cr in SVs led to excess excitation.

For condition 6, the authors made a great effort with Slc6a8. This is a very tough criterion to understand for many synapses and neurotransmitters.

In terms of fundamental neuroscience, the story would be impactful if proven correct. There are certainly more neurotransmitters out there than currently identified.

The impact as framed by the authors in the abstract and introduction for intellectual disability is uncertain (forming a "new basis for ID pathogenesis") and it seems quite speculative beyond the data in this paper.

We deleted this sentence.

Reviewer #1 (Recommendations For The Authors):

To strengthen the manuscript, I suggest the following considerations:

1. The key missing evidence to my mind is a receptor - but this is clearly outside the scope of this paper. Yet, I am surprised that in the list of criteria for neurotransmitters in general there is no mention of a receptor. Furthermore, many receptors have been identified through receptor agonists or antagonists, like neurotoxins or drugs. The authors do not talk about putative receptors except for a sentence in the discussion where they speculate on a GPCR. There are numerous GPCR agonists and antagonists, which may be a long-shot, or something even a bit more designed based on knowledge about creatine? I do not think the publication of this manuscript should have been made dependent on finding an agonist or antagonist of this specific unknown receptor (if it exists), but it would be good to have at least some leads on this from the authors what has been tried or what could be done? How about a manipulation of G-protein-coupled signal transduction to support the idea that there IS such a GPCR? There may be a real opportunity here to test existing compounds in wild type, the slc6a8 and agat mutants.

We will keep trying, but accept the reality that Rome was not built in a single day and that no transmitter was proven by one single paper.

A key new puzzle piece of evidence is the identification of creatine in synaptic vesicles. The experiment relies heavily on the purity of the SV fraction using the anti-synaptophysin antibody. I am quite sure that these preparations contain many other compartments - and of course a big mix of synaptic (and other) vesicles. Would it be possible to purify with an anti slc6a8 antibody?

Sl6a8 is expressed in on the plasma membrane of neurons7-9, instead of synaptic vesicles. Consistent with this, we could not detect obvious Slc6a8-HA signal in our starting material (Lane S in Author response image 6) that was used for SV purification. We have tried to purify SVs by HA antibody in Slc6a8 mice and SV markers could not be detected.

Lack of Slc6a8-HA in our starting material.

In Slc6a8-HA knock-in mice, the HA signal was present in whole brain homogenate (H), but not obvious in supernatants (S) following 35000 × centrifugation. In contrast, SV marker Syp was present in supernatants.

The K stimulation protocol in slices is relatively crude, as all neurons in the slice get simultaneously overactivated - and some of the effects on Ca-dependent release are not very strong (e.g. the 35 neurons that were not responsive to creatine at all). A primary neuronal culture of neurons that respond to creatine would strengthen this section.

To avoid the disadvantage of K stimulation, we also performed optogenetic experiments recently and obtained encouraging preliminary results.

Reviewer #2 (Recommendations For The Authors):

1. The different sections of the manuscript are not separated by headers.

1. The beginning of the results section either does not reference the underlying literature or refers to unpublished data.

We have kept a bit background in the beginning of the Results section.

1. The text contains many opinions and historical information that are not required (e.g., "It has never been easy to discover a new neurotransmitter, especially one in the central nervous system (CNS). We have been searching for new neurotransmitters for 12 years."; l. 17).

This is a field that has been dormant for decades and such background introductions are helpful for at least some readers.

1. Almeida et al. (2008; doi: 10.1002/syn.20280) provided evidence for electrical activity-, and Ca2+-dependent Cr release from rat brain slices. This paper should be introduced in the introduction.

Those were stand-alone papers which have not been reproduced or paid attention to. Our introduction part did not mention them because our research did not begin with those papers. We had no idea that those papers existed when we began. We started with SV purification and only read those papers afterwards. Thus, they were not necessary background to our paper but can be discussed after we discovered Cr in SVs.

1. Fig. 7: A Y-scale for the stimulation protocol is missing.

Revised.

Reviewer #3 (Recommendations For The Authors):

The main suggestion by this reviewer (beyond the details in the public review) is to consider the full spectrum of biology that is consistent with these results. By my reading, creatine could be a neurotransmitter, but other possibilities also exist, and the authors need to highlight those too.

We have discussed non-transmitter role in the discussion.

References

Ghirardini, E., G. Sagona, A. Marquez-Galera, F. Calugi, C. M. Navarron, F. Cacciante, S. Chen, F. Di Vetta, L. Dada, R. Mazziotti, L. Lupori, E. Putignano, P. Baldi, J. P. Lopez-Atalaya, T. Pizzorusso, and L. Baroncelli. 2023. Cell-specific vulnerability to metabolic failure: the crucial role of parvalbumin expressing neurons in creatine transporter deficiency. Acta Neuropathol Commun, 11: 34. doi: 10.1186/s40478-023-01533-w.

Lowe, M. T., Faull, R. L., Christie, D. L. & Waldvogel, H. J. Distribution of the creatine transporter throughout the human brain reveals a spectrum of creatine transporter immunoreactivity. J Comp Neurol 523, 699-725 (2015). https://doi.org:10.1002/cne.23667

Mak, C. S. et al. Immunohistochemical localisation of the creatine transporter in the rat brain. Neuroscience 163, 571-585 (2009). https://doi.org:10.1016/j.neuroscience.2009.06.065.

Molchanova, S. M., Oja, S. S. & Saransaari, P. Mechanisms of enhanced taurine release under Ca2+ depletion. Neurochem Int 47, 343-349 (2005). https://doi.org:10.1016/j.neuint.2005.04.027

Philibert, R. A., Rogers, K. L. & Dutton, G. R. K+-evoked taurine efflux from cerebellar astrocytes: on the roles of Ca2+ and Na+. Neurochem Res 14, 43-48 (1989). https://doi.org:10.1007/BF00969756

Rosko, L. M. et al. Cerebral Creatine Deficiency Affects the Timing of Oligodendrocyte Myelination. J Neurosci 43, 1143-1153 (2023). https://doi.org:10.1523/JNEUROSCI.2120-21.2022

Saransaari, P. & Oja, S. S. Characteristics of taurine release in slices from adult and developing mouse brain stem. Amino Acids 31, 35-43 (2006). https://doi.org:10.1007/s00726-006-0290-5

Schmidt, A. et al. Severely altered guanidino compound levels, disturbed body weight homeostasis and impaired fertility in a mouse model of guanidinoacetate N-methyltransferase (GAMT) deficiency. Hum Mol Genet 13, 905-921 (2004). https://doi.org:10.1093/hmg/ddh112

Speer, O. et al. Creatine transporters: a reappraisal. Mol Cell Biochem 256-257, 407-424 (2004). https://doi.org:10.1023/b:mcbi.0000009886.98508.e7

Takuma, K. et al. Ca2+ depletion facilitates taurine release in cultured rat astrocytes. Jpn J Pharmacol 72, 75-78 (1996). https://doi.org:10.1254/jjp.72.75

Version 2

Abstract

Abstract

The discovery of a new neurotransmitter, especially one in the central nervous system (CNS), is both important and difficult. We have been searching for new neurotransmitters for 12 years. We detected creatine (Cr) in synaptic vesicles (SVs), at a level lower than glutamate (Glu) and gamma-aminobutyric acid (GABA) but higher than acetylcholine (ACh) and 5-hydroxytryptamine (5-HT). SV Cr was reduced in mice lacking either arginine:glycine amidinotransferase (AGAT, a Cr synthetase) or SLC6A8, a Cr transporter with mutations among the most common causes of intellectual disability (ID) in men. Calcium-dependent release of Cr was detected after stimulation in brain slices. Cr release was reduced in SLC6A8 and AGAT mutants. Cr inhibited neocortical pyramidal neurons. SLC6A8 was necessary for Cr uptake into synaptosomes. Cr was found by us to be taken up into SVs in an ATP dependent manner. Our biochemical, chemical, genetic and electrophysiological results are consistent with the possibility of Cr as a neurotransmitter.Our novel approach to discover neurotransmitters is to begin with analysis of contents in SVs before defining their function and physiology.

Introduction

Neural signaling depends on chemical transmission between neurons and their target cells 1–6. Neurotransmission depends on chemicals such as neurotransmitters, neuromodulators and neuropeptides. Decades of work, sometimes with convoluted paths, were involved before a molecule was established as a classic neurotransmitter 7. Initial hints about cholinergic signaling were obtained in the 1800s 8–11. Choline 12,13 and acetylcholine (ACh) 14 were discovered decades before their pharmacological effects were found around the turn of 20th century 15–17. Henry Dale and colleagues realized similarities of ACh and parasympathetic stimulation 5,18 but it was not until 1929 when ACh was detected in the body 19 and 1934 when ACh was proven a neurotransmitter in the peripheral nervous system (PNS) 20–22. It took nearly 100 years from the finding of the effects of supradrenal gland damage 23 or removal 24, the observation of an activity in the supradrenal gland 25, the isolation of an inactive derivative 26–29, and the successful isolation of adrenaline 30–32, the notice of similarities between adrenaline and sympathetic stimulation 3,4,33–35, to the mid-1940s when Ulf von Euler proved that noradrenaline (NA) was the neurotransmitter of the sympathetic nerves 36–38. While it is not easy to establish a molecule as a neurotransmitter in the PNS, it is even harder to establish a central nervous system (CNS) neurotransmitter. Three decades elapsed between the time when ACh was proven to be a PNS neurotransmitter and the time when it was established as a CNS neurotransmitter 39–41 and two decades between NA as a peripheral transmitter and a central transmitter 42,43.

If a neurotransmitter acts only in the CNS, but not in the PNS, it is much more difficult to discover or to prove. Most neurotransmitters were discovered for their effects on peripheral tissues, with muscle contraction or relaxation as a major readout. Glutamate (Glu) 44–47 and gamma-aminobutyric acid (GABA) 46,48–51 were discovered partly because of their peripheral effects and partly because of their effects on spinal neurons. There is no reason for central neurotransmitters to also act peripherally, but relatively little efforts have been reported to find small molecule neurotransmitters acting only on CNS neurons with no peripheral bioassays available. Premature assumptions and technical difficulties are among the major reasons why the hunt for neurotransmitters has not been a highly active area of research over the last three decades.

Are there more neurotransmitters and how can they be discovered? Classic neurotransmitters are stored in synaptic vesicles (SVs) 52–60. They are released upon electric stimulation before being degraded enzymatically or taken up into the presynaptic terminal by cytoplasmic transporters and into SVs by vesicular transporters 61–66. Most of the major textbooks list either 3 67–69 or 4 70–72 criteria of a neurotransmitter: presence in presynaptic neurons, release upon stimulation, action on postsynaptic neurons, mechanism of removal. Some molecules commonly accepted as neurotransmitters still do not meet all the criteria listed in different textbooks, but they nonetheless play important functional roles in the CNS and their defects cause human diseases. Over time, different small molecules have been proposed to function as neurotransmitters (e.g., 46,73), but none satisfies all the criteria. Robust and reliable detection of the candidate molecule in SVs is often, though not always, the problem (cf. 74).

Beginning in 2011, we have been actively searching for new neurotransmitters in the mammalian brain. We have tried different approaches, including searching for neuroactive substances in the cerebral spinal fluid (CSF) and following transporters potentially localized in the SVs. One approach that we have now taken to fruition is the purification of the SVs from mouse brains coupled with chemical analysis of their contents. We have found known transmitters such as Glu, GABA, ACh, and 5-hydroxytryptamine (5-HT). But more importantly, we have reproducibly detected creatine (Cr) in SVs.

Cr was discovered in 1832 by Michel-Eugène Chevreul 75,76 and has long been considered as an energy buffer in the muscle and the brain 77–79. Half of Cr in a mammalian animal is thought to come from diet and the rest from endogenous synthesis 80. Most of the Cr is present in the muscle but it is also present in the brain. Although most of the endogenous Cr is synthesized in the kidney, the pancreas and the liver 77,81, Cr is also synthesized in the brain 80,82,83.

Solute carriers (SLC) contribute to both cytoplasmic and vesicular transporters. With 19 members in humans, family 6 (SLC6) are secondary active transporters relying on electrochemical Na+ or H+ gradients 84–89. SLC6 is also known as the neurotransmitter transporter (NTT) family because some members transport neurotransmitters such as GABA (by SLC6A1 or GABA transporter 1, GAT1; SLC6A13 or GAT2; SLC6A11 or GAT3) 90–93, NA (by SLC6A2 or NA transporter,

NET) 65, dopamine (by SLC6A3 or DAT) 94–96, 5-HT (by SLC6A4 or serotonin transporter, SERT) 66,97, and glycine (by SLC6A9, or GlyT1; SLC6A5 or GlyT2) 90,98,99. Cr is transported by SLC6A8 (also known as CrT, CT1 or CRCT) 100–106. In addition to peripheral organs and tissues, SLC6A8 is also expressed in the nervous system where it is mainly in neurons 82,104,107–109. SLC6A8 protein could be found on the plasma membrane of neurons 109–111.

The functional significance of SLC6A8 in the brain is supported by symptoms of humans defective in SLC6A8. Mutations in SLC6A8 were found in human patients with intellectual disability (ID), delayed language development, epileptic seizures and autistic-like behaviors 112,113. They are collectedly known as Cr transporter deficiency (CTD), with ID as the hallmark. Particular vulnerability of language development has been observed in some SLC6A8 mutations which had mild ID but severe language delay 114. CTD contributes to approximately 1 to 2.1% of X-linked mental retardation 115–123. While CTD is highly prevalent in ID males, it is also present in females, with an estimated carrier frequency of 0.024% 124.

SLC6A8 knockout mice 125–127 showed typical symptoms of human CTD patients with early and progressive impairment in learning and memory. Mice with brain- and neuronal-specific knockout of SLC6A8 showed deficits in learning and memory without changes in locomotion caused by peripheral involvement of SLC6A8 128,129. Deletion of SLC6A8 from dopaminergic neurons in the brain caused hyperactivity 130. These results demonstrate that SLC6A8 is functionally important in neurons.

Cr deficiency syndromes (CDS) are inborn errors of Cr metabolism, which can result from defects in one of the three genes: guanidinoacetate methyltransferase (GAMT) 131, arginine-glycine amidinotransferase (AGAT) 132,133, and SLC6A8 112. That they all show brain disorders indicates the functional importance of Cr in the brain134–137.

Here we first biochemically purified SVs from the mouse brain and discovered the presence of Cr, as well as classic neurotransmitters Glu and GABA, ACh and 5-HT, in SVs. We then detected calcium (Ca2+) dependent releases of Cr, Glu and GABA but not ACh and 5-HT when neurons were depolarized by increased extracellular concentrations of potassium (K+). Both the level of Cr in SVs and that of Cr released upon stimulation were decreased significantly when either the gene for SLC6A8 or the gene for AGAT were eliminated genetically. When Cr was applied to slices from the neocortex, the activities of pyramidal neurons were inhibited. Furthermore, we confirmed that Cr was taken up by synaptosomes and found that Cr uptake was significantly reduced when the SLC6A8 gene was deleted. Finally, we found that Cr was transported into SVs. Thus, multidisciplinary studies with biochemistry, genetics and electrophysiology have suggested that Cr is a new neurotransmitter.

Results

Detection of Cr in SVs from the mouse brain

To search for new neurotransmitters, we tried several approaches. For example, we used Ca2+ imaging to detect neuroactive substances in the cerebrospinal fluid (CSF), but it was difficult to rule out existing neurotransmitters and select responses from potentially new neurotransmitters. We also transfected cDNAs for all human SLCs into dissociated cultures of primary neurons from the mouse brain and found that more than 50 out of all SLCs could be localized in SVs. However, when we used CRISPR-CAS9 to tag some of the candidate SLCs in mice, some of them were found to be expressed outside the CNS, indicating that, while ectopic expression of these candidate SLCs could be localized on SVs, the endogenous counterparts were not localized on SVs.

Here we report our approach using the purification of SVs as the first step (Supplementary Figure 1A) 138–145. Synaptophysin (Syp) is a specific marker for SVs 146–149 and an anti-Syp antibody was used to immunoisolate SVs 143,144,150–152. Visualization by electronic microscopy (EM) (Supplementary Figure 1B, left panel) showed that the purified vesicles were homogenous, with an average diameter of 40.44 ± 0.26 nm (n=596 particles) (Supplementary Figure 1B, right panel), consistent with previous descriptions of SVs 153,154.

Immunoblot analysis with twenty markers of subcellular organelles of neurons and one marker for glia (Supplementary Figure 1C) indicate that our purifications were highly effective, with SV markers detected after purification with the anti-Syp antibody, but not that with the control immunoglobulin G (IgG). SV proteins included Syp 146–149, synaptotagmin (Syt1) 155, synatobrevin2 (Syb2) 156,157, SV2A 158, H+-ATPase 159, and vesicular neurotransmitter transporters for glutamate (VGLUT1, VGLUT2) 160 and GABA (VGAT) 161. Immunoisolation by the anti-Syp antibody did not bring down markers for the synaptic membrane (with SNAP23 as a marker) 162–164, postsynaptic components (with PSD95 and GluN1 as markers) 165,166, the Golgi apparatus (with GM130 and Golgin 97 as markers) 167,168, early endosome (with early endosome-associated 1, EEA1, as a marker) 169, the lysosome (with LC3B and cathepsinB as markers), the cytoplasma (with glyceraldehyde-3-phophate dehydrogenase or GAPDH as a marker), mitochondria (with voltage-dependent anion channel or VDAC, as a marker), cytoplasmic membrane (with calcium voltage-gated channel subunit alpha 1 or CACNA1A as a marker), axonal membrane (with glucose transporter type 4 or GluT4 as a marker) and glia membrane (with myelin basic protein or MBP as a marker). These results indicated that the SVs we obtained were of high integrity and purity.

To detect and quantify small molecules as candidate transmitters present in the purified SVs, capillary electrophoresis-mass spectrometry (CE-MS) was optimized and utilized (Figure 1A, Supplementary Figure 1A) 151,170. We found that the levels of classical neurotransmitters such as Glu, GABA, ACh and 5-HT were significantly higher in SVs pulled down by the anti-Syp antibody than those in lysates pulled down by the control IgG (Figure 1 A-E). Consistent with previous reports 144,151, significant enrichment of neurotransmitters was observed only from SVs immunoisolated at near 0 °C, but not at the room temperature (RT) (Figure1 A-E). By contrast, another small molecule, alanine (Figure 1G), was not elevated in SVs as compared to the control.

Presence of Cr in SVs from the mouse brain.

(A) Representative raw traces from CE-MS of indicated molecules from samples immunoisolated by the control IgG (blue) at 0-2 °C, the monoclonal anti-Syp antibody at 0-2 °C (red) and the anti-Syp antibody at room temperature (RT, green). Q1/Q3 for identifying targets were indicated. (B-G) Quantification of the amounts of indicated molecules. The amount of a molecule was divided by the amount of the anti-Syp antibody bound to magnetic beads. Note Glu (B), GABA(C), ACh (D), 5-HT (E), Cr (F), but not alanine (G) was higher in SVs pulled down by the anti-Syp antibody at 0-2 °C than those pulled down by the IgG control or those pulled down at the RT. n=10 (B-E, and G) or 14 (F) samples per group, ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

The amount of Glu was 171.1 ± 5.4 pmol/μg anti-Syp antibody (n=10, Figure 1B), approximately 10 times that of GABA (n=10,17.81 ± 1.47 pmol/μg anti-Syp antibody, Figure 1C). The amount of ACh was 1.29 ± 0.10 pmol/μg anti-Syp antibody (n=10, Figure 1D), approximately 0.072 that of GABA. The amount of 5-HT was 0.096 ± 0.017 pmol/μg anti-Syp antibody (n=10, Figure 1E). Thus, our purification and detection methods were highly reliable and sensitive enough to detect established neurotransmitters.

Under the same conditions, we also detected Cr in SVs (n=14, Figure 1A and 1F). Amount of Cr in the SVs was found to be 3.43 ± 0.40 pmol/μg anti-Syp antibody (Figure 1F), which was approximately 2% of Glu, 19% of GABA, 266% of ACh and 3573% of 5-HT. It is unlikely that these could be attributable to different levels for different neurotransmitters in each SV, but more likely attributable to the relative abundance of SVs containing different neurotransmitters. 85% to 90% neurons in the mouse brain were glutamatergic while 10%-15% were GABAergic 171–173, which can explain our detection of Glu as approximately 10 times that of GABA (Figure 1 B and C). Similarly, cholinergic neurons (5.67−105) 174 represented 0.81% of total number of neurons (approximately 70 million) in the mouse brain 175, serotonergic neurons (approximately 26,000) for 0.037% of total neurons 175,176. Assuming that the content of each neurotransmitter in a single SV is similar, extrapolation from the above data would suggest that approximately 1.3% to 2.15% of neurons in the mouse brain are creatinergic.

To distinguish whether small molecules co-purified with SVs were in the SVs 143,144, or that they were just associated with the outside of SVs154, we tested the dependence of the presence of these molecules in the SVs on temperature and on the electrochemical gradient of H+. Cr was significantly reduced in SVs purified at the room temperature as compared to that immunoisolated at near 0 °C (Figure 1F), supporting the presence of Cr inside, instead of outside, SVs.

Classical neurotransmitters are stored in SVs with an acidic environment inside (pH of 5.6 ∼ 6.4) 177–179. To further verify the storage of Cr in SVs and examine the role of H+ electrochemical gradient, we applied pharmacological inhibitors during purification 74,180. The proton ionophore FCCP (carbonyl cyanide-4-(tri-fluoromethoxy) phenylhydrazone) was used to dissipate H+ electrochemical gradient 180, 181. FCCP significantly reduced the amount of Cr as well as classical neurotransmitters in SVs (Supplementary Figure 2A-E). The extent of FCCP induced reduction was correlated with the value of pKa or PI (isoelectric point) for different molecule: 5-HT (with pKa predicted to 10 and 9.31, Supplementary Figure 2E) > Cr (PI of ∼7.94, Supplementary Figure 2A) > GABA (PI of 7.33, Supplementary Figure 2C) > Glu (PI of 3.22, Supplementary Figure 2B). Nigericin, a K+/H+ exchanger which dissipates ΔpH180,181, also reduced the amount of Cr and classical neurotransmitters in SVs (Supplementary Figure 2A-E). Furthermore, in the presence of FCCP or nigerin, SV Cr was reduced to a level comparable to that pulled down by control IgG (Supplementary Figure 2A), demonstrating the storage of Cr in SVs was dependent on H+ gradient. As a control, the non-neurotransmitter molecule alanine in SVs was not changed by the inhibitors (Supplementary Figure 2F).

Reduction of SV Cr in mouse mutants lacking slc6a8

SLC6A8, located on the X chromosome, encodes a transporter for Cr and its loss of function (LOF) mutations caused behavioral deficits in humans 112,113 and mice 125–130. To investigate whether SLC6A8 affects Cr in SVs, we generated slc6a8 knockout (KO) mice. Exon 1 of the SLC6A8 gene was partially replaced with CreERT2-WPRE-polyA by CRISPR/Cas9 (Figure 2A). Examination by reverse polymerase chain reaction (RT-PCR) (Supplementary Figure 3A and 3B) and quantitative real-time reverse PCR (qPCR, Supplementary Figure 3C-D) showed that SLC6A8 mRNA was not detected in either male or female mutants, and significantly reduced in female heterozygous (Slc6a8+/-). Consistent with previous reports, the body weights of SLC6A8 KO mice were reduced (Supplementary Figure 4 B and D) 125,182,183. Brain weight was not significantly different between SLC6A8 KO mice and WT mice (Supplementary Figure 4A and C).

SLC6A8 and Cr in SVs.

(A) A schematic illustration of the strategy for generating SLC6A8 knockouts using CRISPR/Cas9. An iCreERT2-WPRE-pA cassette (∼3.5 kb) was inserted immediately downstream of ATG in the SLC6A8 gene, substituting bp 4 to bp 51 in exon 1 (E1). (B) Representative raw traces of Creatine immunoisolated by control IgG from WT mice (blue), the anti-Syp antibody from WT mice (red), IgG from SLC6A8 KO mice (blue) and the anti-Syp antibody from SLC6A8 KO mice (red). (C-H) Quantification of indicated molecules. Note the selective reduction of Cr in SVs from SLC6A8 KO mice. n=14 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

When we examined the contents of SVs isolated by the anti-Syp antibody vs the control IgG, significant reduction was only observed for Cr, but not classical neurotransmitters (Figure 2 B-H, Supplementary Figure 5 A-E). While Cr pulled down by IgG was not significantly different between SLC6A8-/Y and SLC6A8+/Y mice, SV Cr purified by the anti-Syp antibody from SLC6A8-/Y was reduced to approximately 1/3 that of the WT (SLC6A8+/Y) littermates (n =14, Figure 2 B-C). Compared to the IgG control, Cr in SVs was enriched in WT mice, but not in SLC6A8-/Y mice (Figure 2 B and C). In both SLC6A8-/Y and SLC6A8+/Y mice, classical neurotransmitters in SVs were all enriched as compared to IgG controls (Figure 2 C-G, Supplementary Figure 5 A-D). The amounts of Glu (Figure 2D and Supplementary Figure 5A), GABA (Figure 2E and Supplementary Figure 5B), ACh (Figure 2F and Supplementary Figure 5C) and 5-HT (Figure 2G and Supplementary Figure 5D) in SVs were not different between SLC6A8-/Y and SLC6A8+/Y mice. Molecules not enriched in SVs from WT mice, such as alanine, was also unaffected by SLC6A8 KO (Figure 2H and Supplementary Figure 5E).

It is unlikely that the specific reduction of Cr in SVs from SLC6A8 KO mice was due to technical artifacts. First, the possibility of less SVs obtained from SLC6A8 KO mice were precluded by immunoblot analysis, as assessed by SV markers Syp, Syt and H+-ATPase (Supplementary Figure 6A). Second, data collected by high resolution MS (Q Exactive HF-X, Thermoscientific) also revealed selective decrease of SV Cr (m/z=132.0576) from SLC6A8 KO mice (n=8, Supplementary Figure 6 B-G), as quantified by the peak area. Peak areas for Glu (n=8, Supplementary Figure 6B, m/z=148.0604), GABA (n=8, Supplementary Figure 6C, m/z=104.0712), ACh (n=8, Supplementary Figure 6D, m/z = 146.1178) and alanine (n=8, Ala, Supplementary Figure 6E, m/z=90.055) were not significantly different between SVs immmunoisolated with the anti-Syp antibody and control IgG from WT and Slc6a8 KO mice. However, peak areas (Supplementary Figure 5G) and amplitude of Cr (n=8, Supplementary Figure 5F) signal were significantly increased in SVs from WT mice (anti-Syp antibody vs IgG), but not that from SLC6A8 KO mice.

Reduction of SV Cr in mouse mutants lacking AGAT

AGAT is the enzyme catalyzing the first step in Cr synthesis 82,184 and its absence also led to Cr deficiency in the human brain and mental retardation 132,133. To investigate requirement of AGAT for SV Cr, we utilized AGAT ‘knockout-first’ mice (Figure 6A) 185. The targeting cassette containing Frt (Flip recombination sites)-flanked EnS2A, an IRES::lacZ trapping cassette and a floxed neo cassette were inserted downstream of exon 2 to interfere with normal splicing of AGAT pre-mRNA. Examination by RT-PCR (Supplementary Figure 7A) and quantitative RT-PCR (Supplementary Figure 7B) showed reduction of AGAT mRNA in AGAT+/- mice and absence of AGAT mRNA in AGAT-/- mice. Body weight (Supplementary Figure 8B), but not brain weight (Supplementary Figure 8A), of AGAT-/- mice were lower than both AGAT+/+ and AGAT+/- mice, which were similar to SLC6A8 KO mice.

Immunoblot analysis showed SVs purified from the brains were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice (Supplementary Figure 9 A-D), as supported by quantitative analysis of Syp, Syt and H+-ATPase (n=20, with 2 repeats for 10 samples).

We analyzed small molecules present in SVs from AGAT+/+, AGAT+/- and AGAT-/- mice. Cr was significantly enriched in SVs from all three genotypes as compared to the IgG control (Figure 3A and 3B). However, the level of Cr from AGAT-/- mice was significantly lower than those from AGAT+/+ and AGAT+/- mice (n=10, Figure 3A and B).

Reduction of SV Cr in AGAT knockout mice.

(A) Representative raw traces of Creatine, pulled down by IgG or anti-Syp from AGAT+/+, AGAT+/- and AGAT-/- mice. (B-G) Quantification of indicated molecules. Cr was significantly decreased in AGAT-/- mice as compared to Cr in AGAT+/+ or AGAT+/- mice (B). GABA was significantly decreased in SVs from AGAT-/- mice as compared to AGAT+/+ mice (D) but the difference was smaller than that of Cr. Glu (C), ACh (E), 5-HT (F) and alanine was not different among AGAT+/+, AGAT+/- and AGAT-/- mice. n=10 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

Glu (Figure 3C), ACh (Figure 3E) and 5-HT (Figure 3F) were all enriched in SVs (as compared to IgG controls), and not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice. GABA in SVs from AGAT-/- mice was also decreased from AGAT+/+ mice by 30%, to an extent less than that of Cr (78.4%). Alanine was not different among three genotypes of mice (n=6, Figure 3G). Thus, Cr and GABA, but not other neurotransmitters, in SVs were reduced in AGAT KO mice.

Pattern of SLC6A8 expression indicated by knockin mice

We generated SLC6A8-HA knockin mice by CRISPR/Cas9. Three repeats of the hemagglutinin (HA) tag 186, the T2A sequence 187–189 and CreERT2 190–193 were inserted in-frame at the C terminus of the SLC6A8 protein (Figure 4A).

Expression pattern of SLC6a8.

(A) A diagram illustrating the generation of SLC6A8-HA KI mice. 3*HA, T2A and CreERT2 were inserted in-frame, before the stop codon, to the C terminus of SLC6A8 protein. (B-K) Low magnification photomicrographs of coronal sections immunohistochemically labelled with the anti-HA antibody in SLC6A8-HA mice. (I and M) immunostaining with the anti-HA antibody in control WT mice. Abbreviations: Pir, piriform cortex; M, motor cortex; LS: lateral septum; Hp: hippocampus; Hb, habenular nucleus; Vp: ventral posterior nucleus of thalamus; Au auditory cortex; Amyg: amygdala; Ip, interpeduncular nucleus; Pn; pontine nucleus; Cb, cerebellum; Pr, prepositus; SC, Superior colliculus; MVe, medial vestibular nucleus. Scale bar: 500 μm.

To examine the expression pattern of SLC6A8, we performed immunocytochemistry with an antibody against the HA epitope in SLC6A8-HA and WT mice. SLC6A8-HA mice showed positive signals in the olfactory bulb (Figure 4B), the piriform cortex (Figure 4 C-F), the somatosensory cortex (Figure 4 F and G), the ventral posterior thalamus (Figure 4H), the interpeduncular nucleus (Figure 4I) and the pontine nuclei (Figure 4J). In addition, moderate levels of immunoreactivity were observed in the motor cortex (Figure 4 D-H), the medial habenular nucleus (Figure 4H), the hippocampus (Figure 4H) and the cerebellum (Figure 4K). These results were consistent with previous reports 109,111. WT mice were negative for anti-HA antibody staining (Figure 4 L and M).

Ca2+ dependent release of Cr upon stimulation

Classical neurotransmitters are released from the SVs into the synaptic cleft in a Ca2+ dependent manner after stimulation. For example, high extracellular potassium (K+) stimulated Ca2+ dependent release of Glu, GABA and other neurotransmitters in brain slices 194–198.

300 μm thick coronal slices of the mouse brain within 1∼2 millimeters (mm) posterior to the bregma were used, because the cortex, the thalamus, the habenular nucleus and the hippocampus were positive for SLC6A8 (cf., Figure 4H). We monitored the effect of K+ stimulation by recording neurons in the slices. Immediately after K+ stimulation, pyramidal neurons in the CA1 region of the hippocampus were depolarized, firing a train of action potentials and reaching a large depolarization plateau in less than 1 minute (min) (Figure 5A). K+-induced depolarization persisted for several mins before returning to the baseline and being washed within 10 mins. Thus, superfusates in 1-min fraction at the time points of 1.5 min before (control) and after K+ stimulation, and 10 mins after the wash were collected (Figure 5A), and the metabolites in the superfusates were analyzed by CE-MS.

Cr release in brain slices from WT and SLC6a8 knockout mice.

(A) Neuronal depolarization induced by 70 mM K+ and time points for collecting the release sample. “Control” samples were collected 1.5 min to 0.5 min before K+ stimulation, “70 mM K+” ACSF samples were collected during 70 mM K+ stimulation, and “wash” samples were collected 10 mins after washout with ACSF. Efflux of Glu (B), GABA (C) or Cr (D) from WT or SLC6A8 KO male mice (n=7 samples per group). Note that a small amount of Cr released in SLC6A8 KO mice did not return to the baseline after 10 mins of washing. (E-G) Ca2+ dependent release of Glu, GABA and Cr in WT and SLC6A8 KO mice (n=5 samples per group). ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction. #, p<0.05, ##, p<0.01, paired t test.

Cr release in WT and AGAT knockout mice.

(A) A schematic diagram illustrating the strategy of AGAT knockout first. With the AGAT gene (also known as GATM) shown in the upper part, and the gene targeting strategy in the lower part. The homologous arm is approximately 10 kb. A targeting cassette, containing Frt-flanked lacZ and neomycin, was inserted downstream of exon 2. At the same time, exon 3 of AGAT was flanked by two loxP sites. K+ induced release of Glu (C) and GABA (D) were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice, whereas that of Cr (B) was significantly lower in AGAT-/- mice than those in AGAT+/+ and AGAT+/- mice.

In the presence of Ca2+, depolarization with elevated extracellular K+ led to robust release of Glu and GABA in slices from WT (SLC6A8+/Y) mice (n=7 per group, Figure 5 B and C). After 10 mins wash, levels of Glu and GABA returned to the baseline (Figure 5 B and C). In the presence of Ca2+, depolarization with elevated K+ led to robust release of Cr. Extracellular Cr returned to the baseline after 10 min wash (Figure 5D). For quantification, the stimulated releases of metabolites were calculated by subtracting the basal levels from the total releases in response to K+ stimulation. In the presence of Ca2+, K+ stimulation induced the efflux of Glu, GABA and Cr at 0.46, 0.33 and 0.086 nmol/min, respectively (n=7 per group) (Figure 5 B, C, D). From the detection limits of ACh and 5-HT in our system, we inferred that the efflux rate for ACh was lower than 0.001 nmol/min and that for 5-HT lower than 0.003 nmol/min. The efflux rate for Cr in brain slices is lower than those of Glu and GABA, but higher than those for ACh and 5-HT.

Ca2+ dependence of transmitter release was examined by comparing responses to ACSF without Ca2+ or elevated K+ (supplemented with 1 mM EGTA), elevated extracellular K+ in the absence of Ca2+ (supplemented with 1 mM EGTA), or K+ in presence of 2.5 mM Ca2+ (Figure 5 E-G, n=5 per group). In the absence of Ca2+, elevated K+ stimulated the release of a small but significant amount of Glu and GABA, with efflux rates at 0.056 nmol/min and 0.066 respectively (Figure 5 E, F). In the presence of 2.5 mM Ca2+, elevated K+ further augmented the release of Glu and GABA by 5 to 6 times, confirming previously reported Ca2+ dependent release of neurotransmitters in response to depolarization 194,196.

Cr was also released both in a Ca2+ dependent and a Ca2+ independent manner (Figure 5G). More Cr was released in response to K+ stimulation in the presence of 2.5 mM Ca2+ than that in the absence of Ca2+. These results demonstrate Ca2+ dependent release of Cr upon stimulation.

Reduced Cr release in SLC6A8 and AGAT mutant mice

We examined whether SLC6A8 KO affected K+-induced release of Cr. While Glu and GABA were released in slices from SLC6A8 KO (SLC6A8-/Y) mice at levels not significantly different from those of WT mice (Figure 5 B, C), release of Cr in response to K+ stimulation was significantly reduced in SLC6A8-/Y mice as compared to SLC6A8+/Y mice (Figure 5D). The basal level of Cr in SLC6A8 KO mice was lower than that of WT mice. In addition, K+ stimulation induced release of Cr persisted to some extent even after 10 mins of washout (Figure 5D), possibly due to the inability of presynaptic terminals in SLC6A8 KO mice to reuptake Cr in the synaptic cleft (Figure 8).

Experiments with slices from brains of SLC6A8 KO (SLC6A8-/Y) mice showed that Ca2+ dependent release of either Glu or GABA was not affected by the genotype of SLC6A8 (Figure 5 E and F). By contrast, Ca2+ dependent release of Cr was abolished in SLC6A8-/Y slices. Interestingly, Ca2+ independent release of Cr was reduced by a third, but did not reach statistical significance, in SLC6A8-/Y slices. In the absence of Ca2+, the basal level of Cr was not changed in SLC6A8 KO mice. Taken together, these results indicate that there is Ca2+ dependent release of Cr upon stimulation and that SLC6A8 is required specifically for Ca2+ dependent release of Cr, but not for Ca2+ dependent release of other neurotransmitters such as Glu and GABA, or for Ca2+ independent release of Cr. Knockout of AGAT (Figure 6A) selectively reduced K+ evoked release of Cr, but not those of Glu or GABA (n=5 per group, Figure 6 B, C, D). Although K+ stimulation still elicited Cr release from brain slices of AGAT+/- mice, the efflux rate in AGAT-/- mice was reduced to less than 10% that in AGAT+/+ mice and 20% that in AGAT+/- mice.

Cr inhibition of neocortical neurons

Our own data (Figure 4H) and previous reports 109,111 have shown SLC6A8 in the neocortex, with dense SLA6A8-HA immunoreactive fibers in layer 4 (Supplementary Figure 10A) 199,200. Layer 5 neurons in the somatosensory cortex have been reported to express SLC6A8 previously 109,111. To investigate electrophysiological effects of Cr, we performed whole cell patch clamp recordings from the pyramidal neurons in layer 4/5 of the somatosensory cortex (Figure 7 A and B, Supplementary Figure 10).

Inhibitory effects of Cr on cortical neurons.

(A) A photograph showing recording at layer 4 in the somatosensory cortex. Scale bar: 10 μm. (B) Patch-clamp recording of a pyramidal neuron. Scale bar: 10 μm. (C) Ratios of Cr-responsive and -unresponsive neurons in the region. (D) Representative raw electrophysiological traces showing inhibition of evoked firing by Cr, with the lower panel showing the stimulus protocol. (E) Evoked spike numbers in response to different current injections from (D). (F) Relationship between evoked spike numbers and different current injections to neurons that were inhibited by Cr (n=16). (G) The same for Cr-unresponsive neurons (n=35). (H) Rheobase for Cr responsive neurons. (I) Evoked spike number when these neurons were injected with current of rheobase + 50 pA. (J) input resistance. *, p<0.05; **, p<0.01; ***, p<0.001, paired t test.

Cr uptake into synaptosomes and SVs.

(A) Markers for subcellular organelles detected in synaptosomes prepared from WT mice or mice with Slc6a8 gene fused in frame to the HA epitope. SLC6A8-HA was enriched in synaptosomes (Sy or 4-Sy) or crude synaptosomes (P2) (see methods), as were other markers for the subcellular organelles in synaptosomes but not the marker for myelin (MBP). (B) Representative electron micrographs and histograms of size distribution in synaptosomes isolated from WT (Slc6a8+/Y) and Slc6a8 KO (Slc6a8-/Y) mice by Ficoll density-gradient centrifugation. Sy, synaptosome; SV, synaptic vesicles; Mt, mitochondria; PSD, post-synaptic density. Bar, 20 nm. (C) Cr uptake into synaptosomes (n=5 per group). The two left columns were results from WT mice and the two right columns from Slc6a8 knockout mice. The control baseline was [14C]-Cr uptake at 0 °C at 10 mins (205). Cr uptake into synaptosomes at 37 °C measured at 10 mins was observed in WT synaptosomes. Uptake into Slc6a8 knockout synaptosomes was significantly reduced when compared to the WT synaptosomes. ***, p<0.001, one-way ANOVA with Tukey’s correction. (D) uptake of [13C]-Cr into immunoisolated SVs in the presence or absence of ATP. (n=11 samples per group)*, p<0.05, paired-t test.

Medium-sized pyramidal neurons (Figure 7B) with a membrane capacitance (Cm) of 114.96 ± 3.92 pF (n=51, Supplementary Figure 11A) were recorded. These neurons exhibited regular firing patterns 201,202 in response to depolarization current injection (Figure 7D and Supplementary Figure 11A) with moderate maximal evoked spiking frequencies of 10-30 spikes per 500 milliseconds (ms) (Figure 7 E-G), increasing of inter-spike intervals during depolarizing steps (Figure 7D and Supplementary Figure 11A), high action potential amplitude (81.64 ± 1.06 mV, Supplementary Figure 12E) and large spike half-width (1.12 ± 0.031 ms, Supplementary Figure 12F).

Cr was bath applied only after the evoked firing pattern reached a steady state. Of the 51 neurons, 16 were inhibited by 100 μM creatine (Figure 7 B-F). Fewer spikes were evoked in Cr responsive neurons in response to depolarizing current injections during Cr application (25 pA step, 500 ms) (Figure 7 D-F). The inhibitory effect of Cr was reversible (Figure 7 D-F), typically observed within 2-3 min following Cr application (with maximal effect from 2 to 8 mins) and disappeared after 10 to 25 min washout. This could be repeated by a second application of Cr. The rheobase, defined as the minimal electrical current necessary to elicit an action potential, was increased during bath application of Cr (Figure 7 H). The inhibitory effect was most obvious at near spike threshold. When a neuron was depolarized with a current of 50 pA above rheobase, the number of evoked spikes was decreased dramatically during Cr application (Figure 7I). Cr also mildly inhibited the input resistance (Figure 7J), slightly hyperpolarized resting membrane potential (Supplementary Figure 12C) or reduced amplitude of afterhyperploarization (AHP) followed by the first evoked action potential (Supplementary Figure 12G). The spike threshold (Supplementary Figure 12D), amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F) were not changed by Cr.

The remaining 35 neurons were not responsive to Cr (Figure 7C and G, Supplementary Figure 11). Cr did not change electrical parameters tested, including evoked firing rates (Figure 7G, Supplementary Figure 11 A, B and D), rheobase (Supplementary Figure 11C), resting membrane potential (Supplementary Figure 12C), spike threshold amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F). In addition, electrical properties of responsive neurons and unresponsive neurons were not significantly different. With the limited number of neurons recorded, the ratio of responsive neurons appeared higher in layer 4 or border of layer4/5, than the deeper layer in layer 5 (Supplementary Figure 10).

SLC6A8 dependent uptake of Cr into the synaptosomes

Along with enzymatic degradation, re-uptake by transporters serves as an important way to remove neurotransmitters released into the synaptic cleft. As synaptosomes contain the apparatus for neurotransmission, they are often used for studying uptake of neurotransmitters 203.

To investigate whether Cr uptake into synaptosomes required SLC6A8, we first examined whether SLC6A8 was present in synaptosomes. Using Slc6a8-HA knockin mice and an anti-HA antibody, we found that SLC6A8-HA was present and enriched in crude synaptosomal fraction (P2 fraction in Figure 8A, enrichment score: P2/H = 1.76 ± 0.15, n = 4) and synaptosomal fraction prepared using a discontinuous Ficoll gradient (Sy and 4-Sy fractions in Figure 8A, enrichment score: Sy/H = 2.02 ±0.14, n = 4). The integrity of synaptosomes was confirmed by multiple markers of the synaptosomes 203, including the presynaptic membrane marker SNAP25 204 and the SV marker Syp (synaptophysin Syp) 146–149 in presynaptic terminals, the postsynaptic density marker PSD95 165,166 and the postsynaptic membrane protein GluN1 205, the synaptic membrane protein SNAP23 162–164, the plasma membrane marker Na+-K+-ATPase 206–208, and the mitochondria marker VDAC 203. These were all enriched in our synaptosomal preparations. The cytosol marker GAPDH was also present in synaptosomes, whereas the oligodendrocyte marker MBP was nearly absent, suggesting myelin pollution was largely avoided 203.

We have also used EM to confirm the quality of our synaptosome preparations. As reported previously 203,209, synaptosomes were composed of membrane bounded structures (Sy in Figure 8B) filled with synaptic vesicles (SV in Figure 8B), sometimes with a segment of postsynaptic membrane along with the postsynaptic density (PSD in Figure 8B) and mitochondria (Mt in Figure 8B). The sizes of synaptosomes from WT mice and Slc6a8 knockout mice were similar, with areas of 0.245 ± 0.01 μm2 (n=302 particles) and 0.247 ± 0.01 μm2 (n=317 particles), respectively (Figure 8B).

We then examined whether SLC6a8 participated in Cr uptake into the synaptosomes. A mixture of 18 μM [14C]-Cr (with a total radioactivity of 0.4 μCi) and 5 μM Cr was used, and uptake at 0 °C measured at 10 mins was the baseline 210. Cr uptake into synaptosomes from WT mice was stimulated approximately sevenfold at 37 °C (Uptake, Figure 8C) as compared with 0 °C (Ctrl, Figure 8C). Cr uptake into synaptosomes from SLC6a8 knockout mice was less than 3 times as compared to its control, and was decreased to approximately 1/3 of that of WT mice (Figure 8C). Thus, SLC6A8 is necessary for uptake of Cr into the synaptosomes.

Cr uptake into SVs

Classical neurotransmitters were taken up in SVs in an ATP-dependent manner144,159,181,210,211. We examined whether Cr could be transported into SVs.

We used 10 μg anti-Syp antibody to purify SVs from mouse brains. Purified SVs were preincubated for 30 min to allow sufficient leakage of endogenous Cr, before being mixed with 1 mM [13C]-Cr in the presence or absence of 4 mM ATP and placed at 25 L for 10 mins to allow adequate uptake. The SV content of [13C]-Cr was then examined by CE-MS and high-performance liquid chromatography-mass spectrometry (HPLC-MS). Significantly more [13C]-Cr were taken up by SVs in the presence of ATP, with about 10.3 pmol [13C]-Cr transported into SVs (1.03 pmol/μg α-Syp or transportation rate of 0.103 pmol/min, n=11, Figure 8D).

In summary, Cr could be transported into SVs in ATP-dependent manner. At this point, we do not know what is the transporter(s) on the SVs for Cr uptake. SLC6A8 is only found in plasma membrane, not on SVs, and is not a candidate for Cr uptake into SVs.

Discussion

While no neurotransmitter has been proven in a single paper, supportive evidence suggesting Cr as a possible new neurotransmitter has been presented here to the extent of any single previous papers.

At various times and by different researchers, taurine 46,212, proline 73, D-aspartic acid 213, hydrogen sulfide 214, agmatine 215, DOPA 216, estradiol 217, β-alanine 218, protons 219 have been suspected as neurotransmitters, but they do not meet all the criteria. Some of the suspected molecules can be released upon stimulation or removed by transporters. Often, they have not been reproducibly found in SVs 74.

Our discovery of Cr in SVs significantly raised the priority of testing the candidacy of Cr, and our further investigations have led to more evidence suggesting Cr as a neurotransmitter: 1) Cr is stored in SVs; 2) Ca++ dependent release of Cr upon stimulation has been observed; 3) both Cr storage in SVs and Cr release are reduced when either the gene for SLC6A8 or the gene for AGAT was deficient; and 4) Cr inhibits activities of pyramidal neurons in the neocortex; 5) Cr uptake into synaptosomes requires SLC6A8; 6) Cr uptake into SVs was ATP-dependent.

Of the above results, 1, 3, 4 and 6 are reported for the first time in this paper. Furthermore, we have demonstrated that detection of Cr in SVs was lower than those for Glu and GABA, but higher than those for ACh and 5-HT, placing Cr at a level in the middle of known central transmitters (Figures 1-3). The storage of Cr in SVs is dependent on preserved H+ gradient (Supplementary Figure 2) and Cr can be transported into SVs (Figure 8D).

There was a single previous report of Ca++ dependent release of [3H]Cr and endogenous Cr in response to electrical stimulation 220. We now provide evidence that Cr was released in response to extracellular K+ stimulation (within 1-2 min) (Figure 5-6). Furthermore, Cr release was reduced when either the Slc6A8 or AGAT gene was removed (Figures 5 and 6). Although the Ca2+ dependent component of K+ evoked Cr release was smaller as compared to those of Glu and GABA, it nevertheless existed and was totally abolished by Slc6A8 knockout (Figure 5). The reported electrically evoked Cr release showed more Ca2+ dependence 220. Taken together, our data and previous report 220 supported a role of Cr as a neurotransmitter. Our observation of extremely low efflux rates of 5-HT or ACh may have arisen from very limited numbers of cholinergic174 or serotoninergic neurons176 in the sliced sections and rapid enzymatic degradation of these neurotransmitters.

Cr uptake from the extracellular space into the cells was reported twice previously, once with brain slices showing sodium dependent uptake of [3H]Cr 220 and once with synaptosomes 221. Our new results have not only replicated the synaptosome Cr uptake experiment but also shown the requirement of SLC6A8, a membrane transporter expressed in synaptosomes (Figure 8A), for Cr uptake into synaptosomes. Transportation of Cr into synaptosomes by Slc6A8 may function for both the clearance of Cr from the synaptic cleft and recycling of Cr into SVs residing in neurons (Figure 8B, Figure 6D).

In summary, in addition to confirming and extending previous results which have stood alone for more than a decade without replication or follow-up, we have obtained entirely new results suggesting the candidacy of Cr as a neurotransmitter. We discuss below the criteria for a neurotransmitter, Cr as a neurotransmitter, and the implications of Cr as a neurotransmitter.

Criteria of a neurotransmitter

The criteria for establishing a non-peptide small molecule as a neurotransmitter have varied from time to time and from author to author.

Some textbooks simply state that a neurotransmitter is stored presynaptically, released upon stimulation and active on postsynaptic neurons. The details of these 3 criteria can vary. For example, one textbook stipulates that “the substance must be present within the presynaptic neuron; the substance must be released in response to presynaptic depolarization, and the release must be Ca2+ dependent; specific receptors for the substance be present on the postsynaptic cell” 67,68. Another states that “the molecule must be synthesized and stored in the presynaptic neuron; the molecule must be released by the presynaptic axon terminal upon stimulation; the molecule, when experimentally applied, must produce a response in the postsynaptic cell that mimics the response produced by the release of neurotransmitter from the presynaptic neuron69. The neuroscience textbook most widely used internationally for the last 4 decades lists 4 criteria for a neurotransmitter 70,71: it is synthesized in the presynaptic neuron; it is present within vesicles and is released in amounts sufficient to exert a defined action on the postsynaptic neuron or effector organ; when administered exogenously in reasonable concentrations it mimics the action of the endogenous transmitter; a specific mechanism usually exists for removing the substance from the synaptic cleft. These are similar, but not identical, to the classic textbook on neurotransmitters: a neurotransmitter “should be synthesized and released presynaptically; it must mimic the action of the endogenous compound that is release on nerve stimulation; and where possible, a pharmacological identity is required where drugs that either potentiate or block postsynaptic responses to the endogenously released agent also act identically to the suspected neurotransmitter that is administered” 222. The pharmacological criterion is listed in another textbook 223.

Some authors note difficulties of establishing a CNS neurotransmitter. For example, a specialized neurotransmitter book states that “the candidate neurotransmitter should be present in the presynaptic terminal, be released when the presynaptic terminal is active, and when applied experimentally, induce faithful responses in the postsynaptic neuron. In practice, since central nervous system neurons continuously integrate diverse excitations and inhibitions, the last criterion is relaxed to demonstrating merely changes in such activity” 224.

Solomon Snyder, a leading scientist of classic neurotransmitters, neuropeptides and their receptors, wrote that “designating a molecule as a transmitter depends on the criteria employed, the most common of which are that the substance is synthesized in neurons, released by their terminals, mimics the effects of physiologic neurotransmission and possess a mechanism for inactivation. However, with each new candidate the rules have been modified and broadened” 225.

Evidence supporting Cr as a neurotransmitter

16 small molecules have been listed as neurotransmitters in the classic textbook 70,71. Among them, adenosine, arachidonic acid, nitric oxide and carbon monoxide do not meet all four criteria, at the present. Cr appears to be better than these in meeting the criteria for a central neurotransmitter.

The results obtained by us in this paper have satisfied the criteria of Robinson 224 for Cr to be a CNS neurotransmitter.

The four criteria of Snyder and colleagues 225 have been mostly met but the physiological neurotransmission would require more research because a specific synapse(s) would have to be defined and studied for putative creatinergic neurotransmission. This can take much longer in the CNS than the PNS. Some commonly accepted neurotransmitters have never satisfied this criterion in a strict sense. The mechanism of Cr removal criterion is met not only by the Cr uptake in brain slices 220 and in synaptosomes 221, but also by our demonstration that SLC6A8 is required for synaptosome uptake of Cr.

The four criteria of Kandel et al.70,71 are mostly satisfied with some details requiring further research. The synthesis requirement is usually not strict because there are transmitters synthesized in some cells and transported into others where they function as transmitters. Our discovery of Cr in SVs can replace the synthesis requirement because the presence in neuronal SVs provide sufficient evidence that Cr is located in the right location to function as a neurotransmitter. The level of Cr in SVs is higher than those of ACh and 5-HT (Figs. 1 and 2). The amount of released Cr is in the same order of magnitude as those of Glu and GABA (Figs. 5 and 6). The criterion of a specific mechanism of removal was met by Cr uptake experiments in slices 220 and in synaptosomes 221, and further strengthened by our finding of SLC6A8 involvement in synaptosome uptake of Cr (Figure 8).

Here we report that Cr, at a concentration comparable to classical neurotransmitters, inhibits pyramidal neurons in specific regions of the mouse brain, with approximate 1/3 of pyramidal neurons responding to 100 μM Cr (Figure 7). In previous reports, 100 μM to 2 mM of GABA226–228, 50 μM to 2 mM of Glu 229,230, 1 to 100 μM of DA229, 0.1-100 μM of 5-HT231–236 were bath applied to investigate the physiological functions of neurotransmitters. Our results revealed that, when bath applied, Cr could inhibit cortical neurons at 100 μM within several minutes, with a time course similar to that of 5-HT237 234 236 and DA 229, but significantly slower than that of Glu230, GABA48,227 and 5-HT233.

In a recent report, knockout of the Slc6a8 gene increased excitation of cortical neurons267. Electrophysiological characterization of pyramidal neurons in the prefrontal cortex (PFC) found increased evoked firing frequency. Because we have shown that Cr inhibit a fraction of pyramidal neurons in the neocortex (Figure 8), this paper provides in vivo evidence consistent with the possibility of Cr as an inhibitory neurotransmitter.

Differences between Cr and classic neurotransmitters

At this point, we do not have a molecularly defined receptor for Cr, only inferring its presence from the electrophysiological responses to Cr. We speculate that Cr may act on G-protein coupled receptors (GPCRs), rather than the fast-acting ligand-gated ion channels, such as AMPA or NMDA receptors for Glu and GABAA receptor for GABA. There have been previous reports of Cr effects on neurons, including Cr as a partial agonist for GABAA receptors 238–241. These effects require very high concentrations of Cr (in the ten millimolar range). There was also a report of the opposite effect: that Cr (at a concentration above 500 μM) increased neuronal excitability through NMDA receptors after incubation for 60 mins, with a time course significantly slower than those of classic neurotransmitters 242.

Ca2+ independent component of Cr release induced by extracellular K+ was more prominent than those of Glu or GABA. One possibility was that Ca2+-independent Cr release came from glia, because high GAMT levels were reported in astrocytes261 and oligodendrites 261, 262. As reported, other neuromodulators such as taurine can be released from astrocytes 263 or slices 264 in an Ca2+ independent manner. In addition, in the absence of potassium stimulation, Ca2+ depletion increased release of taurine in cultured astrocytes 265 or in striatum in vivo266. Similarly, in SLC6A8 KO slices, Ca2+ depletion (Figure 5G) also increased Cr baseline as compared to that in normal ACSF (Figure 5D).

With much longer history of research, ACh and 5-HT now have more evidence in other aspects than Cr as a central transmitter, especially because there are many agonists and antagonists for ACh and 5-HT to prove an additional criterion that is required in some 222,223, but not the majority of, textbooks for a neurotransmitter. The pharmacology criterion will take some time and efforts, because, so far, no efforts have been made to find agonists or antagonists for Cr.

Implications of SLC6A8 and Cr

It is notable that SLC6A8 belongs to the neurotransmitter transporter family, with multiple members already shown to transport neurotransmitters 65,66,90–99.

The uptake experiments by others and us indicate that SLC6A8 transports Cr into neurons within the brain. AGAT is also expressed in the brain, but in cells not expressing SLC6A8 243,244. Cr and its precursor were thought to be transported between different cells in the nervous system. When SLC6A8 was completely missing, such as in homozygous SLC6A8 deficient patients, Cr treatment was not effective. But, if

SLC6A8 was partially active, Cr was effective 245. Intractable epilepsy in a female with heterozygous SLC6A8 mutation was completely treated by Cr 246. Our data of inhibitory effect of creatine on cortical neurons might provide a new mechanism to its anti-epileptic activity 247.

The absence of SLC6A8 expression in astrocytes whose endfeet lining microcapillary endothelial cells (MCEC) form the blood-brain barrier (BBB) indicates that Cr in the brain does not rely on import from the periphery and is instead mainly synthesized in the brain 80,82,248. SLC6A8 functions within the brain to transport Cr and its precursors, not as a major contributor of Cr transport across the BBB. It is thought to mediate Cr uptake into the presynaptic terminal, based on studies of synaptosomes (Figure 8) 221.

Cr is known to have effects other than an energy source, and Cr supplement has been thought to be beneficial for children, pregnant and lactating women, and old people 79,249. Cr has been reported to improve human mental performance 250–255. Cr has been used as potential treatment in animal models of neurodegenerative diseases 256,257. Our work will stimulate further research to distinguish which of the previously suspected effects of Cr is not attributed to its role as an energy storage, but can be attributed to its role as a neurotransmitter.

Search for new neurotransmitters

Our work may stimulate the search for more neurotransmitters. Our discovery indicates that the hunt for neurotransmitters stopped decades ago because of technical difficulties, not due to absence of more neurotransmitters. The fact that most of the known small molecule neurotransmitters have been found because of their peripheral effects also argues that what is missing is the concerted efforts to uncover central neurotransmitters with no peripheral effects. New neurotransmitters may be discovered from candidates which have been long suspected, and from previously unsuspected molecules or even previously unknown molecules.

Innovative approaches should be taken to uncover molecules with no previous suspicions or hints. Highly purified SVs, SVs from different regions of the brain, SVs with specific SLCs offer some of the starting points for future research.

Materials and methods

Generation of knockout and knockin mice.

SLC6A8 knockout and knockin mice were generated using CRISPR-Cas9-mediated genome engineering techniques by Beijing Biocytogen (Beijing, China). AGAT ‘knockout-first’ 185 mice were purchased from CAM-SU GRC (Suzhou, China). All mutations were validated by southern blot analysis, tail junction PCR and DNA sequencing.

Reverse-transcription polymerase chain reaction (RT-PCR) and quantitative real-time PCR (qPCR)

Total RNA of whole brains from mice of different genotypes was extracted using the Buffer RZ (Tiangen, no. RK14, Beijing, China) and reverse transcribed into complementary DNA (cDNA) using the RevertAid First-Strand cDNA synthesis kit (Thermo Scientific, K1622, USA). qPCR was performed using the Taq Pro Universal SYBR qPCR Master Mix (Vazyme, Q712-02) on Bio-Rad CFX-96 Touch Real-time PCR System (Bio-Rad, USA). Glyceraldehyde-3-phosphate dehydrogenase (GAPDH) was used as an internal control. ΔCt (difference in cycle threshold) was calculated for each sample (ΔCt = Ct Target gene − Ct GAPDH) for further evaluation of relative mRNA expression levels in different genotypes. The sequence specificities of the primers were examined. Three pairs of primers targeting different genes were used: SLC6a8 forward, 5’-GTCTGGTGACGAGAAGAAGGG-3’, Slc6a8 reverse, 5’-CCACGCACGACATGATGAAGT −3’; AGAT forward, 5’-CACAGTGGAGGTGAAGGCCAATACATAT-3’, AGAT reverse, 5’-CCGCCTCACGGTCACTCCT-3’; GAPDH forward, 5’-AGGTCGGTGTGAACGGATTTG −3’, GAPDH reverse, 5’-TGTAGACCATGTAGTTGAGGTCA-3’.

Primers for reverse PCR were designed to obtain complete coding sequences based on information obtained from National Center for Biotechnology Information (NCBI): SLC6A8 forward, 5’-ATGGCGAAAAAGAGCGCTGAAAACG-3’; Slc6a8 reverse, 5’-TTACATGACACTCTCCACCACGACGACC-3’; AGAT forward, 5’-ATGCTACGGGTGCGGTGTCT-3’; AGAT reverse, 5’-TCAGTCAAAGTAGGACTGAAGGGTGCCT-3’. PCR products were electrophoresed on 1% agarose gels, stained with GelRed, visualized under UV illumination and photographed.

Immunoblot analysis

Samples were loaded onto 10% polyacrylamide gels with the PAGE system (#1610183, Bio-Rad Laboratories, USA) and run in the SDS running buffer (25 mM Tris, 192 mM glycine, 0.1% SDS, pH 8.8) for 25 min at 80 V followed by 25 - 45 min at 200 V. Afterward, proteins were transferred to immobilon NC transfer membranes (HATF00010, Millipore) at 400 mA for 2 h in transfer buffer (25 mM Tris, 192 mM glycine, 20% methanol). Membranes were blocked in 5% fat-free milk powder in TBST (25 mM Tris, 150 mM NaCl, 0.2%Tween-20 (P1397, Sigma), pH 7.4 adjusted with HCl) and incubated overnight with the indicated primary antibodies dissolved in TBST containing 2% BSA.

Primary antibodies are listed below: rabbit anti-Synaptophysin (dilution 1:5000, cat no. 101002, SySy, Germany), rabbit anti-Synaptotagmin1/2 (dilution 1:2000, cat no.105002, SySy, Germany), rabbit anti-proton ATPase (dilution 1:1000, cat no.109002, SySy, Germany), rabbit anti-Synaptobrevin 2 (dilution 1:5000, cat no.104202, SySy, Germany), rabbit anti-SV2A (dilution 1:2000, cat no.109003, SySy, Germany), rabbit anti-VGlut1(dilution 1:4000, 135302, SySy, Germany), rabbit anti-VGlut2 (dilution 1:2000, 135402, SySy, Germany), rabbit anti-VGAT(dilution

1:4000, 131002, SySy, Germany), rabbit anti-SNAP23 (dilution 1:2000, cat no. 111202, SySy, Germany), mouse anti-PSD95 (dilution 1:5000, cat no 75028,NeuroMab), mouse anti-GluN1(dilution 1:5000, cat no 114011, SySy, Germany), rabbit anti-GM130 (dilution 1:1000, cat no ab52649, Abcam), rabbit anti-Golgin-97 (dilution 1:2000, cat no13192, Cell Signaling Technology, USA), rabbit anti-EEA1 (dilution 1:2000, cat no 3288, Cell Signaling Technology, USA), rabbit anti-LC3B (dilution 1:1000, cat no 2775S, Cell Signaling Technology, USA), goat anti-CathepsinB (dilution 1:2000, AF965, R&D Systems), rabbit anti-GAPDH (dilution 1:1000, cat no 2118S, Cell Signaling Technology, USA), rabbit anti-GluT4 (dilution 1:1000, ab33780, Abcam), rabbit anti-CACNA1A (dilution 1:300, 152103, SySy, Germany), rabbit anti-VDAC (dilution 1:1000, cat no 4661S, Cell Signaling Technology, USA), rabbit anti-MBP (1:1000, cat no. 295003, SySy, Germany), mouse anti Creatine Kinase B (dilution 1:5000, cat no MAB9076, R&D Systems), rabbit anti HA (dilution 1:2000, CST3724; Cell Signaling Technology, USA) and rabbit anti SNAP25 (dilution 1:2000, ab109105, Abcam) antibodies.

Membranes were washed in three washing steps in TBST (each for 5 mins) and incubated with peroxidase-conjugated secondary antibodies for 2-3 h at 4 °C. The second antibody used were anti-rabbit (dilution 1:5000, A6154, Sigma), anti-mouse (dilution 1:5000, 715-035-150, Jackson Immuno Research) or rabbit anti-goat IgG secondary antibodies (dilution 1:1000, cat no ab6741, Abcam). After repeated washing, signals were visualized using a ChemiDoc XRS+ System (Bio-RAD Laboratories, USA).

Isolation of synaptic vesicles

Our purification procedures for synaptic vesicles were based on previously established immunoisolation methods 143,151. Protein G magnetic beads (cat no.88848, Thermofisher Scienctific) were washed three times with IP buffer (100 mM potassium tartrate, 4 mM HEPES-KOH, 2 mM MgCl2, pH 7.4) supplemented with a complete protease inhibitor cocktail (Roche). 5 μg monoclonal anti-Syp antibody directed against a cytoplasmic epitope (cat no 101011, synaptic systems, Germany) or control mouse IgG (10400C, Thermofisher Scientific) was used to incubate with 20-30 μl beads for 30 mins at RT in 2% BSA dissolved in IP buffer. Under this condition, 4-4.5 μg of antibody was coupled, as determined by Western blot and Commassie Blue staining. Immunoisolation of SVs were carried out at 0 °C-2 °C to prevent vesicular content leakage (with RT as a control). Briefly, the whole mouse brain was homogenized in 3 ml of IP buffer with a glass/Teflon homogenizer (20 strokes at 2000 rpm, WHEATON, USA and WIGGENS WB2000-M, Germany) immediately after decapitation. Homogenates were centrifuged for 25 mins at 35,000 x g and the supernatant was adjusted to approximately 3 mg/ml protein (Nanodrop 2000 C, Thermofisher Scientific). To capture the SVs for content detection, about 200 μl of supernatants (per 5 μg anti-Syp/IgG) were incubated with pre-coupled beads for 2.25 h under slow rotation at 2 °C. Beads were washed 6 times for further Western blot analysis and vesicular content detection. For pharmacological blockade of H+-gradient across SV membrane, the mix of supernatants and pre-coupled beads was diluted into 1.2 ml before the addition of inhibitors.

Determination of vesicular contents

To extract SV contents, immunoisolates were treated with 50 μl ultra-pure water. 100 μl methanol together with 100 μl acetonitrile were added to precipitate proteins in samples. After centrifugation for 20 mins at 168000 × g, supernatants were collected and centrifuged for 20 mins at 2000 × g to remove beads and proteins. Samples were pre-frozen with liquid nitrogen and vacuum dried at −45 °C overnight. Dried samples were kept frozen and re-suspended with 50 μl of 0.2 μM 13C-creatine (internal control) immediately before detection.

Capillary electrophoresis-mass spectrometry (CE-MS) was used to verify and quantify small molecules. CE/MS detection was applied with the coupling of PA800 plus CE system (Beckman Coulter, Brea, CA, USA) and mass spectrometry (TRIPLE QUAD 5500, AB SCIEX or Q Exactive HF-X, Thermo Scientific). Before SV content detection, we optimized MS detection of classical neuro transmitters, Cr and amino acids in positive ion mode. Firstly, the fragment ions (Q3) for a given molecule (precursor ions, Q1) were determined by either systematic scanning of standard sample solution (0.1 μM in 10% acetate acid) or referring to database (www.mzcloud.org). Secondly, optimal values of collision energy (CE), collision cell exit potential (CXP) and declustering potentials (DP) were determined for each pair of Q1/Q3. Thirdly, optimal combination of parameters (Q1/Q3, CE, CXP, DP) was chosen for each molecule. In addition, parameters were adjusted every 2 to 3 months for best signal to noise ratios.

CE/MS separations were carried out by capillaries (OptiMS silica surface cartridge, BECKMAN COULTER). The CE background electrolyte was 10 % acetate acid. Each new separation capillary was activated with rinsing under100 psi sequentially with methanol for 10 mins (forward), methanol for 3 mins (reverse), H2O for 10 mins (forward), H2O for 3 mins (reverse), 0.1 M NaOH for 10 mins (forward), water for 5 mins (reverse), 0.1 M HCl for 10 mins (forward), followed by water for 10 mins and then 10 % acetate acid for 10 mins (forward) and 3 mins (reverse), prior to the first use. Between analyses, the capillary was rinsed with 10% acetate acid under a 100 psi pressure for 5 min (forward) flowed by 75 psi for 4 min. The sample (50 μl) was injected with 2.5-4 psi for 30 s. Separation voltage of 25 kV was applied for 25 mins. To maintain stably spray during CE separation, ion spray voltage was applied at 1.7-1.9 kV. MS data were collected 5 mins after CE separation. Finally, the capillary was washed with 10% acetate acid for 10 mins, followed by methanol for 20 min and then10% acetate acid for 20 mins.

Standard solutions of 0.2 μM 13C-Cr (internal control) and analytes were used to plot standard curves. Linear standard curves (R2>0.98, for most cases, R2>0.99), calculated from peak area ratios corresponding to analytes and internal standards, were obtained for all molecules tested. The concentration ranges used for standards of Glu, GABA, ACh, 5-HT, Cr and alanine were 0.03-10 μM, 0.003-1 μM, 0.0003-0.1 μM, 0.003-1 μM, 0.03-1 μM and 0.03-1 μM, respectively. Standard curves were made at least twice for a given capillary. Analytes of SV contents were calculated using the standard curves and normalized to the amount of anti-Syp antibody conjugated to the beads.

Electron microscopy

All EM grids were glow discharged for 30 s using a plasma cleaner (Harrick PDC-32G-2). To free SVs from beads, 25 μl 0.1 M glycine-HCl (PH=2) was incubated for 1 min and quickly neutralized with 25 μl 0.1 M Tris (pH=10). Beads were quickly removed and 2-4 μL aliquots of SVs were applied to the carbon coated copper grids (Zhong Jing Ke Yi, Beijing). After 1 min, the grid was dried with a filter paper (Whatman No.1), and placed in the water, and then immediately stained using 2 % uranyl acetate for 30 s. At last uranyl acetate was removed and the grid was air dried. The grids were examined on a JEM-F200 electron microscope (JEOL, Japan) operated at 200 kV. Images were recorded using a 4k × 4k COMS One view camera (Gatan). Fixation of synaptosomal pellets was performed by immersion with pre-warmed 2.5% glutaraldehyde in 0.1 M phosphate buffer (pH 7.4) at RT for 2 h. After washing 4 times with 0.1 M phosphate buffer (pH 7.4) every 15 mins, samples were post-fixed with 1% osmium tetroxide (w/v) at 4°C for 1 h and then washed 3 times. Following en bloc staining with 2% uranyl acetate (w/v) at 4°C overnight, samples were dehydrated and embedded in fresh resin, polymerized at 65°C for 24 h. Ultrathin (70 nm) sections were obtained by Leica UC7 ultramicrotome (Leica Microsystems, German) and recorded on 80 kV in a JEOL Jem-1400 transmission electron-microscope (JEOL, Japan) using a CMOS camera (XAROSA, EMSIS)

Immunohistochemistry

Adult mice were anesthetized by i.p. injection with 2 % 2,2,2-tribromoethanol (T48402, Sigma Aldrich) in saline at a dose of 400 mg/kg, and perfused trancardially with 0.9 % saline followed by 4 % PFA in PBS (137 mM NaCl, 2.7 mM KCl, 10 mM Na2HPO4, 1.8 mM KH2PO4, pH=7.4).

Brains were cryoprotected with 30 % sucrose in 30 % sucrose 0.1 M PB (81 mM Na2HPO4, 19 mM NaH2PO4) and sectioned in the coronal plane (40 μm thick) using a Cryostat (Leica 3050S). For anti-HA immunostaining, we used a rabbit monoclonal anti-HA antibody (1:500 in 0.3 % Triton in PBS; 48 h incubation at 4 °C; #3724, Cell Signaling Technology), followed by a goat anti-rabbit Alexa Fluor 546 secondary antibody (1:1000; overnight at 4 °C; # A-11035, Invitrogen). Sections were mounted in a medium containing 50 % glycerol, cover-slipped, and sealed with nail polish. Images were acquired using virtual slide microscope (Olympus VS120-S6-W) and a laser-scanning confocal microscope (Zesis 710) and brain structures inferred with an established mouse brain atlas 258.

Preparations of brain slices

Male C57 mice (of 30-38 days-old) were anesthetized with pentobarbital (250 mg/kg) and decapitated. Brains were quickly removed and placed into ice-cold, low calcium, high magnesium artificial cerebrospinal fluid (ACSF) with sodium replaced by choline.

The medium consisted of 120 mM choline chloride, 2.5 mM KCl, 7 mM MgSO4, 0.5 mM CaCl2, 1.25 mM NaH2PO4, 5 mM sodium ascorbate, 3 mM sodium pyruvate, 26 mM NaHCO3, and 25 mM D-(+)-glucose, and was pre-equilibrated with 95% O2-5% CO2. Coronal brain slices (300 μm thick) were cut with a vibratome (Leica VT1200S). Slices were incubated for 1 h at 34°C with oxygenated ACSF containing: 124 mM NaCl, 2.5 mM KCl, 2 mM MgSO4, 2.5 mM CaCl2, 1.25 mM NaH2PO4, 26 mM NaHCO3, and 10 mM D-(+)-glucose.

Evoked release from brain slices

Coronal brain slices (each 300 μm thick, typically with a wet weight of 17-20 mg) were transferred into a specially designed superfusion chamber with a volume of approximately 200 μl, containing freshly 95% O2/5% CO2 oxygenated ACSF. Slices were equilibrated for 10 min in ACSF at a superfusion rate of 0.9-1.25 ml/min. The “control” sample was collected for 1 min just before high K+ stimulation (K-ACSF, 70 mM KCl replacing equal amount of NaCl). We waited for 30 seconds to allow K+ stimulus to immerse the slices (dead volume for solution transition of 200 μl and chamber volume of 200 μl), then the sample “70 mM K” in response to K-ACSF was collected for another 1 min. Following 10 min of washout period, we collected the third sample of “wash” for 1 min.

To detect Ca2+ dependent release, slices were pre-incubated for 10 mins with normal ACSF and equilibrated with Ca2+ free ACSF (containing 1 mM EGTA to chelate extracellular Ca2+) for 10 mins. The baseline sample “0 Ca2+ ACSF” was collected for 1 min. Superfusion solution was changed to Ca2+ free K-ACSF for 2 mins and sample “0 Ca2+ 70 mM K” was collected (dead volume for solution transition of 400 μl and chamber volume of 200 μl). After that, the solution was changed back to normal ACSF for 10 mins and K-ACSF for 2 mins. The sample “2.5 mM Ca 70 mM K” for the last min was collected.

Samples were subjected to CE-MS in a method similar to SV content detection, except for the following: 1) standards were dissolved in ACSF or other buffers used in release experiment; 2) concentration ranges used for standards of Glu was from 0.003-1 μM. 3) To protect the MS from salt pollution, data were collected from 10 mins −20 mins during CE separation.

Patch-Clamp recordings

Slices were transferred to a recording chamber on an upright fluorescent microscope equipped with differential interference contrast optics (DIC; Olympus BX51WI). Slices were submerged and superfused with ACSF at about 2.8 ml/min at 24-26 °C. Whole cell patch recordings were routinely achieved from layer 4/5 medium sized pyramidal neurons from the somatosensory cortex. Patch pipettes (3-5 MΩ) contained: 140 mM K-gluconate, 10 mM HEPES, 0.5 mM EGTA, 5 mM KCl, 3 mM Na2-ATP, 0.5 mM Na3GTP, and 4 mM MgCl2 (with pH adjusted to 7.3 and osmolarity of 290 mOsm/kg). Current-clamp recordings were carried out with a computer-controlled amplifier (Multiclamp 700B, Molecular Devices) and traces were digitized at 10 kHz (DigiData 1550B, Molecular Devices). Data were collected and analyzed using Clampfitor Clampex 10 software (Molecular Devices).

Cells were characterized by their membrane responses and firing patterns during hyperpolarizing and depolarizing current steps (−100 to +500 pA, increment: 50 pA or 25 pA, 500 ms). Regular spiking pyramidal neurons were identified by moderate maximal spiking frequencies (20-60 Hz, that is, 10-30 spikes per 500 ms, Figure 7E-G), increasing of inter-spike intervals during depolarizing step (Figure 7B and Supplementary Figure 11A), high action amplitude (Supplementary Figure 12E) and large half width (Supplementary Figure 12F) 201,202. After the mean firing frequency evoked by current injections reached the steady state for at least 5 mins (typically 20-30 mins following the formation of whole-cell configuration), 100 μM Cr was bath-applied for 6 mins. Typically, Cr was applied for a 2nd time following washout to re-confirm the effects.

Synaptosome preparation

Synaptosomes were isolated by Ficoll/sucrose density-gradient centrifugation 203,209,221,259. Whole brains from adult male mice were homogenized with 15 strokes at 900 rpm in buffer A (320 mM sucrose, 1 mM EDTA, 1mM EGTA, 10 mM Tris–HCl, pH 7.4, with a complete protease inhibitor cocktail (Roche)). The homogenate (H fraction) was centrifuged at 1000 g for 10 min to precipitate the membrane fragments and nuclei (P1 fraction). Supernatant was centrifuged again at 1000 g for 10 min and the resulting supernatant (S1) was centrifuged at 12,000 g for 20 min. Supernatant was the S2 fraction, and the pellet was re-suspended with buffer A and centrifuged at 12,000 g for 20 min. The resulting pellet was crude synaptosomes (P2 fraction), containing synaptosomes with mitochondria and microsomes.

Crude synaptosomes (P2 fraction) was re-suspended with 150-200 μl buffer B (320 mM sucrose and 10 mM Tris–HCI (pH 7.4)). The sample was carefully overlaid on the top of a gradient of 2 ml of 7.5 % (wt/vol in buffer B) Ficoll and 1.8 ml of 13% (wt/vol in buffer B) Ficoll and centrifuged at 98,000 g for 45 min at 2-4 °C in a swinging-bucket rotor. A myelin band was present near the surface, and th synaptosomes band (fraction Sy) was present at the interface between the 13% and 7.5 % Ficoll layers, with the mitochondria being pelleted at the bottom. For further western analysis, the supernatant was divided into 6 fractions (600 μl for each fraction) and the mitochondria pellet was discarded. The isolated synaptosomes was included in fraction 4.

For Western analysis, fractions H, S1, P1, S2, P2 and Sy were adjusted to 0.5 mg/ml by bicinchoninic acid assay (BCA) method with reference to NanoDrop™ 2000 Spectrophotometers. 3.35 μg protein was loaded for each lane. Fractions 1-6 were loaded with the same volume (10 μl composed of 6.7 μl sample and 3.3 μl loading buffer) for each lane.

Creatine uptake into synaptosomes

To remove Ficoll, we diluted the synaptosomal band (480 μl) with 4.3 ml of a pH 7.4 buffer C containing (in mM) 240 mannitol, 10 glucose, 4.8 Kgluconate, 2.2 Cagluconate, 1.2 MgSO4, 1.2 KH2PO4 and 25 HEPES-Tris. The sample was then centrifuged at 12,000 g and the pellet was re-suspended with buffer C. Uptake experiments were either performed at 37°C or at 0°C (control). For each sample, 25-43 μg of synaptosomes (with a volume of 40-50 μl) were added to 360 μl buffer containing (in mmol/L) 100 NaCl, 40 mannitol, 10 glucose, 4.8 Kgluconate, 2.2 Cagluconate, 1.2 MgSO4, 1.2 KH2PO4, 25 HEPES and 25 Tris (pH adjusted to 7.4). A mixture of 18 μM [14C]-creatine (0.4 μCi) and 5 μM creatine was quickly added. After 10 min, uptake was terminated by the addition of 1 ml of NaCl-free ice-cold buffer C. Samples were immediately filtered, under vacuum, through a Whatman GF/C glass filter (1825-025) pre-wetted with buffer C. Filters were further washed with 10 ml of ice-cold buffer C, dissolved in scintillation fluid and the radioactivity determined by liquid scintillation spectrometry.

Creatine uptake into SVs

The uptake of 13C-creatine was assayed according to a conventional procedure 260 with slight modifications: the immunoisolated synaptic vesicles by 10 μg Syp antibody (101011, Synaptic Systems, Germany) were resuspended with the uptake buffer (150mM meglumine-tartrate, 4 mM KCl, 4 mM MgSO4, 10mM HEPES-KOH pH 7.4, and complete EDTA-free protease inhibitor cocktail) containing 4mM Mg-ATP or additional 4mM MgSO4, followed by preincubation for 30 min at 25L. The uptake reaction was started by addition of 1mM 13C-creatine dissolved in the uptake buffer with a final volume of 125 μL (pH at 6.8). After 10min at 25L, 1mL of ice-cold uptake buffer was added to the incubation to stop the reaction, followed by 5 more times washing. The SV contents were extracted using the protocol described in the determination of vesicular contents part. 100 nM Cr was used as the internal control. CE-MS and LC-MS was used to verify and quantify the creatine contents of samples. A Vanquish UHPLC system coupled to a Q Exactive HF-X mass spectrometer (both instrument Thermo Fisher Scientific, USA) were used for LC-MS analysis along with SeQuant ZIC-HILIC column (150 mm × 2.1 mm, 3.5 μm, Merck Millipore, 150442) in the positive mode and SeQuant ZIC-pHILIC column (150 mm × 2.1 mm, 5 μm, Merck Millipore, 150460) in the negative mode. For ZIC-HILIC column, the mobile phase A was 0.1% formic acid in water and the mobile phase B was 0.1% formic acid in acetonitrile. The linear gradient was as following: 0 min, 80% B; 6 min, 50% B; 13 min, 50% B; 14 min, 20% B; 18 min, 20% B; 18.5 min, 80% B; 30 min, 80% B. The flow rate used was 300 μl/min and the column temperature was maintained at 30°C. For ZIC-pHILIC column, the mobile phase A is 20 mM ammonium carbonate in water, adjusted to pH 9.0 with 0.1% ammonium hydroxide solution (25%) and the mobile phase B is 100% acetonitrile. The linear gradient was as follows: 0 min, 80% B; 2 min, 80% B; 19 min, 20% B; 20 min, 80% B; 30 min, 80% B. The flow rate used is 150 μl/min and the column temperature is 25°C. Samples were maintained at 4°C in Vanquish autosampler. 3 µl of extracted metabolites were injected for each run. IP samples were subjected to ZIC-HILIC column in positive mode for major metabolites detection, and then subject to ZIC-pHILIC column in negative mode for orthogonal detection.

Supporting information

eLife Assessment

This study presents valuable observations on the role of creatine (Cr) in the context of synaptic transmission. Overall, the data are solid in support of the conclusion that Cr is present in synaptic vesicles, although the evidence for Cr release and Cr-dependent modulation of neuronal function was considered incomplete. The work will be of general interest to the field of neuroscience.

Reviewer #1 (Public Review):

The overall tone of the rebuttal and lack of responses on several questions was surprising. Clearly, the authors took umbrage at the phrase 'no smoking gun' and provided a lengthy repetition of the fair argument about 'ticking boxes' on the classic list of criteria. They also make repeated historical references that descriptions of neurotransmitters include many papers, typically over decades, e.g. in the case of ACh and its discovery by Sir Henry Dale. While I empathize with the authors' apparent frustration (I quote: '...accept the reality that Rome was not built in a single day and that no transmitter was proven by a one single paper') I am a bit surprised at the complete brushing away of the argument, and in fact the discussion. In the original paper, the notion of a receptor was mentioned only in a single sentence and all three reviewers brought up this rather obvious question. The historical comparisons are difficult: Of course many papers contribute to the identification of a neurotransmitter, but there is a much higher burden of proof in 2023 compared to the work by Otto Loewi and Sir Henry Dale: most, if not all, currently accepted neurotransmitter have a clear biological function at the level of the brain and animal behavior or function - and were in fact first proposed to exist based on a functional biological experiment (e.g. Loewi's heart rate change). This, and the isolation of the chemical that does the job, were clear, unquestionable 'smoking guns' a hundred years ago. Fast forward 2023: Creatine has been carefully studied by the authors to tick many of the boxes for neurotransmitters, but there is no clear role for its function in an animal. The authors show convincing effects upon K+ stimulation and electrophysiological recordings that show altered neuronal activity using the slc6a8 and agat mutants as well as Cr application - but, as has been pointed out by other reviewers, these effects are not a clear-cut demonstration of a chemical transmitter function, however many boxes are ticked. The identification of a role of a neurotransmitter for brain function and animal behavior has reasonably more advanced possibilities in 2023 than a hundred years ago - and e.g. a discussion of approaches for possible receptor candidates should be possible.

Again, I reviewed this positively and agree that a lot of cumulative data are great to be put out there and allow the discovery to be more broadly discussed and tested. But I have to note, that the authors simply respond with the 'Rome was not built in a single day' statement to my suggestions on at least 'have some lead' how to approach the question of a receptor e.g. through agonists or antagonists (while clearly stating 'I do not think the publication of this manuscript should not be made dependent' on this). Similarly, in response to reviewer 2's concerns about a missing receptor, the authors' only (may I say snarky) response is ' We have deleted this sentence, though what could mediate postsynaptic responses other than receptors?' The bullet point by reviewer 3 ' • No candidate receptor for creatine has been identified postsynaptically.' is the one point by that reviewer that is simply ignored by the authors completely. Finally, I note that my reivew question on the K stimulation issues (e.g. 35 neurons that simply did not respond at all) was: ' Response: To avoid the disadvantage of K stimulation, we also performed optogenetic experiments recently and obtained encouraging preliminary results.' No details, not data - no response really.

In sum, I find this all a bit strange and the rebuttal surprising - all three reviewers were supportive and have carefully listed points of discussion that I found all valid and thoughtful. In response, the authors selectively responded scientifically to some experimental questions, but otherwise simply rather non-scientifically dismissed questions with 'Rome was not built in a day'-type answers, or less. I my view, the authors have disregarded the review process and the effort of three supportive reviewers, which should be part of the permanent record of this paper.

Reviewer #2 (Public Review):

Summary:

Bian et al studied creatine (Cr) in the context of central nervous system (CNS) function. They detected Cr in synaptic vesicles purified from mouse brains with anti-Synaptophysin using capillary electrophoresis-mass spectrometry. Cr levels in the synaptic vesicle fraction was reduced in mice lacking the Cr synthetase AGAT, or the Cr transporter SLC6A8. They provide evidence for Cr release within several minutes after treating brain slices with KCl. This KCl-induced Cr release was partially calcium dependent and was attenuated in slices obtained from AGAT and SLC6A8 mutant mice. Cr application also decreased the excitability of cortical pyramidal cells in one third of the cells tested. Finally, they provide evidence for SLC6A8-dependent Cr uptake into synaptosomes, and ATP-dependent Cr loading into synaptic vesicles. Based on these data, the authors propose that Cr may act as neurotransmitter in the CNS.

Strengths:

1. A major strength of the paper is the broad spectrum of tools used to investigate Cr.

2. The study provides evidence that Cr is present in/loaded into synaptic vesicles.

Weaknesses:

1. There is no significant decrease in Cr content pulled down by anti-Syp in AGAT-/- mice when normalized to IgG controls. Hence, blocking AGAT activity/Cr synthesis does not affect Cr levels in the synaptic vesicle fraction, arguing against a Cr enrichment.

2. There is no difference in KCl-induced Cr release between SLC6A8-/Y and SLC6A8+/Y when normalizing the data to the respective controls. Thus, the data are not consistent with the idea that depolarization-induced Cr release requires SLC6A8.

3. The rationale of grouping the excitability data into responders and non-responders is not convincing because the threshold of 10% decrease in AP rate is arbitrary. The data do therefore not support the conclusion that Cr reduces neuronal excitability.

Reviewer #3 (Public Review):

SUMMARY:

The manuscript by Bian et al. promotes the idea that creatine is a new neurotransmitter. The authors conduct an impressive combination of mass spectrometry (Fig. 1), genetics (Figs. 2, 3, 6), biochemistry (Figs. 2, 3, 8), immunostaining (Fig. 4), electrophysiology (Figs. 5, 6, 7), and EM (Fig. 8) in order to offer support for the hypothesis that creatine is a CNS neurotransmitter.

STRENGTHS:

There are many strengths to this study.

• The combinatorial approach is a strength. There is no shortage of data in this study.

• The careful consideration of specific criteria that creatine would need to meet in order to be considered a neurotransmitter is a strength.

• The comparison studies that the authors have done in parallel with classical neurotransmitters is helpful.

• Demonstration that creatine has inhibitory effects is another strength.

• The new genetic mutations for Slc6a8 and AGAT are strengths and potentially incredibly helpful for downstream work.

WEAKNESSES:

• Some data are indirect. Even though Slc6a8 and AGAT are helpful sentinels for the presence of creatine, they are not creatine themselves. Of note, these molecules themselves are not essential for making the case that creatine is a neurotransmitter.

• Regarding Slc6a8, it seems to work only as a reuptake transporter - not as a transporter into SVs. Therefore, we do not know what the transporter into the TVs is.

• Puzzlingly, Slc6a8 and AGAT are in different cells, setting up the complicated model that creatine is created in one cell type and then processed as a neurotransmitter in another. This matter will likely need to be resolved in future studies.

• No candidate receptor for creatine has been identified postsynaptically. This will likely need to be resolved in future studies.

• Because no candidate receptor has been identified, it is important to fully consider other possibilities for roles of creatine that would explain these observations other than it being a neurotransmitter? There is some attention to this in the Discussion.

There are several criteria that define a neurotransmitter. The authors nicely delineated many criteria in their discussion, but it is worth it for readers to do the same with their own understanding of the data.

By this reviewer's understanding (and combining some textbook definitions together) a neurotransmitter: 1) must be present within the presynaptic neuron and stored in vesicles; 2) must be released by depolarization of the presynaptic terminal; 3) must require Ca2+ influx upon depolarization prior to release; 4) must bind specific receptors present on the postsynaptic cell; 5) exogenous transmitter can mimic presynaptic release; 6) there exists a mechanism of removal of the neurotransmitter from the synaptic cleft.

For a paper to claim that the published work has identified a new neurotransmitter, several of these criteria would be met - and the paper would acknowledge in the discussion which ones have not been met. For this particular paper, this reviewer finds that condition 1 is clearly met.

Conditions 2 and 3 seem to be met by electrophysiology, but there are caveats here. High KCl stimulation is a blunt instrument that will depolarize absolutely everything in the prep all at once and could result in any number of non-specific biological reactions as a result of K+ rushing into all neurons in the prep. Moreover, the results in 0 Ca2+ are puzzling. For creatine (and for the other neurotransmitters), why is there such a massive uptick in release, even when the extracellular saline is devoid of calcium?

Condition 4 is not discussed in detail at all. In the discussion, the authors elide the criterion of receptors specified by Purves by inferring that the existence of postsynaptic responses implies the existence of receptors. True, but does it specifically imply the existence of creatinergic receptors? This reviewer does not think that is necessarily the case. The authors should be appropriately circumspect and consider other modes of inhibition that are induced by activation or potentiation of other receptors (e.g., GABAergic or glycinergic).

Condition 5 may be met, because authors applied exogenous creatine and observed inhibition. However, this is tough to know without understanding the effects of endogenous release of creatine. if they were to test if the absence of creatine caused excess excitation (at putative creatinergic synapses), then that would be supportive of the same. Nicely, Ghirardini et al., 2023 study cited by the reviewers does provide support for this exact notion in pyramidal neurons.

For condition 6, the authors made a great effort with Slc6a8. This is a very tough criterion to understand or prove for many synapses and neurotransmitters.

In terms of fundamental neuroscience, the story should be impactful. There are certainly more neurotransmitters out there than currently identified and by textbook criteria, creatine seems to be one of them taking all of the data in this study and others into account.

Author Response

The following is the authors’ response to the previous reviews

Point-to-Point Responses to Reviewers’ Comments

We are a bit surprised by the comments of Reviewer 1, but that our further responses can help communications with Reviewer 1. We have also responded to comments of Reviewers 2 and 3.

Public Reviews:

*Reviewer #1 (Public Review):

The overall tone of the rebuttal and lack of responses on several questions was surprising. Clearly, the authors took umbrage at the phrase 'no smoking gun' and provided a lengthy repetition of the fair argument about 'ticking boxes' on the classic list of criteria. They also make repeated historical references that descriptions of neurotransmitters include many papers, typically over decades, e.g. in the case of ACh and its discovery by Sir Henry Dale. While I empathize with the authors' apparent frustration (I quote: '...accept the reality that Rome was not built in a single day and that no transmitter was proven by a one single paper') I am a bit surprised at the complete brushing away of the argument, and in fact the discussion. In the original paper, the notion of a receptor was mentioned only in a single sentence and all three reviewers brought up this rather obvious question. The historical comparisons are difficult: Of course many papers contribute to the identification of a neurotransmitter, but there is a much higher burden of proof in 2023 compared to the work by Otto Loewi and Sir Henry Dale: most, if not all, currently accepted neurotransmitter have a clear biological function at the level of the brain and animal behavior or function - and were in fact first proposed to exist based on a functional biological experiment (e.g. Loewi's heart rate change). This, and the isolation of the chemical that does the job, were clear, unquestionable 'smoking guns' a hundred years ago. Fast forward 2023: Creatine has been carefully studied by the authors to tick many of the boxes for neurotransmitters, but there is no clear role for its function in an animal. The authors show convincing effects upon K+ stimulation and electrophysiological recordings that show altered neuronal activity using the slc6a8 and agat mutants as well as Cr application - but, as has been pointed out by other reviewers, these effects are not a clear-cut demonstration of a chemical transmitter function, however many boxes are ticked. The identification of a role of a neurotransmitter for brain function and animal behavior has reasonably more advanced possibilities in 2023 than a hundred years ago - and e.g. a discussion of approaches for possible receptor candidates should be possible.

Again, I reviewed this positively and agree that a lot of cumulative data are great to be put out there and allow the discovery to be more broadly discussed and tested. But I have to note, that the authors simply respond with the 'Rome was not built in a single day' statement to my suggestions on at least 'have some lead' how to approach the question of a receptor e.g. through agonists or antagonists (while clearly stating 'I do not think the publication of this manuscript should not be made dependent' on this). Similarly, in response to reviewer 2's concerns about a missing receptor, the authors' only (may I say snarky) response is ' We have deleted this sentence, though what could mediate postsynaptic responses other than receptors?' The bullet point by reviewer 3 ' • No candidate receptor for creatine has been identified postsynaptically.' is the one point by that reviewer that is simply ignored by the authors completely. Finally, I note that my reivew question on the K stimulation issues (e.g. 35 neurons that simply did not respond at all) was: ' Response: To avoid the disadvantage of K stimulation, we also performed optogenetic experiments recently and obtained encouraging preliminary results.' No details, not data - no response really.

In sum, I find this all a bit strange and the rebuttal surprising - all three reviewers were supportive and have carefully listed points of discussion that I found all valid and thoughtful. In response, the authors selectively responded scientifically to some experimental questions, but otherwise simply rather non-scientifically dismissed questions with 'Rome was not built in a day'-type answers, or less. I my view, the authors have disregarded the review process and the effort of three supportive reviewers, which should be part of the permanent record of this paper.

Response:

We were very surprised by the tone of Reviewer 1 in the second round of reviewing. The corresponding author has spent some time including a long holiday to cool down and re-read our earlier responses. The following is entirely by the correspond author.

I have finally checked the term “smoking gun”, and found out that I interpreted it wrongly while I had thought that Reviewer 1 was wrong. This came from a long story in that I was lectured by a native speaker for my English when submitting the first paper from my own paper. In that case, the Reviewer was wrong (in arguing that only adjectives but not nouns can be used to define nouns), I was quite offended and remembered it vividly. In the case of “smoking gun”, I wrongly believed that it meant a hint (while the definite evidence would be “the final nail in the coffin”). By interpreting is as a hint, I was then rebutting Reviewer 1 for negating all our experimental results as “not a single piece of suggestive evidence”.

For the above, I apologize.

I have another disagreement about “smoking gun”. For a transmitter, multiple criteria have to be met. For example, finding a receptor for a small molecule would not be definitive for a transmitter because if it is not present in the SVs, it is unlikely to be a typical transmitter. If a molecule has a receptor but they are not even in the nervous system, it is definitely no a transmitter.

The title of our paper is “Evidence suggesting creatine as a new central neurotransmitter”, not “Evidence proving creatine as a new central neurotransmitter”. In the Abstract, after “Our biochemical, chemical, genetic and electrophysiological results are consistent with the possibility of Cr as a neurotransmitter”, we are adding “though not yet reaching the level of proof for the now classic transmitters”. In the last sentence of the introduction, we have now added “though the discovery of a receptor for Cr would prove it”.

I do, however, believe that, however strong the wordings are, criticisms and rebuttals in science are normal and should be conducted even when emotions are involved.

One of my major point of differences with at least two of the reviewers is that the criteria for neurotransmitters should be those listed in major textbooks. While everyone can have one’s own opinions, the textbooks, especially those accepted by readers of the field for more than 40 years, should be the standards. Kandel has listed the 4 criteria not only 40 years ago but also just 2 years ago in their latest 6th edition. The reviewers have asked for more, while discounting Kandel et al. (2021). So, in essence, the Reviewer is not shy in scientific criticisms when stating “The identification of a role of a neurotransmitter for brain function and animal behavior has reasonably more advanced possibilities in 2023 than a hundred years ago”.

Reviewer 1 raised another new criterion: brain function and behavior, while this is not in any textbook lists. However, lack of Cr caused behavioral problems, as cited by us in the introduction: both humans and mice were defective in brain function with loss of function mutations in the gene for the specific Cr transporter SLC6A8. If the reviewer meant behavioral abnormalities caused by Cr injection, that was unclear. But that criterion may not be met by other transmitters which is the likely reason that it was not a criterion in any textbook.

Reviewer #2 (Public Review):

Summary:

Bian et al studied creatine (Cr) in the context of central nervous system (CNS) function. They detected Cr in synaptic vesicles purified from mouse brains with anti-Synaptophysin using capillary electrophoresis-mass spectrometry. Cr levels in the synaptic vesicle fraction was reduced in mice lacking the Cr synthetase AGAT, or the Cr transporter SLC6A8. They provide evidence for Cr release within several minutes after treating brain slices with KCl. This KCl-induced Cr release was partially calcium dependent and was attenuated in slices obtained from AGAT and SLC6A8 mutant mice. Cr application also decreased the excitability of cortical pyramidal cells in one third of the cells tested. Finally, they provide evidence for SLC6A8-dependent Cr uptake into synaptosomes, and ATP-dependent Cr loading into synaptic vesicles. Based on these data, the authors propose that Cr may act as neurotransmitter in the CNS.

Strengths:

1. A major strength of the paper is the broad spectrum of tools used to investigate Cr.

2. The study provides evidence that Cr is present in/loaded into synaptic vesicles.

Weaknesses:

1. There is no significant decrease in Cr content pulled down by anti-Syp in AGAT-/- mice when normalized to IgG controls. Hence, blocking AGAT activity/Cr synthesis does not affect Cr levels in the synaptic vesicle fraction, arguing against a Cr enrichment.

Response: Evidence for Cr enrichment in the SVS was obtained robustly with wild type mice. When brain Cr is very low in AGAT-/- mutant mice, because there is little Cr, there is also little Cr in the SVs. One does not require that as a criterion: it does not argue against the normal levels of Cr could be transported into the SVs even if when the much reduced levels of AGAT-/- Cr in mutant mice could be enriched in SVs.

1. There is no difference in KCl-induced Cr release between SLC6A8-/Y and SLC6A8+/Y when normalizing the data to the respective controls. Thus, the data are not consistent with the idea that depolarization-induced Cr release requires SLC6A8.

Response: This comment of Reviewer 2 was based on Figure 5D. But if one carefully examines Figure 5G, it was clear that the Ca++ dependent component of KCl -induced Cr release was lower in SLC6A8-/Y than that in SLC6A8+/Y.

1. The rationale of grouping the excitability data into responders and non-responders is not convincing because the threshold of 10% decrease in AP rate is arbitrary. The data do therefore not support the conclusion that Cr reduces neuronal excitability.

Response: Comparison of the same neuron, before and after Cr did show effects on neuronal excitability though that would have no statistics if one does not group multiple cells into the same categories.

Reviewer #3 (Public Review):

SUMMARY:

The manuscript by Bian et al. promotes the idea that creatine is a new neurotransmitter. The authors conduct an impressive combination of mass spectrometry (Fig. 1), genetics (Figs. 2, 3, 6), biochemistry (Figs. 2, 3, 8), immunostaining (Fig. 4), electrophysiology (Figs. 5, 6, 7), and EM (Fig. 8) in order to offer support for the hypothesis that creatine is a CNS neurotransmitter.

STRENGTHS:

There are many strengths to this study. • The combinatorial approach is a strength. There is no shortage of data in this study. • The careful consideration of specific criteria that creatine would need to meet in order to be considered a neurotransmitter is a strength. • The comparison studies that the authors have done in parallel with classical neurotransmitters is helpful. • Demonstration that creatine has inhibitory effects is another strength. • The new genetic mutations for Slc6a8 and AGAT are strengths and potentially incredibly helpful for downstream work.

WEAKNESSES: • Some data are indirect. Even though Slc6a8 and AGAT are helpful sentinels for the presence of creatine, they are not creatine themselves. Of note, these molecules themselves are not essential for making the case that creatine is a neurotransmitter.

Response: We agree, but those data are not inconsistent with the possibility.

• Regarding Slc6a8, it seems to work only as a reuptake transporter - not as a transporter into SVs. Therefore, we do not know what the transporter into the TVs is.

Response: SLC6A8 is not the transporter on the SVs, but is an excellent candidate for the transporter on the presynaptic cytoplasmic membrane for uptake of Cr into the presynaptic structure.

• Puzzlingly, Slc6a8 and AGAT are in different cells, setting up the complicated model that creatine is created in one cell type and then processed as a neurotransmitter in another. This matter will likely need to be resolved in future studies.

Response: We agree.

• No candidate receptor for creatine has been identified postsynaptically. This will likely need to be resolved in future studies.

Response: We agree.

• Because no candidate receptor has been identified, it is important to fully consider other possibilities for roles of creatine that would explain these observations other than it being a neurotransmitter? There is some attention to this in the Discussion.

Response: We agree.

There are several criteria that define a neurotransmitter. The authors nicely delineated many criteria in their discussion, but it is worth it for readers to do the same with their own understanding of the data.

By this reviewer's understanding (and combining some textbook definitions together) a neurotransmitter: 1) must be present within the presynaptic neuron and stored in vesicles; 2) must be released by depolarization of the presynaptic terminal; 3) must require Ca2+ influx upon depolarization prior to release; 4) must bind specific receptors present on the postsynaptic cell; 5) exogenous transmitter can mimic presynaptic release; 6) there exists a mechanism of removal of the neurotransmitter from the synaptic cleft.

Response: While any of us can come up with a list according to our own understanding, the paper copies lists from textbooks, especially from Kandel et al. (2021), which lists the same 4 criteria as Kandel et al. (1983), providing consistency and consensus.

For a paper to claim that the published work has identified a new neurotransmitter, several of these criteria would be met - and the paper would acknowledge in the discussion which ones have not been met. For this particular paper, this reviewer finds that condition 1 is clearly met.

Conditions 2 and 3 seem to be met by electrophysiology, but there are caveats here. High KCl stimulation is a blunt instrument that will depolarize absolutely everything in the prep all at once and could result in any number of non-specific biological reactions as a result of K+ rushing into all neurons in the prep. Moreover, the results in 0 Ca2+ are puzzling. For creatine (and for the other neurotransmitters), why is there such a massive uptick in release, even when the extracellular saline is devoid of calcium?

Response: Classic transmitters are released in a Ca++ dependent manner when stimulated by KCl, though they also had a Ca++ independent component as also shown in our Figure 5 E and F.

Condition 4 is not discussed in detail at all. In the discussion, the authors elide the criterion of receptors specified by Purves by inferring that the existence of postsynaptic responses implies the existence of receptors. True, but does it specifically imply the existence of creatinergic receptors? This reviewer does not think that is necessarily the case. The authors should be appropriately circumspect and consider other modes of inhibition that are induced by activation or potentiation of other receptors (e.g., GABAergic or glycinergic).

Response: Kandel et al. did not list this.

Condition 5 may be met, because authors applied exogenous creatine and observed inhibition. However, this is tough to know without understanding the effects of endogenous release of creatine. if they were to test if the absence of creatine caused excess excitation (at putative creatinergic synapses), then that would be supportive of the same. Nicely, Ghirardini et al., 2023 study cited by the reviewers does provide support for this exact notion in pyramidal neurons.

Response: For most commonly accepted transmitters, this criterion has never been met. For example, the simplest case would be ACh at the neuromuscular junction. Howver, we have now found that choline is clearly present in SVs. So, how does anyone be sure that only ACh is released only, or how does anyone rule out effects of choline on postsynaptic cells when cholinergic neurons are stimulated?

Many synapses are now known to release more than one transmitter, making it difficult to define the effect of one transmitter released endogenously.

These are perhaps reasons why some textbooks do not emphasize similarities of endogenously released vs exogenously applied molecules.

For condition 6, the authors made a great effort with Slc6a8. This is a very tough criterion to understand or prove for many synapses and neurotransmitters.

Response: SLC6A8 is a transporter on the cytoplasmic membrane, thus a good candidate for removal of Cr from the synaptic cleft.

In terms of fundamental neuroscience, the story should be impactful. There are certainly more neurotransmitters out there than currently identified and by textbook criteria, creatine seems to be one of them taking all of the data in this study and others into account.

Response: We hope that more will join our lonely efforts in trying to discover more transmitters.

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

Since the authors largely disregarded questions in the review process, I do not see a point in listing recommendation for the authors again.

Reviewer #2 (Recommendations For The Authors):

1. The different sections of the manuscript are not separated by headers.

Response: We do have separate subheadings.

1. The beginning of the results section either does not reference the underlying literature or refers to unpublished data.

Response: We have a very long introduction which was criticized for being too long and with too much historical citations. We therefore refrained from citation again in the beginning part of the Results section.

1. The text contains many opinions and historical information that are not required (e.g., "It has never been easy to discover a new neurotransmitter, especially one in the central nervous system (CNS). We have been searching for new neurotransmitters for 12 years."; l. 17).

Response: We would like to keep these because most readers are young and do not know the history and difficulties of discovering transmitters.

1. Almeida et al. (2008; doi: 10.1002/syn.20280) provided evidence for electrical activity-, and Ca2+-dependent Cr release from rat brain slices. This paper should be introduced in the introduction.

Response: Done.

1. Fig. 7: A Y-scale for the stimulation protocol is missing.

Response: Done.

Reviewer #3 (Recommendations For The Authors):

The main suggestion by this reviewer (beyond the details in the public review) was to consider the full spectrum of biology that is consistent with these results. By my reading, creatine could be a neurotransmitter, but other possibilities also exist. The authors have highlighted some of those for their Discussion.

Version 3

Abstract

Abstract

The discovery of a new neurotransmitter, especially one in the central nervous system (CNS), is both important and difficult. We have been searching for new neurotransmitters for 12 years. We detected creatine (Cr) in synaptic vesicles (SVs), at a level lower than glutamate (Glu) and gamma-aminobutyric acid (GABA) but higher than acetylcholine (ACh) and 5-hydroxytryptamine (5-HT). SV Cr was reduced in mice lacking either arginine:glycine amidinotransferase (AGAT, a Cr synthetase) or SLC6A8, a Cr transporter with mutations among the most common causes of intellectual disability (ID) in men. Calcium-dependent release of Cr was detected after stimulation in brain slices. Cr release was reduced in SLC6A8 and AGAT mutants. Cr inhibited neocortical pyramidal neurons. SLC6A8 was necessary for Cr uptake into synaptosomes. Cr was found by us to be taken up into SVs in an ATP dependent manner. Our biochemical, chemical, genetic and electrophysiological results are consistent with the possibility of Cr as a neurotransmitter, though not yet reaching the level of proof for the now classic transmitters. Our novel approach to discover neurotransmitters is to begin with analysis of contents in SVs before defining their function and physiology.

Introduction

Neural signaling depends on chemical transmission between neurons and their target cells 1–6. Neurotransmission depends on chemicals such as neurotransmitters, neuromodulators and neuropeptides. Decades of work, sometimes with convoluted paths, were involved before a molecule was established as a classic neurotransmitter 7. Initial hints about cholinergic signaling were obtained in the 1800s 8–11. Choline 12,13 and acetylcholine (ACh) 14 were discovered decades before their pharmacological effects were found around the turn of 20th century 15–17. Henry Dale and colleagues realized similarities of ACh and parasympathetic stimulation 5,18 but it was not until 1929 when ACh was detected in the body 19 and 1934 when ACh was proven a neurotransmitter in the peripheral nervous system (PNS) 20–22. It took nearly 100 years from the finding of the effects of supradrenal gland damage 23 or removal 24, the observation of an activity in the supradrenal gland 25, the isolation of an inactive derivative 26–29, and the successful isolation of adrenaline 30–32, the notice of similarities between adrenaline and sympathetic stimulation 3,4,33–35, to the mid-1940s when Ulf von Euler proved that noradrenaline (NA) was the neurotransmitter of the sympathetic nerves 36–38. While it is not easy to establish a molecule as a neurotransmitter in the PNS, it is even harder to establish a central nervous system (CNS) neurotransmitter. Three decades elapsed between the time when ACh was proven to be a PNS neurotransmitter and the time when it was established as a CNS neurotransmitter 39–41 and two decades between NA as a peripheral transmitter and a central transmitter 42,43.

If a neurotransmitter acts only in the CNS, but not in the PNS, it is much more difficult to discover or to prove. Most neurotransmitters were discovered for their effects on peripheral tissues, with muscle contraction or relaxation as a major readout. Glutamate (Glu) 44–47 and gamma-aminobutyric acid (GABA) 46,48–51 were discovered partly because of their peripheral effects and partly because of their effects on spinal neurons. There is no reason for central neurotransmitters to also act peripherally, but relatively little efforts have been reported to find small molecule neurotransmitters acting only on CNS neurons with no peripheral bioassays available. Premature assumptions and technical difficulties are among the major reasons why the hunt for neurotransmitters has not been a highly active area of research over the last three decades.

Are there more neurotransmitters and how can they be discovered? Classic neurotransmitters are stored in synaptic vesicles (SVs) 52–60. They are released upon electric stimulation before being degraded enzymatically or taken up into the presynaptic terminal by cytoplasmic transporters and into SVs by vesicular transporters 61–66. Most of the major textbooks list either 3 67–69 or 4 70–72 criteria of a neurotransmitter: presence in presynaptic neurons, release upon stimulation, action on postsynaptic neurons, mechanism of removal. Some molecules commonly accepted as neurotransmitters still do not meet all the criteria listed in different textbooks, but they nonetheless play important functional roles in the CNS and their defects cause human diseases. Over time, different small molecules have been proposed to function as neurotransmitters (e.g., 46,73), but none satisfies all the criteria. Robust and reliable detection of the candidate molecule in SVs is often, though not always, the problem (cf. 74).

Beginning in 2011, we have been actively searching for new neurotransmitters in the mammalian brain. We have tried different approaches, including searching for neuroactive substances in the cerebral spinal fluid (CSF) and following transporters potentially localized in the SVs. One approach that we have now taken to fruition is the purification of the SVs from mouse brains coupled with chemical analysis of their contents. We have found known transmitters such as Glu, GABA, ACh, and 5-hydroxytryptamine (5-HT). But more importantly, we have reproducibly detected creatine (Cr) in SVs.

Cr was discovered in 1832 by Michel-Eugène Chevreul 75,76 and has long been considered as an energy buffer in the muscle and the brain 77–79. Half of Cr in a mammalian animal is thought to come from diet and the rest from endogenous synthesis 80. Most of the Cr is present in the muscle but it is also present in the brain. Although most of the endogenous Cr is synthesized in the kidney, the pancreas and the liver 77,81, Cr is also synthesized in the brain 80,82,83.

Solute carriers (SLC) contribute to both cytoplasmic and vesicular transporters. With 19 members in humans, family 6 (SLC6) are secondary active transporters relying on electrochemical Na+ or H+ gradients 84–89. SLC6 is also known as the neurotransmitter transporter (NTT) family because some members transport neurotransmitters such as GABA (by SLC6A1 or GABA transporter 1, GAT1; SLC6A13 or GAT2; SLC6A11 or GAT3) 90–93, NA (by SLC6A2 or NA transporter,

NET) 65, dopamine (by SLC6A3 or DAT) 94–96, 5-HT (by SLC6A4 or serotonin transporter, SERT) 66,97, and glycine (by SLC6A9, or GlyT1; SLC6A5 or GlyT2) 90,98,99. Cr is transported by SLC6A8 (also known as CrT, CT1 or CRCT) 100–106. In addition to peripheral organs and tissues, SLC6A8 is also expressed in the nervous system where it is mainly in neurons 82,104,107–109. SLC6A8 protein could be found on the plasma membrane of neurons 109–111.

The functional significance of SLC6A8 in the brain is supported by symptoms of humans defective in SLC6A8. Mutations in SLC6A8 were found in human patients with intellectual disability (ID), delayed language development, epileptic seizures and autistic-like behaviors 112,113. They are collectedly known as Cr transporter deficiency (CTD), with ID as the hallmark. Particular vulnerability of language development has been observed in some SLC6A8 mutations which had mild ID but severe language delay 114. CTD contributes to approximately 1 to 2.1% of X-linked mental retardation 115–123. While CTD is highly prevalent in ID males, it is also present in females, with an estimated carrier frequency of 0.024% 124.

SLC6A8 knockout mice 125–127 showed typical symptoms of human CTD patients with early and progressive impairment in learning and memory. Mice with brain- and neuronal-specific knockout of SLC6A8 showed deficits in learning and memory without changes in locomotion caused by peripheral involvement of SLC6A8 128,129. Deletion of SLC6A8 from dopaminergic neurons in the brain caused hyperactivity 130. These results demonstrate that SLC6A8 is functionally important in neurons.

Cr deficiency syndromes (CDS) are inborn errors of Cr metabolism, which can result from defects in one of the three genes: guanidinoacetate methyltransferase (GAMT) 131, arginine-glycine amidinotransferase (AGAT) 132,133, and SLC6A8 112. That they all show brain disorders indicates the functional importance of Cr in the brain134–137. Cr uptake was observed previously 220, 221.

Here we first biochemically purified SVs from the mouse brain and discovered the presence of Cr, as well as classic neurotransmitters Glu and GABA, ACh and 5-HT, in SVs. We then detected calcium (Ca2+) dependent releases of Cr, Glu and GABA but not ACh and 5-HT when neurons were depolarized by increased extracellular concentrations of potassium (K+). Both the level of Cr in SVs and that of Cr released upon stimulation were decreased significantly when either the gene for SLC6A8 or the gene for AGAT were eliminated genetically. When Cr was applied to slices from the neocortex, the activities of pyramidal neurons were inhibited. Furthermore, we confirmed that Cr was taken up by synaptosomes and found that Cr uptake was significantly reduced when the SLC6A8 gene was deleted. Finally, we found that Cr was transported into SVs. Thus, multidisciplinary studies with biochemistry, genetics and electrophysiology have suggested that Cr is a new neurotransmitter, though the discovery of a receptor for Cr would prove it.

Results

Detection of Cr in SVs from the mouse brain

To search for new neurotransmitters, we tried several approaches. For example, we used Ca2+ imaging to detect neuroactive substances in the cerebrospinal fluid (CSF), but it was difficult to rule out existing neurotransmitters and select responses from potentially new neurotransmitters. We also transfected cDNAs for all human SLCs into dissociated cultures of primary neurons from the mouse brain and found that more than 50 out of all SLCs could be localized in SVs. However, when we used CRISPR-CAS9 to tag some of the candidate SLCs in mice, some of them were found to be expressed outside the CNS, indicating that, while ectopic expression of these candidate SLCs could be localized on SVs, the endogenous counterparts were not localized on SVs.

Here we report our approach using the purification of SVs as the first step (Supplementary Figure 1A) 138–145. Synaptophysin (Syp) is a specific marker for SVs 146–149 and an anti-Syp antibody was used to immunoisolate SVs 143,144,150–152. Visualization by electronic microscopy (EM) (Supplementary Figure 1B, left panel) showed that the purified vesicles were homogenous, with an average diameter of 40.44 ± 0.26 nm (n=596 particles) (Supplementary Figure 1B, right panel), consistent with previous descriptions of SVs 153,154.

Immunoblot analysis with twenty markers of subcellular organelles of neurons and one marker for glia (Supplementary Figure 1C) indicate that our purifications were highly effective, with SV markers detected after purification with the anti-Syp antibody, but not that with the control immunoglobulin G (IgG). SV proteins included Syp 146–149, synaptotagmin (Syt1) 155, synatobrevin2 (Syb2) 156,157, SV2A 158, H+-ATPase 159, and vesicular neurotransmitter transporters for glutamate (VGLUT1, VGLUT2) 160 and GABA (VGAT) 161. Immunoisolation by the anti-Syp antibody did not bring down markers for the synaptic membrane (with SNAP23 as a marker) 162–164, postsynaptic components (with PSD95 and GluN1 as markers) 165,166, the Golgi apparatus (with GM130 and Golgin 97 as markers) 167,168, early endosome (with early endosome-associated 1, EEA1, as a marker) 169, the lysosome (with LC3B and cathepsinB as markers), the cytoplasma (with glyceraldehyde-3-phophate dehydrogenase or GAPDH as a marker), mitochondria (with voltage-dependent anion channel or VDAC, as a marker), cytoplasmic membrane (with calcium voltage-gated channel subunit alpha 1 or CACNA1A as a marker), axonal membrane (with glucose transporter type 4 or GluT4 as a marker) and glia membrane (with myelin basic protein or MBP as a marker). These results indicated that the SVs we obtained were of high integrity and purity.

To detect and quantify small molecules as candidate transmitters present in the purified SVs, capillary electrophoresis-mass spectrometry (CE-MS) was optimized and utilized (Figure 1A, Supplementary Figure 1A) 151,170. We found that the levels of classical neurotransmitters such as Glu, GABA, ACh and 5-HT were significantly higher in SVs pulled down by the anti-Syp antibody than those in lysates pulled down by the control IgG (Figure 1 A-E). Consistent with previous reports 144,151, significant enrichment of neurotransmitters was observed only from SVs immunoisolated at near 0 °C, but not at the room temperature (RT) (Figure1 A-E). By contrast, another small molecule, alanine (Figure 1G), was not elevated in SVs as compared to the control.

Presence of Cr in SVs from the mouse brain.

(A) Representative raw traces from CE-MS of indicated molecules from samples immunoisolated by the control IgG (blue) at 0-2 °C, the monoclonal anti-Syp antibody at 0-2 °C (red) and the anti-Syp antibody at room temperature (RT, green). Q1/Q3 for identifying targets were indicated. (B-G) Quantification of the amounts of indicated molecules. The amount of a molecule was divided by the amount of the anti-Syp antibody bound to magnetic beads. Note Glu (B), GABA(C), ACh (D), 5-HT (E), Cr (F), but not alanine (G) was higher in SVs pulled down by the anti-Syp antibody at 0-2 °C than those pulled down by the IgG control or those pulled down at the RT. n=10 (B-E, and G) or 14 (F) samples per group, ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

The amount of Glu was 171.1 ± 5.4 pmol/μg anti-Syp antibody (n=10, Figure 1B), approximately 10 times that of GABA (n=10,17.81 ± 1.47 pmol/μg anti-Syp antibody, Figure 1C). The amount of ACh was 1.29 ± 0.10 pmol/μg anti-Syp antibody (n=10, Figure 1D), approximately 0.072 that of GABA. The amount of 5-HT was 0.096 ± 0.017 pmol/μg anti-Syp antibody (n=10, Figure 1E). Thus, our purification and detection methods were highly reliable and sensitive enough to detect established neurotransmitters.

Under the same conditions, we also detected Cr in SVs (n=14, Figure 1A and 1F). Amount of Cr in the SVs was found to be 3.43 ± 0.40 pmol/μg anti-Syp antibody (Figure 1F), which was approximately 2% of Glu, 19% of GABA, 266% of ACh and 3573% of 5-HT. It is unlikely that these could be attributable to different levels for different neurotransmitters in each SV, but more likely attributable to the relative abundance of SVs containing different neurotransmitters. 85% to 90% neurons in the mouse brain were glutamatergic while 10%-15% were GABAergic 171–173, which can explain our detection of Glu as approximately 10 times that of GABA (Figure 1 B and C). Similarly, cholinergic neurons (5.67−105) 174 represented 0.81% of total number of neurons (approximately 70 million) in the mouse brain 175, serotonergic neurons (approximately 26,000) for 0.037% of total neurons 175,176. Assuming that the content of each neurotransmitter in a single SV is similar, extrapolation from the above data would suggest that approximately 1.3% to 2.15% of neurons in the mouse brain are creatinergic.

To distinguish whether small molecules co-purified with SVs were in the SVs 143,144, or that they were just associated with the outside of SVs154, we tested the dependence of the presence of these molecules in the SVs on temperature and on the electrochemical gradient of H+. Cr was significantly reduced in SVs purified at the room temperature as compared to that immunoisolated at near 0 °C (Figure 1F), supporting the presence of Cr inside, instead of outside, SVs.

Classical neurotransmitters are stored in SVs with an acidic environment inside (pH of 5.6 ∼ 6.4) 177–179. To further verify the storage of Cr in SVs and examine the role of H+ electrochemical gradient, we applied pharmacological inhibitors during purification 74,180. The proton ionophore FCCP (carbonyl cyanide-4-(tri-fluoromethoxy) phenylhydrazone) was used to dissipate H+ electrochemical gradient 180, 181. FCCP significantly reduced the amount of Cr as well as classical neurotransmitters in SVs (Supplementary Figure 2A-E). The extent of FCCP induced reduction was correlated with the value of pKa or PI (isoelectric point) for different molecule: 5-HT (with pKa predicted to 10 and 9.31, Supplementary Figure 2E) > Cr (PI of ∼7.94, Supplementary Figure 2A) > GABA (PI of 7.33, Supplementary Figure 2C) > Glu (PI of 3.22, Supplementary Figure 2B). Nigericin, a K+/H+ exchanger which dissipates ΔpH180,181, also reduced the amount of Cr and classical neurotransmitters in SVs (Supplementary Figure 2A-E). Furthermore, in the presence of FCCP or nigerin, SV Cr was reduced to a level comparable to that pulled down by control IgG (Supplementary Figure 2A), demonstrating the storage of Cr in SVs was dependent on H+ gradient. As a control, the non-neurotransmitter molecule alanine in SVs was not changed by the inhibitors (Supplementary Figure 2F).

Reduction of SV Cr in mouse mutants lacking slc6a8

SLC6A8, located on the X chromosome, encodes a transporter for Cr and its loss of function (LOF) mutations caused behavioral deficits in humans 112,113 and mice 125–130. To investigate whether SLC6A8 affects Cr in SVs, we generated slc6a8 knockout (KO) mice. Exon 1 of the SLC6A8 gene was partially replaced with CreERT2-WPRE-polyA by CRISPR/Cas9 (Figure 2A). Examination by reverse polymerase chain reaction (RT-PCR) (Supplementary Figure 3A and 3B) and quantitative real-time reverse PCR (qPCR, Supplementary Figure 3C-D) showed that SLC6A8 mRNA was not detected in either male or female mutants, and significantly reduced in female heterozygous (Slc6a8+/-). Consistent with previous reports, the body weights of SLC6A8 KO mice were reduced (Supplementary Figure 4 B and D) 125,182,183. Brain weight was not significantly different between SLC6A8 KO mice and WT mice (Supplementary Figure 4A and C).

SLC6A8 and Cr in SVs.

(A) A schematic illustration of the strategy for generating SLC6A8 knockouts using CRISPR/Cas9. An iCreERT2-WPRE-pA cassette (∼3.5 kb) was inserted immediately downstream of ATG in the SLC6A8 gene, substituting bp 4 to bp 51 in exon 1 (E1). (B) Representative raw traces of Creatine immunoisolated by control IgG from WT mice (blue), the anti-Syp antibody from WT mice (red), IgG from SLC6A8 KO mice (blue) and the anti-Syp antibody from SLC6A8 KO mice (red). (C-H) Quantification of indicated molecules. Note the selective reduction of Cr in SVs from SLC6A8 KO mice. n=14 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

When we examined the contents of SVs isolated by the anti-Syp antibody vs the control IgG, significant reduction was only observed for Cr, but not classical neurotransmitters (Figure 2 B-H, Supplementary Figure 5 A-E). While Cr pulled down by IgG was not significantly different between SLC6A8-/Y and SLC6A8+/Y mice, SV Cr purified by the anti-Syp antibody from SLC6A8-/Y was reduced to approximately 1/3 that of the WT (SLC6A8+/Y) littermates (n =14, Figure 2 B-C). Compared to the IgG control, Cr in SVs was enriched in WT mice, but not in SLC6A8-/Y mice (Figure 2 B and C). In both SLC6A8-/Y and SLC6A8+/Y mice, classical neurotransmitters in SVs were all enriched as compared to IgG controls (Figure 2 C-G, Supplementary Figure 5 A-D). The amounts of Glu (Figure 2D and Supplementary Figure 5A), GABA (Figure 2E and Supplementary Figure 5B), ACh (Figure 2F and Supplementary Figure 5C) and 5-HT (Figure 2G and Supplementary Figure 5D) in SVs were not different between SLC6A8-/Y and SLC6A8+/Y mice. Molecules not enriched in SVs from WT mice, such as alanine, was also unaffected by SLC6A8 KO (Figure 2H and Supplementary Figure 5E).

It is unlikely that the specific reduction of Cr in SVs from SLC6A8 KO mice was due to technical artifacts. First, the possibility of less SVs obtained from SLC6A8 KO mice were precluded by immunoblot analysis, as assessed by SV markers Syp, Syt and H+-ATPase (Supplementary Figure 6A). Second, data collected by high resolution MS (Q Exactive HF-X, Thermoscientific) also revealed selective decrease of SV Cr (m/z=132.0576) from SLC6A8 KO mice (n=8, Supplementary Figure 6 B-G), as quantified by the peak area. Peak areas for Glu (n=8, Supplementary Figure 6B, m/z=148.0604), GABA (n=8, Supplementary Figure 6C, m/z=104.0712), ACh (n=8, Supplementary Figure 6D, m/z = 146.1178) and alanine (n=8, Ala, Supplementary Figure 6E, m/z=90.055) were not significantly different between SVs immmunoisolated with the anti-Syp antibody and control IgG from WT and Slc6a8 KO mice. However, peak areas (Supplementary Figure 5G) and amplitude of Cr (n=8, Supplementary Figure 5F) signal were significantly increased in SVs from WT mice (anti-Syp antibody vs IgG), but not that from SLC6A8 KO mice.

Reduction of SV Cr in mouse mutants lacking AGAT

AGAT is the enzyme catalyzing the first step in Cr synthesis 82,184 and its absence also led to Cr deficiency in the human brain and mental retardation 132,133. To investigate requirement of AGAT for SV Cr, we utilized AGAT ‘knockout-first’ mice (Figure 6A) 185. The targeting cassette containing Frt (Flip recombination sites)-flanked EnS2A, an IRES::lacZ trapping cassette and a floxed neo cassette were inserted downstream of exon 2 to interfere with normal splicing of AGAT pre-mRNA. Examination by RT-PCR (Supplementary Figure 7A) and quantitative RT-PCR (Supplementary Figure 7B) showed reduction of AGAT mRNA in AGAT+/- mice and absence of AGAT mRNA in AGAT-/- mice. Body weight (Supplementary Figure 8B), but not brain weight (Supplementary Figure 8A), of AGAT-/- mice were lower than both AGAT+/+ and AGAT+/- mice, which were similar to SLC6A8 KO mice.

Immunoblot analysis showed SVs purified from the brains were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice (Supplementary Figure 9 A-D), as supported by quantitative analysis of Syp, Syt and H+-ATPase (n=20, with 2 repeats for 10 samples).

We analyzed small molecules present in SVs from AGAT+/+, AGAT+/- and AGAT-/- mice. Cr was significantly enriched in SVs from all three genotypes as compared to the IgG control (Figure 3A and 3B). However, the level of Cr from AGAT-/- mice was significantly lower than those from AGAT+/+ and AGAT+/- mice (n=10, Figure 3A and B).

Reduction of SV Cr in AGAT knockout mice.

(A) Representative raw traces of Creatine, pulled down by IgG or anti-Syp from AGAT+/+, AGAT+/- and AGAT-/- mice. (B-G) Quantification of indicated molecules. Cr was significantly decreased in AGAT-/- mice as compared to Cr in AGAT+/+ or AGAT+/- mice (B). GABA was significantly decreased in SVs from AGAT-/- mice as compared to AGAT+/+ mice (D) but the difference was smaller than that of Cr. Glu (C), ACh (E), 5-HT (F) and alanine was not different among AGAT+/+, AGAT+/- and AGAT-/- mice. n=10 samples per group. ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction.

Glu (Figure 3C), ACh (Figure 3E) and 5-HT (Figure 3F) were all enriched in SVs (as compared to IgG controls), and not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice. GABA in SVs from AGAT-/- mice was also decreased from AGAT+/+ mice by 30%, to an extent less than that of Cr (78.4%). Alanine was not different among three genotypes of mice (n=6, Figure 3G). Thus, Cr and GABA, but not other neurotransmitters, in SVs were reduced in AGAT KO mice.

Pattern of SLC6A8 expression indicated by knockin mice

We generated SLC6A8-HA knockin mice by CRISPR/Cas9. Three repeats of the hemagglutinin (HA) tag 186, the T2A sequence 187–189 and CreERT2 190–193 were inserted in-frame at the C terminus of the SLC6A8 protein (Figure 4A).

Expression pattern of SLC6a8.

(A) A diagram illustrating the generation of SLC6A8-HA KI mice. 3*HA, T2A and CreERT2 were inserted in-frame, before the stop codon, to the C terminus of SLC6A8 protein. (B-K) Low magnification photomicrographs of coronal sections immunohistochemically labelled with the anti-HA antibody in SLC6A8-HA mice. (I and M) immunostaining with the anti-HA antibody in control WT mice. Abbreviations: Pir, piriform cortex; M, motor cortex; LS: lateral septum; Hp: hippocampus; Hb, habenular nucleus; Vp: ventral posterior nucleus of thalamus; Au auditory cortex; Amyg: amygdala; Ip, interpeduncular nucleus; Pn; pontine nucleus; Cb, cerebellum; Pr, prepositus; SC, Superior colliculus; MVe, medial vestibular nucleus. Scale bar: 500 μm.

To examine the expression pattern of SLC6A8, we performed immunocytochemistry with an antibody against the HA epitope in SLC6A8-HA and WT mice. SLC6A8-HA mice showed positive signals in the olfactory bulb (Figure 4B), the piriform cortex (Figure 4 C-F), the somatosensory cortex (Figure 4 F and G), the ventral posterior thalamus (Figure 4H), the interpeduncular nucleus (Figure 4I) and the pontine nuclei (Figure 4J). In addition, moderate levels of immunoreactivity were observed in the motor cortex (Figure 4 D-H), the medial habenular nucleus (Figure 4H), the hippocampus (Figure 4H) and the cerebellum (Figure 4K). These results were consistent with previous reports 109,111. WT mice were negative for anti-HA antibody staining (Figure 4 L and M).

Ca2+ dependent release of Cr upon stimulation

Classical neurotransmitters are released from the SVs into the synaptic cleft in a Ca2+ dependent manner after stimulation. For example, high extracellular potassium (K+) stimulated Ca2+ dependent release of Glu, GABA and other neurotransmitters in brain slices 194–198.

300 μm thick coronal slices of the mouse brain within 1∼2 millimeters (mm) posterior to the bregma were used, because the cortex, the thalamus, the habenular nucleus and the hippocampus were positive for SLC6A8 (cf., Figure 4H). We monitored the effect of K+ stimulation by recording neurons in the slices. Immediately after K+ stimulation, pyramidal neurons in the CA1 region of the hippocampus were depolarized, firing a train of action potentials and reaching a large depolarization plateau in less than 1 minute (min) (Figure 5A). K+-induced depolarization persisted for several mins before returning to the baseline and being washed within 10 mins. Thus, superfusates in 1-min fraction at the time points of 1.5 min before (control) and after K+ stimulation, and 10 mins after the wash were collected (Figure 5A), and the metabolites in the superfusates were analyzed by CE-MS.

Cr release in brain slices from WT and SLC6a8 knockout mice.

(A) Neuronal depolarization induced by 70 mM K+ and time points for collecting the release sample. “Control” samples were collected 1.5 min to 0.5 min before K+ stimulation, “70 mM K+” ACSF samples were collected during 70 mM K+ stimulation, and “wash” samples were collected 10 mins after washout with ACSF. Efflux of Glu (B), GABA (C) or Cr (D) from WT or SLC6A8 KO male mice (n=7 samples per group). Note that a small amount of Cr released in SLC6A8 KO mice did not return to the baseline after 10 mins of washing. (E-G) Ca2+ dependent release of Glu, GABA and Cr in WT and SLC6A8 KO mice (n=5 samples per group). ***, p<0.001, ns, not significant. one-way ANOVA with Tukey’s correction. #, p<0.05, ##, p<0.01, paired t test.

Cr release in WT and AGAT knockout mice.

(A) A schematic diagram illustrating the strategy of AGAT knockout first. With the AGAT gene (also known as GATM) shown in the upper part, and the gene targeting strategy in the lower part. The homologous arm is approximately 10 kb. A targeting cassette, containing Frt-flanked lacZ and neomycin, was inserted downstream of exon 2. At the same time, exon 3 of AGAT was flanked by two loxP sites. K+ induced release of Glu (C) and GABA (D) were not significantly different among AGAT+/+, AGAT+/- and AGAT-/- mice, whereas that of Cr (B) was significantly lower in AGAT-/- mice than those in AGAT+/+ and AGAT+/- mice.

In the presence of Ca2+, depolarization with elevated extracellular K+ led to robust release of Glu and GABA in slices from WT (SLC6A8+/Y) mice (n=7 per group, Figure 5 B and C). After 10 mins wash, levels of Glu and GABA returned to the baseline (Figure 5 B and C). In the presence of Ca2+, depolarization with elevated K+ led to robust release of Cr. Extracellular Cr returned to the baseline after 10 min wash (Figure 5D). For quantification, the stimulated releases of metabolites were calculated by subtracting the basal levels from the total releases in response to K+ stimulation. In the presence of Ca2+, K+ stimulation induced the efflux of Glu, GABA and Cr at 0.46, 0.33 and 0.086 nmol/min, respectively (n=7 per group) (Figure 5 B, C, D). From the detection limits of ACh and 5-HT in our system, we inferred that the efflux rate for ACh was lower than 0.001 nmol/min and that for 5-HT lower than 0.003 nmol/min. The efflux rate for Cr in brain slices is lower than those of Glu and GABA, but higher than those for ACh and 5-HT.

Ca2+ dependence of transmitter release was examined by comparing responses to ACSF without Ca2+ or elevated K+ (supplemented with 1 mM EGTA), elevated extracellular K+ in the absence of Ca2+ (supplemented with 1 mM EGTA), or K+ in presence of 2.5 mM Ca2+ (Figure 5 E-G, n=5 per group). In the absence of Ca2+, elevated K+ stimulated the release of a small but significant amount of Glu and GABA, with efflux rates at 0.056 nmol/min and 0.066 respectively (Figure 5 E, F). In the presence of 2.5 mM Ca2+, elevated K+ further augmented the release of Glu and GABA by 5 to 6 times, confirming previously reported Ca2+ dependent release of neurotransmitters in response to depolarization 194,196.

Cr was also released both in a Ca2+ dependent and a Ca2+ independent manner (Figure 5G). More Cr was released in response to K+ stimulation in the presence of 2.5 mM Ca2+ than that in the absence of Ca2+. These results demonstrate Ca2+ dependent release of Cr upon stimulation.

Reduced Cr release in SLC6A8 and AGAT mutant mice

We examined whether SLC6A8 KO affected K+-induced release of Cr. While Glu and GABA were released in slices from SLC6A8 KO (SLC6A8-/Y) mice at levels not significantly different from those of WT mice (Figure 5 B, C), release of Cr in response to K+ stimulation was significantly reduced in SLC6A8-/Y mice as compared to SLC6A8+/Y mice (Figure 5D). The basal level of Cr in SLC6A8 KO mice was lower than that of WT mice. In addition, K+ stimulation induced release of Cr persisted to some extent even after 10 mins of washout (Figure 5D), possibly due to the inability of presynaptic terminals in SLC6A8 KO mice to reuptake Cr in the synaptic cleft (Figure 8).

Experiments with slices from brains of SLC6A8 KO (SLC6A8-/Y) mice showed that Ca2+ dependent release of either Glu or GABA was not affected by the genotype of SLC6A8 (Figure 5 E and F). By contrast, Ca2+ dependent release of Cr was abolished in SLC6A8-/Y slices. Interestingly, Ca2+ independent release of Cr was reduced by a third, but did not reach statistical significance, in SLC6A8-/Y slices. In the absence of Ca2+, the basal level of Cr was not changed in SLC6A8 KO mice. Taken together, these results indicate that there is Ca2+ dependent release of Cr upon stimulation and that SLC6A8 is required specifically for Ca2+ dependent release of Cr, but not for Ca2+ dependent release of other neurotransmitters such as Glu and GABA, or for Ca2+ independent release of Cr. Knockout of AGAT (Figure 6A) selectively reduced K+ evoked release of Cr, but not those of Glu or GABA (n=5 per group, Figure 6 B, C, D). Although K+ stimulation still elicited Cr release from brain slices of AGAT+/- mice, the efflux rate in AGAT-/- mice was reduced to less than 10% that in AGAT+/+ mice and 20% that in AGAT+/- mice.

Cr inhibition of neocortical neurons

Our own data (Figure 4H) and previous reports 109,111 have shown SLC6A8 in the neocortex, with dense SLA6A8-HA immunoreactive fibers in layer 4 (Supplementary Figure 10A) 199,200. Layer 5 neurons in the somatosensory cortex have been reported to express SLC6A8 previously 109,111. To investigate electrophysiological effects of Cr, we performed whole cell patch clamp recordings from the pyramidal neurons in layer 4/5 of the somatosensory cortex (Figure 7 A and B, Supplementary Figure 10).

Inhibitory effects of Cr on cortical neurons.

(A) A photograph showing recording at layer 4 in the somatosensory cortex. Scale bar: 10 μm. (B) Patch-clamp recording of a pyramidal neuron. Scale bar: 10 μm. (C) Ratios of Cr-responsive and -unresponsive neurons in the region. (D) Representative raw electrophysiological traces showing inhibition of evoked firing by Cr, with the lower panel showing the stimulus protocol. (E) Evoked spike numbers in response to different current injections from (D). (F) Relationship between evoked spike numbers and different current injections to neurons that were inhibited by Cr (n=16). (G) The same for Cr-unresponsive neurons (n=35). (H) Rheobase for Cr responsive neurons. (I) Evoked spike number when these neurons were injected with current of rheobase + 50 pA. (J) input resistance. *, p<0.05; **, p<0.01; ***, p<0.001, paired t test.

Cr uptake into synaptosomes and SVs.

(A) Markers for subcellular organelles detected in synaptosomes prepared from WT mice or mice with Slc6a8 gene fused in frame to the HA epitope. SLC6A8-HA was enriched in synaptosomes (Sy or 4-Sy) or crude synaptosomes (P2) (see methods), as were other markers for the subcellular organelles in synaptosomes but not the marker for myelin (MBP). (B) Representative electron micrographs and histograms of size distribution in synaptosomes isolated from WT (Slc6a8+/Y) and Slc6a8 KO (Slc6a8-/Y) mice by Ficoll density-gradient centrifugation. Sy, synaptosome; SV, synaptic vesicles; Mt, mitochondria; PSD, post-synaptic density. Bar, 20 nm. (C) Cr uptake into synaptosomes (n=5 per group). The two left columns were results from WT mice and the two right columns from Slc6a8 knockout mice. The control baseline was [14C]-Cr uptake at 0 °C at 10 mins (205). Cr uptake into synaptosomes at 37 °C measured at 10 mins was observed in WT synaptosomes. Uptake into Slc6a8 knockout synaptosomes was significantly reduced when compared to the WT synaptosomes. ***, p<0.001, one-way ANOVA with Tukey’s correction. (D) uptake of [13C]-Cr into immunoisolated SVs in the presence or absence of ATP. (n=11 samples per group)*, p<0.05, paired-t test.

Medium-sized pyramidal neurons (Figure 7B) with a membrane capacitance (Cm) of 114.96 ± 3.92 pF (n=51, Supplementary Figure 11A) were recorded. These neurons exhibited regular firing patterns 201,202 in response to depolarization current injection (Figure 7D and Supplementary Figure 11A) with moderate maximal evoked spiking frequencies of 10-30 spikes per 500 milliseconds (ms) (Figure 7 E-G), increasing of inter-spike intervals during depolarizing steps (Figure 7D and Supplementary Figure 11A), high action potential amplitude (81.64 ± 1.06 mV, Supplementary Figure 12E) and large spike half-width (1.12 ± 0.031 ms, Supplementary Figure 12F).

Cr was bath applied only after the evoked firing pattern reached a steady state. Of the 51 neurons, 16 were inhibited by 100 μM creatine (Figure 7 B-F). Fewer spikes were evoked in Cr responsive neurons in response to depolarizing current injections during Cr application (25 pA step, 500 ms) (Figure 7 D-F). The inhibitory effect of Cr was reversible (Figure 7 D-F), typically observed within 2-3 min following Cr application (with maximal effect from 2 to 8 mins) and disappeared after 10 to 25 min washout. This could be repeated by a second application of Cr. The rheobase, defined as the minimal electrical current necessary to elicit an action potential, was increased during bath application of Cr (Figure 7 H). The inhibitory effect was most obvious at near spike threshold. When a neuron was depolarized with a current of 50 pA above rheobase, the number of evoked spikes was decreased dramatically during Cr application (Figure 7I). Cr also mildly inhibited the input resistance (Figure 7J), slightly hyperpolarized resting membrane potential (Supplementary Figure 12C) or reduced amplitude of afterhyperploarization (AHP) followed by the first evoked action potential (Supplementary Figure 12G). The spike threshold (Supplementary Figure 12D), amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F) were not changed by Cr.

The remaining 35 neurons were not responsive to Cr (Figure 7C and G, Supplementary Figure 11). Cr did not change electrical parameters tested, including evoked firing rates (Figure 7G, Supplementary Figure 11 A, B and D), rheobase (Supplementary Figure 11C), resting membrane potential (Supplementary Figure 12C), spike threshold amplitude (Supplementary Figure 12E) and half width (Supplementary Figure 12F). In addition, electrical properties of responsive neurons and unresponsive neurons were not significantly different. With the limited number of neurons recorded, the ratio of responsive neurons appeared higher in layer 4 or border of layer4/5, than the deeper layer in layer 5 (Supplementary Figure 10).

SLC6A8 dependent uptake of Cr into the synaptosomes

Along with enzymatic degradation, re-uptake by transporters serves as an important way to remove neurotransmitters released into the synaptic cleft. As synaptosomes contain the apparatus for neurotransmission, they are often used for studying uptake of neurotransmitters 203.

To investigate whether Cr uptake into synaptosomes required SLC6A8, we first examined whether SLC6A8 was present in synaptosomes. Using Slc6a8-HA knockin mice and an anti-HA antibody, we found that SLC6A8-HA was present and enriched in crude synaptosomal fraction (P2 fraction in Figure 8A, enrichment score: P2/H = 1.76 ± 0.15, n = 4) and synaptosomal fraction prepared using a discontinuous Ficoll gradient (Sy and 4-Sy fractions in Figure 8A, enrichment score: Sy/H = 2.02 ±0.14, n = 4). The integrity of synaptosomes was confirmed by multiple markers of the synaptosomes 203, including the presynaptic membrane marker SNAP25 204 and the SV marker Syp (synaptophysin Syp) 146–149 in presynaptic terminals, the postsynaptic density marker PSD95 165,166 and the postsynaptic membrane protein GluN1 205, the synaptic membrane protein SNAP23 162–164, the plasma membrane marker Na+-K+-ATPase 206–208, and the mitochondria marker VDAC 203. These were all enriched in our synaptosomal preparations. The cytosol marker GAPDH was also present in synaptosomes, whereas the oligodendrocyte marker MBP was nearly absent, suggesting myelin pollution was largely avoided 203.

We have also used EM to confirm the quality of our synaptosome preparations. As reported previously 203,209, synaptosomes were composed of membrane bounded structures (Sy in Figure 8B) filled with synaptic vesicles (SV in Figure 8B), sometimes with a segment of postsynaptic membrane along with the postsynaptic density (PSD in Figure 8B) and mitochondria (Mt in Figure 8B). The sizes of synaptosomes from WT mice and Slc6a8 knockout mice were similar, with areas of 0.245 ± 0.01 μm2 (n=302 particles) and 0.247 ± 0.01 μm2 (n=317 particles), respectively (Figure 8B).

We then examined whether SLC6a8 participated in Cr uptake into the synaptosomes. A mixture of 18 μM [14C]-Cr (with a total radioactivity of 0.4 μCi) and 5 μM Cr was used, and uptake at 0 °C measured at 10 mins was the baseline 210. Cr uptake into synaptosomes from WT mice was stimulated approximately sevenfold at 37 °C (Uptake, Figure 8C) as compared with 0 °C (Ctrl, Figure 8C). Cr uptake into synaptosomes from SLC6a8 knockout mice was less than 3 times as compared to its control, and was decreased to approximately 1/3 of that of WT mice (Figure 8C). Thus, SLC6A8 is necessary for uptake of Cr into the synaptosomes.

Cr uptake into SVs

Classical neurotransmitters were taken up in SVs in an ATP-dependent manner144,159,181,210,211. We examined whether Cr could be transported into SVs.

We used 10 μg anti-Syp antibody to purify SVs from mouse brains. Purified SVs were preincubated for 30 min to allow sufficient leakage of endogenous Cr, before being mixed with 1 mM [13C]-Cr in the presence or absence of 4 mM ATP and placed at 25 L for 10 mins to allow adequate uptake. The SV content of [13C]-Cr was then examined by CE-MS and high-performance liquid chromatography-mass spectrometry (HPLC-MS). Significantly more [13C]-Cr were taken up by SVs in the presence of ATP, with about 10.3 pmol [13C]-Cr transported into SVs (1.03 pmol/μg α-Syp or transportation rate of 0.103 pmol/min, n=11, Figure 8D).

In summary, Cr could be transported into SVs in ATP-dependent manner. At this point, we do not know what is the transporter(s) on the SVs for Cr uptake. SLC6A8 is only found in plasma membrane, not on SVs, and is not a candidate for Cr uptake into SVs.

Discussion

While no neurotransmitter has been proven in a single paper, supportive evidence suggesting Cr as a possible new neurotransmitter has been presented here to the extent of any single previous papers.

At various times and by different researchers, taurine 46,212, proline 73, D-aspartic acid 213, hydrogen sulfide 214, agmatine 215, DOPA 216, estradiol 217, β-alanine 218, protons 219 have been suspected as neurotransmitters, but they do not meet all the criteria. Some of the suspected molecules can be released upon stimulation or removed by transporters. Often, they have not been reproducibly found in SVs 74.

Our discovery of Cr in SVs significantly raised the priority of testing the candidacy of Cr, and our further investigations have led to more evidence suggesting Cr as a neurotransmitter: 1) Cr is stored in SVs; 2) Ca++ dependent release of Cr upon stimulation has been observed; 3) both Cr storage in SVs and Cr release are reduced when either the gene for SLC6A8 or the gene for AGAT was deficient; and 4) Cr inhibits activities of pyramidal neurons in the neocortex; 5) Cr uptake into synaptosomes requires SLC6A8; 6) Cr uptake into SVs was ATP-dependent.

Of the above results, 1, 3, 4 and 6 are reported for the first time in this paper. Furthermore, we have demonstrated that detection of Cr in SVs was lower than those for Glu and GABA, but higher than those for ACh and 5-HT, placing Cr at a level in the middle of known central transmitters (Figures 1-3). The storage of Cr in SVs is dependent on preserved H+ gradient (Supplementary Figure 2) and Cr can be transported into SVs (Figure 8D).

There was a single previous report of Ca++ dependent release of [3H]Cr and endogenous Cr in response to electrical stimulation 220. We now provide evidence that Cr was released in response to extracellular K+ stimulation (within 1-2 min) (Figure 5-6). Furthermore, Cr release was reduced when either the Slc6A8 or AGAT gene was removed (Figures 5 and 6). Although the Ca2+ dependent component of K+ evoked Cr release was smaller as compared to those of Glu and GABA, it nevertheless existed and was totally abolished by Slc6A8 knockout (Figure 5). The reported electrically evoked Cr release showed more Ca2+ dependence 220. Taken together, our data and previous report 220 supported a role of Cr as a neurotransmitter. Our observation of extremely low efflux rates of 5-HT or ACh may have arisen from very limited numbers of cholinergic174 or serotoninergic neurons176 in the sliced sections and rapid enzymatic degradation of these neurotransmitters.

Cr uptake from the extracellular space into the cells was reported twice previously, once with brain slices showing sodium dependent uptake of [3H]Cr 220 and once with synaptosomes 221. Our new results have not only replicated the synaptosome Cr uptake experiment but also shown the requirement of SLC6A8, a membrane transporter expressed in synaptosomes (Figure 8A), for Cr uptake into synaptosomes. Transportation of Cr into synaptosomes by Slc6A8 may function for both the clearance of Cr from the synaptic cleft and recycling of Cr into SVs residing in neurons (Figure 8B, Figure 6D).

In summary, in addition to confirming and extending previous results which have stood alone for more than a decade without replication or follow-up, we have obtained entirely new results suggesting the candidacy of Cr as a neurotransmitter. We discuss below the criteria for a neurotransmitter, Cr as a neurotransmitter, and the implications of Cr as a neurotransmitter.

Criteria of a neurotransmitter

The criteria for establishing a non-peptide small molecule as a neurotransmitter have varied from time to time and from author to author.

Some textbooks simply state that a neurotransmitter is stored presynaptically, released upon stimulation and active on postsynaptic neurons. The details of these 3 criteria can vary. For example, one textbook stipulates that “the substance must be present within the presynaptic neuron; the substance must be released in response to presynaptic depolarization, and the release must be Ca2+ dependent; specific receptors for the substance be present on the postsynaptic cell” 67,68. Another states that “the molecule must be synthesized and stored in the presynaptic neuron; the molecule must be released by the presynaptic axon terminal upon stimulation; the molecule, when experimentally applied, must produce a response in the postsynaptic cell that mimics the response produced by the release of neurotransmitter from the presynaptic neuron69. The neuroscience textbook most widely used internationally for the last 4 decades lists 4 criteria for a neurotransmitter 70,71: it is synthesized in the presynaptic neuron; it is present within vesicles and is released in amounts sufficient to exert a defined action on the postsynaptic neuron or effector organ; when administered exogenously in reasonable concentrations it mimics the action of the endogenous transmitter; a specific mechanism usually exists for removing the substance from the synaptic cleft. These are similar, but not identical, to the classic textbook on neurotransmitters: a neurotransmitter “should be synthesized and released presynaptically; it must mimic the action of the endogenous compound that is release on nerve stimulation; and where possible, a pharmacological identity is required where drugs that either potentiate or block postsynaptic responses to the endogenously released agent also act identically to the suspected neurotransmitter that is administered” 222. The pharmacological criterion is listed in another textbook 223.

Some authors note difficulties of establishing a CNS neurotransmitter. For example, a specialized neurotransmitter book states that “the candidate neurotransmitter should be present in the presynaptic terminal, be released when the presynaptic terminal is active, and when applied experimentally, induce faithful responses in the postsynaptic neuron. In practice, since central nervous system neurons continuously integrate diverse excitations and inhibitions, the last criterion is relaxed to demonstrating merely changes in such activity” 224.

Solomon Snyder, a leading scientist of classic neurotransmitters, neuropeptides and their receptors, wrote that “designating a molecule as a transmitter depends on the criteria employed, the most common of which are that the substance is synthesized in neurons, released by their terminals, mimics the effects of physiologic neurotransmission and possess a mechanism for inactivation. However, with each new candidate the rules have been modified and broadened” 225.

Evidence supporting Cr as a neurotransmitter

16 small molecules have been listed as neurotransmitters in the classic textbook 70,71. Among them, adenosine, arachidonic acid, nitric oxide and carbon monoxide do not meet all four criteria, at the present. Cr appears to be better than these in meeting the criteria for a central neurotransmitter.

The results obtained by us in this paper have satisfied the criteria of Robinson 224 for Cr to be a CNS neurotransmitter.

The four criteria of Snyder and colleagues 225 have been mostly met but the physiological neurotransmission would require more research because a specific synapse(s) would have to be defined and studied for putative creatinergic neurotransmission. This can take much longer in the CNS than the PNS. Some commonly accepted neurotransmitters have never satisfied this criterion in a strict sense. The mechanism of Cr removal criterion is met not only by the Cr uptake in brain slices 220 and in synaptosomes 221, but also by our demonstration that SLC6A8 is required for synaptosome uptake of Cr.

The four criteria of Kandel et al.70,71 are mostly satisfied with some details requiring further research. The synthesis requirement is usually not strict because there are transmitters synthesized in some cells and transported into others where they function as transmitters. Our discovery of Cr in SVs can replace the synthesis requirement because the presence in neuronal SVs provide sufficient evidence that Cr is located in the right location to function as a neurotransmitter. The level of Cr in SVs is higher than those of ACh and 5-HT (Figs. 1 and 2). The amount of released Cr is in the same order of magnitude as those of Glu and GABA (Figs. 5 and 6). The criterion of a specific mechanism of removal was met by Cr uptake experiments in slices 220 and in synaptosomes 221, and further strengthened by our finding of SLC6A8 involvement in synaptosome uptake of Cr (Figure 8).

Here we report that Cr, at a concentration comparable to classical neurotransmitters, inhibits pyramidal neurons in specific regions of the mouse brain, with approximate 1/3 of pyramidal neurons responding to 100 μM Cr (Figure 7). In previous reports, 100 μM to 2 mM of GABA226–228, 50 μM to 2 mM of Glu 229,230, 1 to 100 μM of DA229, 0.1-100 μM of 5-HT231–236 were bath applied to investigate the physiological functions of neurotransmitters. Our results revealed that, when bath applied, Cr could inhibit cortical neurons at 100 μM within several minutes, with a time course similar to that of 5-HT237 234 236 and DA 229, but significantly slower than that of Glu230, GABA48,227 and 5-HT233.

In a recent report, knockout of the Slc6a8 gene increased excitation of cortical neurons267. Electrophysiological characterization of pyramidal neurons in the prefrontal cortex (PFC) found increased evoked firing frequency. Because we have shown that Cr inhibit a fraction of pyramidal neurons in the neocortex (Figure 8), this paper provides in vivo evidence consistent with the possibility of Cr as an inhibitory neurotransmitter.

Differences between Cr and classic neurotransmitters

At this point, we do not have a molecularly defined receptor for Cr, only inferring its presence from the electrophysiological responses to Cr. We speculate that Cr may act on G-protein coupled receptors (GPCRs), rather than the fast-acting ligand-gated ion channels, such as AMPA or NMDA receptors for Glu and GABAA receptor for GABA. There have been previous reports of Cr effects on neurons, including Cr as a partial agonist for GABAA receptors 238–241. These effects require very high concentrations of Cr (in the ten millimolar range). There was also a report of the opposite effect: that Cr (at a concentration above 500 μM) increased neuronal excitability through NMDA receptors after incubation for 60 mins, with a time course significantly slower than those of classic neurotransmitters 242.

Ca2+ independent component of Cr release induced by extracellular K+ was more prominent than those of Glu or GABA. One possibility was that Ca2+-independent Cr release came from glia, because high GAMT levels were reported in astrocytes261 and oligodendrites 261, 262. As reported, other neuromodulators such as taurine can be released from astrocytes 263 or slices 264 in an Ca2+ independent manner. In addition, in the absence of potassium stimulation, Ca2+ depletion increased release of taurine in cultured astrocytes 265 or in striatum in vivo266. Similarly, in SLC6A8 KO slices, Ca2+ depletion (Figure 5G) also increased Cr baseline as compared to that in normal ACSF (Figure 5D).

With much longer history of research, ACh and 5-HT now have more evidence in other aspects than Cr as a central transmitter, especially because there are many agonists and antagonists for ACh and 5-HT to prove an additional criterion that is required in some 222,223, but not the majority of, textbooks for a neurotransmitter. The pharmacology criterion will take some time and efforts, because, so far, no efforts have been made to find agonists or antagonists for Cr.

Implications of SLC6A8 and Cr

It is notable that SLC6A8 belongs to the neurotransmitter transporter family, with multiple members already shown to transport neurotransmitters 65,66,90–99.

The uptake experiments by others and us indicate that SLC6A8 transports Cr into neurons within the brain. AGAT is also expressed in the brain, but in cells not expressing SLC6A8 243,244. Cr and its precursor were thought to be transported between different cells in the nervous system. When SLC6A8 was completely missing, such as in homozygous SLC6A8 deficient patients, Cr treatment was not effective. But, if

SLC6A8 was partially active, Cr was effective 245. Intractable epilepsy in a female with heterozygous SLC6A8 mutation was completely treated by Cr 246. Our data of inhibitory effect of creatine on cortical neurons might provide a new mechanism to its anti-epileptic activity 247.

The absence of SLC6A8 expression in astrocytes whose endfeet lining microcapillary endothelial cells (MCEC) form the blood-brain barrier (BBB) indicates that Cr in the brain does not rely on import from the periphery and is instead mainly synthesized in the brain 80,82,248. SLC6A8 functions within the brain to transport Cr and its precursors, not as a major contributor of Cr transport across the BBB. It is thought to mediate Cr uptake into the presynaptic terminal, based on studies of synaptosomes (Figure 8) 221.

Cr is known to have effects other than an energy source, and Cr supplement has been thought to be beneficial for children, pregnant and lactating women, and old people 79,249. Cr has been reported to improve human mental performance 250–255. Cr has been used as potential treatment in animal models of neurodegenerative diseases 256,257. Our work will stimulate further research to distinguish which of the previously suspected effects of Cr is not attributed to its role as an energy storage, but can be attributed to its role as a neurotransmitter.

Search for new neurotransmitters

Our work may stimulate the search for more neurotransmitters. Our discovery indicates that the hunt for neurotransmitters stopped decades ago because of technical difficulties, not due to absence of more neurotransmitters. The fact that most of the known small molecule neurotransmitters have been found because of their peripheral effects also argues that what is missing is the concerted efforts to uncover central neurotransmitters with no peripheral effects. New neurotransmitters may be discovered from candidates which have been long suspected, and from previously unsuspected molecules or even previously unknown molecules.

Innovative approaches should be taken to uncover molecules with no previous suspicions or hints. Highly purified SVs, SVs from different regions of the brain, SVs with specific SLCs offer some of the starting points for future research.

Materials and methods

Generation of knockout and knockin mice.

SLC6A8 knockout and knockin mice were generated using CRISPR-Cas9-mediated genome engineering techniques by Beijing Biocytogen (Beijing, China). AGAT ‘knockout-first’ 185 mice were purchased from CAM-SU GRC (Suzhou, China). All mutations were validated by southern blot analysis, tail junction PCR and DNA sequencing.

Reverse-transcription polymerase chain reaction (RT-PCR) and quantitative real-time PCR (qPCR)

Total RNA of whole brains from mice of different genotypes was extracted using the Buffer RZ (Tiangen, no. RK14, Beijing, China) and reverse transcribed into complementary DNA (cDNA) using the RevertAid First-Strand cDNA synthesis kit (Thermo Scientific, K1622, USA). qPCR was performed using the Taq Pro Universal SYBR qPCR Master Mix (Vazyme, Q712-02) on Bio-Rad CFX-96 Touch Real-time PCR System (Bio-Rad, USA). Glyceraldehyde-3-phosphate dehydrogenase (GAPDH) was used as an internal control. ΔCt (difference in cycle threshold) was calculated for each sample (ΔCt = Ct Target gene − Ct GAPDH) for further evaluation of relative mRNA expression levels in different genotypes. The sequence specificities of the primers were examined. Three pairs of primers targeting different genes were used: SLC6a8 forward, 5’-GTCTGGTGACGAGAAGAAGGG-3’, Slc6a8 reverse, 5’-CCACGCACGACATGATGAAGT −3’; AGAT forward, 5’-CACAGTGGAGGTGAAGGCCAATACATAT-3’, AGAT reverse, 5’-CCGCCTCACGGTCACTCCT-3’; GAPDH forward, 5’-AGGTCGGTGTGAACGGATTTG −3’, GAPDH reverse, 5’-TGTAGACCATGTAGTTGAGGTCA-3’.

Primers for reverse PCR were designed to obtain complete coding sequences based on information obtained from National Center for Biotechnology Information (NCBI): SLC6A8 forward, 5’-ATGGCGAAAAAGAGCGCTGAAAACG-3’; Slc6a8 reverse, 5’-TTACATGACACTCTCCACCACGACGACC-3’; AGAT forward, 5’-ATGCTACGGGTGCGGTGTCT-3’; AGAT reverse, 5’-TCAGTCAAAGTAGGACTGAAGGGTGCCT-3’. PCR products were electrophoresed on 1% agarose gels, stained with GelRed, visualized under UV illumination and photographed.

Immunoblot analysis

Samples were loaded onto 10% polyacrylamide gels with the PAGE system (#1610183, Bio-Rad Laboratories, USA) and run in the SDS running buffer (25 mM Tris, 192 mM glycine, 0.1% SDS, pH 8.8) for 25 min at 80 V followed by 25 - 45 min at 200 V. Afterward, proteins were transferred to immobilon NC transfer membranes (HATF00010, Millipore) at 400 mA for 2 h in transfer buffer (25 mM Tris, 192 mM glycine, 20% methanol). Membranes were blocked in 5% fat-free milk powder in TBST (25 mM Tris, 150 mM NaCl, 0.2%Tween-20 (P1397, Sigma), pH 7.4 adjusted with HCl) and incubated overnight with the indicated primary antibodies dissolved in TBST containing 2% BSA.

Primary antibodies are listed below: rabbit anti-Synaptophysin (dilution 1:5000, cat no. 101002, SySy, Germany), rabbit anti-Synaptotagmin1/2 (dilution 1:2000, cat no.105002, SySy, Germany), rabbit anti-proton ATPase (dilution 1:1000, cat no.109002, SySy, Germany), rabbit anti-Synaptobrevin 2 (dilution 1:5000, cat no.104202, SySy, Germany), rabbit anti-SV2A (dilution 1:2000, cat no.109003, SySy, Germany), rabbit anti-VGlut1(dilution 1:4000, 135302, SySy, Germany), rabbit anti-VGlut2 (dilution 1:2000, 135402, SySy, Germany), rabbit anti-VGAT(dilution

1:4000, 131002, SySy, Germany), rabbit anti-SNAP23 (dilution 1:2000, cat no. 111202, SySy, Germany), mouse anti-PSD95 (dilution 1:5000, cat no 75028,NeuroMab), mouse anti-GluN1(dilution 1:5000, cat no 114011, SySy, Germany), rabbit anti-GM130 (dilution 1:1000, cat no ab52649, Abcam), rabbit anti-Golgin-97 (dilution 1:2000, cat no13192, Cell Signaling Technology, USA), rabbit anti-EEA1 (dilution 1:2000, cat no 3288, Cell Signaling Technology, USA), rabbit anti-LC3B (dilution 1:1000, cat no 2775S, Cell Signaling Technology, USA), goat anti-CathepsinB (dilution 1:2000, AF965, R&D Systems), rabbit anti-GAPDH (dilution 1:1000, cat no 2118S, Cell Signaling Technology, USA), rabbit anti-GluT4 (dilution 1:1000, ab33780, Abcam), rabbit anti-CACNA1A (dilution 1:300, 152103, SySy, Germany), rabbit anti-VDAC (dilution 1:1000, cat no 4661S, Cell Signaling Technology, USA), rabbit anti-MBP (1:1000, cat no. 295003, SySy, Germany), mouse anti Creatine Kinase B (dilution 1:5000, cat no MAB9076, R&D Systems), rabbit anti HA (dilution 1:2000, CST3724; Cell Signaling Technology, USA) and rabbit anti SNAP25 (dilution 1:2000, ab109105, Abcam) antibodies.

Membranes were washed in three washing steps in TBST (each for 5 mins) and incubated with peroxidase-conjugated secondary antibodies for 2-3 h at 4 °C. The second antibody used were anti-rabbit (dilution 1:5000, A6154, Sigma), anti-mouse (dilution 1:5000, 715-035-150, Jackson Immuno Research) or rabbit anti-goat IgG secondary antibodies (dilution 1:1000, cat no ab6741, Abcam). After repeated washing, signals were visualized using a ChemiDoc XRS+ System (Bio-RAD Laboratories, USA).

Isolation of synaptic vesicles

Our purification procedures for synaptic vesicles were based on previously established immunoisolation methods 143,151. Protein G magnetic beads (cat no.88848, Thermofisher Scienctific) were washed three times with IP buffer (100 mM potassium tartrate, 4 mM HEPES-KOH, 2 mM MgCl2, pH 7.4) supplemented with a complete protease inhibitor cocktail (Roche). 5 μg monoclonal anti-Syp antibody directed against a cytoplasmic epitope (cat no 101011, synaptic systems, Germany) or control mouse IgG (10400C, Thermofisher Scientific) was used to incubate with 20-30 μl beads for 30 mins at RT in 2% BSA dissolved in IP buffer. Under this condition, 4-4.5 μg of antibody was coupled, as determined by Western blot and Commassie Blue staining. Immunoisolation of SVs were carried out at 0 °C-2 °C to prevent vesicular content leakage (with RT as a control). Briefly, the whole mouse brain was homogenized in 3 ml of IP buffer with a glass/Teflon homogenizer (20 strokes at 2000 rpm, WHEATON, USA and WIGGENS WB2000-M, Germany) immediately after decapitation. Homogenates were centrifuged for 25 mins at 35,000 x g and the supernatant was adjusted to approximately 3 mg/ml protein (Nanodrop 2000 C, Thermofisher Scientific). To capture the SVs for content detection, about 200 μl of supernatants (per 5 μg anti-Syp/IgG) were incubated with pre-coupled beads for 2.25 h under slow rotation at 2 °C. Beads were washed 6 times for further Western blot analysis and vesicular content detection. For pharmacological blockade of H+-gradient across SV membrane, the mix of supernatants and pre-coupled beads was diluted into 1.2 ml before the addition of inhibitors.

Determination of vesicular contents

To extract SV contents, immunoisolates were treated with 50 μl ultra-pure water. 100 μl methanol together with 100 μl acetonitrile were added to precipitate proteins in samples. After centrifugation for 20 mins at 168000 × g, supernatants were collected and centrifuged for 20 mins at 2000 × g to remove beads and proteins. Samples were pre-frozen with liquid nitrogen and vacuum dried at −45 °C overnight. Dried samples were kept frozen and re-suspended with 50 μl of 0.2 μM 13C-creatine (internal control) immediately before detection.

Capillary electrophoresis-mass spectrometry (CE-MS) was used to verify and quantify small molecules. CE/MS detection was applied with the coupling of PA800 plus CE system (Beckman Coulter, Brea, CA, USA) and mass spectrometry (TRIPLE QUAD 5500, AB SCIEX or Q Exactive HF-X, Thermo Scientific). Before SV content detection, we optimized MS detection of classical neuro transmitters, Cr and amino acids in positive ion mode. Firstly, the fragment ions (Q3) for a given molecule (precursor ions, Q1) were determined by either systematic scanning of standard sample solution (0.1 μM in 10% acetate acid) or referring to database (www.mzcloud.org). Secondly, optimal values of collision energy (CE), collision cell exit potential (CXP) and declustering potentials (DP) were determined for each pair of Q1/Q3. Thirdly, optimal combination of parameters (Q1/Q3, CE, CXP, DP) was chosen for each molecule. In addition, parameters were adjusted every 2 to 3 months for best signal to noise ratios.

CE/MS separations were carried out by capillaries (OptiMS silica surface cartridge, BECKMAN COULTER). The CE background electrolyte was 10 % acetate acid. Each new separation capillary was activated with rinsing under100 psi sequentially with methanol for 10 mins (forward), methanol for 3 mins (reverse), H2O for 10 mins (forward), H2O for 3 mins (reverse), 0.1 M NaOH for 10 mins (forward), water for 5 mins (reverse), 0.1 M HCl for 10 mins (forward), followed by water for 10 mins and then 10 % acetate acid for 10 mins (forward) and 3 mins (reverse), prior to the first use. Between analyses, the capillary was rinsed with 10% acetate acid under a 100 psi pressure for 5 min (forward) flowed by 75 psi for 4 min. The sample (50 μl) was injected with 2.5-4 psi for 30 s. Separation voltage of 25 kV was applied for 25 mins. To maintain stably spray during CE separation, ion spray voltage was applied at 1.7-1.9 kV. MS data were collected 5 mins after CE separation. Finally, the capillary was washed with 10% acetate acid for 10 mins, followed by methanol for 20 min and then10% acetate acid for 20 mins.

Standard solutions of 0.2 μM 13C-Cr (internal control) and analytes were used to plot standard curves. Linear standard curves (R2>0.98, for most cases, R2>0.99), calculated from peak area ratios corresponding to analytes and internal standards, were obtained for all molecules tested. The concentration ranges used for standards of Glu, GABA, ACh, 5-HT, Cr and alanine were 0.03-10 μM, 0.003-1 μM, 0.0003-0.1 μM, 0.003-1 μM, 0.03-1 μM and 0.03-1 μM, respectively. Standard curves were made at least twice for a given capillary. Analytes of SV contents were calculated using the standard curves and normalized to the amount of anti-Syp antibody conjugated to the beads.

Electron microscopy

All EM grids were glow discharged for 30 s using a plasma cleaner (Harrick PDC-32G-2). To free SVs from beads, 25 μl 0.1 M glycine-HCl (PH=2) was incubated for 1 min and quickly neutralized with 25 μl 0.1 M Tris (pH=10). Beads were quickly removed and 2-4 μL aliquots of SVs were applied to the carbon coated copper grids (Zhong Jing Ke Yi, Beijing). After 1 min, the grid was dried with a filter paper (Whatman No.1), and placed in the water, and then immediately stained using 2 % uranyl acetate for 30 s. At last uranyl acetate was removed and the grid was air dried. The grids were examined on a JEM-F200 electron microscope (JEOL, Japan) operated at 200 kV. Images were recorded using a 4k × 4k COMS One view camera (Gatan). Fixation of synaptosomal pellets was performed by immersion with pre-warmed 2.5% glutaraldehyde in 0.1 M phosphate buffer (pH 7.4) at RT for 2 h. After washing 4 times with 0.1 M phosphate buffer (pH 7.4) every 15 mins, samples were post-fixed with 1% osmium tetroxide (w/v) at 4°C for 1 h and then washed 3 times. Following en bloc staining with 2% uranyl acetate (w/v) at 4°C overnight, samples were dehydrated and embedded in fresh resin, polymerized at 65°C for 24 h. Ultrathin (70 nm) sections were obtained by Leica UC7 ultramicrotome (Leica Microsystems, German) and recorded on 80 kV in a JEOL Jem-1400 transmission electron-microscope (JEOL, Japan) using a CMOS camera (XAROSA, EMSIS)

Immunohistochemistry

Adult mice were anesthetized by i.p. injection with 2 % 2,2,2-tribromoethanol (T48402, Sigma Aldrich) in saline at a dose of 400 mg/kg, and perfused trancardially with 0.9 % saline followed by 4 % PFA in PBS (137 mM NaCl, 2.7 mM KCl, 10 mM Na2HPO4, 1.8 mM KH2PO4, pH=7.4).

Brains were cryoprotected with 30 % sucrose in 30 % sucrose 0.1 M PB (81 mM Na2HPO4, 19 mM NaH2PO4) and sectioned in the coronal plane (40 μm thick) using a Cryostat (Leica 3050S). For anti-HA immunostaining, we used a rabbit monoclonal anti-HA antibody (1:500 in 0.3 % Triton in PBS; 48 h incubation at 4 °C; #3724, Cell Signaling Technology), followed by a goat anti-rabbit Alexa Fluor 546 secondary antibody (1:1000; overnight at 4 °C; # A-11035, Invitrogen). Sections were mounted in a medium containing 50 % glycerol, cover-slipped, and sealed with nail polish. Images were acquired using virtual slide microscope (Olympus VS120-S6-W) and a laser-scanning confocal microscope (Zesis 710) and brain structures inferred with an established mouse brain atlas 258.

Preparations of brain slices

Male C57 mice (of 30-38 days-old) were anesthetized with pentobarbital (250 mg/kg) and decapitated. Brains were quickly removed and placed into ice-cold, low calcium, high magnesium artificial cerebrospinal fluid (ACSF) with sodium replaced by choline.

The medium consisted of 120 mM choline chloride, 2.5 mM KCl, 7 mM MgSO4, 0.5 mM CaCl2, 1.25 mM NaH2PO4, 5 mM sodium ascorbate, 3 mM sodium pyruvate, 26 mM NaHCO3, and 25 mM D-(+)-glucose, and was pre-equilibrated with 95% O2-5% CO2. Coronal brain slices (300 μm thick) were cut with a vibratome (Leica VT1200S). Slices were incubated for 1 h at 34°C with oxygenated ACSF containing: 124 mM NaCl, 2.5 mM KCl, 2 mM MgSO4, 2.5 mM CaCl2, 1.25 mM NaH2PO4, 26 mM NaHCO3, and 10 mM D-(+)-glucose.

Evoked release from brain slices

Coronal brain slices (each 300 μm thick, typically with a wet weight of 17-20 mg) were transferred into a specially designed superfusion chamber with a volume of approximately 200 μl, containing freshly 95% O2/5% CO2 oxygenated ACSF. Slices were equilibrated for 10 min in ACSF at a superfusion rate of 0.9-1.25 ml/min. The “control” sample was collected for 1 min just before high K+ stimulation (K-ACSF, 70 mM KCl replacing equal amount of NaCl). We waited for 30 seconds to allow K+ stimulus to immerse the slices (dead volume for solution transition of 200 μl and chamber volume of 200 μl), then the sample “70 mM K” in response to K-ACSF was collected for another 1 min. Following 10 min of washout period, we collected the third sample of “wash” for 1 min.

To detect Ca2+ dependent release, slices were pre-incubated for 10 mins with normal ACSF and equilibrated with Ca2+ free ACSF (containing 1 mM EGTA to chelate extracellular Ca2+) for 10 mins. The baseline sample “0 Ca2+ ACSF” was collected for 1 min. Superfusion solution was changed to Ca2+ free K-ACSF for 2 mins and sample “0 Ca2+ 70 mM K” was collected (dead volume for solution transition of 400 μl and chamber volume of 200 μl). After that, the solution was changed back to normal ACSF for 10 mins and K-ACSF for 2 mins. The sample “2.5 mM Ca 70 mM K” for the last min was collected.

Samples were subjected to CE-MS in a method similar to SV content detection, except for the following: 1) standards were dissolved in ACSF or other buffers used in release experiment; 2) concentration ranges used for standards of Glu was from 0.003-1 μM. 3) To protect the MS from salt pollution, data were collected from 10 mins −20 mins during CE separation.

Patch-Clamp recordings

Slices were transferred to a recording chamber on an upright fluorescent microscope equipped with differential interference contrast optics (DIC; Olympus BX51WI). Slices were submerged and superfused with ACSF at about 2.8 ml/min at 24-26 °C. Whole cell patch recordings were routinely achieved from layer 4/5 medium sized pyramidal neurons from the somatosensory cortex. Patch pipettes (3-5 MΩ) contained: 140 mM K-gluconate, 10 mM HEPES, 0.5 mM EGTA, 5 mM KCl, 3 mM Na2-ATP, 0.5 mM Na3GTP, and 4 mM MgCl2 (with pH adjusted to 7.3 and osmolarity of 290 mOsm/kg). Current-clamp recordings were carried out with a computer-controlled amplifier (Multiclamp 700B, Molecular Devices) and traces were digitized at 10 kHz (DigiData 1550B, Molecular Devices). Data were collected and analyzed using Clampfitor Clampex 10 software (Molecular Devices).

Cells were characterized by their membrane responses and firing patterns during hyperpolarizing and depolarizing current steps (−100 to +500 pA, increment: 50 pA or 25 pA, 500 ms). Regular spiking pyramidal neurons were identified by moderate maximal spiking frequencies (20-60 Hz, that is, 10-30 spikes per 500 ms, Figure 7E-G), increasing of inter-spike intervals during depolarizing step (Figure 7B and Supplementary Figure 11A), high action amplitude (Supplementary Figure 12E) and large half width (Supplementary Figure 12F) 201,202. After the mean firing frequency evoked by current injections reached the steady state for at least 5 mins (typically 20-30 mins following the formation of whole-cell configuration), 100 μM Cr was bath-applied for 6 mins. Typically, Cr was applied for a 2nd time following washout to re-confirm the effects.

Synaptosome preparation

Synaptosomes were isolated by Ficoll/sucrose density-gradient centrifugation 203,209,221,259. Whole brains from adult male mice were homogenized with 15 strokes at 900 rpm in buffer A (320 mM sucrose, 1 mM EDTA, 1mM EGTA, 10 mM Tris–HCl, pH 7.4, with a complete protease inhibitor cocktail (Roche)). The homogenate (H fraction) was centrifuged at 1000 g for 10 min to precipitate the membrane fragments and nuclei (P1 fraction). Supernatant was centrifuged again at 1000 g for 10 min and the resulting supernatant (S1) was centrifuged at 12,000 g for 20 min. Supernatant was the S2 fraction, and the pellet was re-suspended with buffer A and centrifuged at 12,000 g for 20 min. The resulting pellet was crude synaptosomes (P2 fraction), containing synaptosomes with mitochondria and microsomes.

Crude synaptosomes (P2 fraction) was re-suspended with 150-200 μl buffer B (320 mM sucrose and 10 mM Tris–HCI (pH 7.4)). The sample was carefully overlaid on the top of a gradient of 2 ml of 7.5 % (wt/vol in buffer B) Ficoll and 1.8 ml of 13% (wt/vol in buffer B) Ficoll and centrifuged at 98,000 g for 45 min at 2-4 °C in a swinging-bucket rotor. A myelin band was present near the surface, and th synaptosomes band (fraction Sy) was present at the interface between the 13% and 7.5 % Ficoll layers, with the mitochondria being pelleted at the bottom. For further western analysis, the supernatant was divided into 6 fractions (600 μl for each fraction) and the mitochondria pellet was discarded. The isolated synaptosomes was included in fraction 4.

For Western analysis, fractions H, S1, P1, S2, P2 and Sy were adjusted to 0.5 mg/ml by bicinchoninic acid assay (BCA) method with reference to NanoDrop™ 2000 Spectrophotometers. 3.35 μg protein was loaded for each lane. Fractions 1-6 were loaded with the same volume (10 μl composed of 6.7 μl sample and 3.3 μl loading buffer) for each lane.

Creatine uptake into synaptosomes

To remove Ficoll, we diluted the synaptosomal band (480 μl) with 4.3 ml of a pH 7.4 buffer C containing (in mM) 240 mannitol, 10 glucose, 4.8 Kgluconate, 2.2 Cagluconate, 1.2 MgSO4, 1.2 KH2PO4 and 25 HEPES-Tris. The sample was then centrifuged at 12,000 g and the pellet was re-suspended with buffer C. Uptake experiments were either performed at 37°C or at 0°C (control). For each sample, 25-43 μg of synaptosomes (with a volume of 40-50 μl) were added to 360 μl buffer containing (in mmol/L) 100 NaCl, 40 mannitol, 10 glucose, 4.8 Kgluconate, 2.2 Cagluconate, 1.2 MgSO4, 1.2 KH2PO4, 25 HEPES and 25 Tris (pH adjusted to 7.4). A mixture of 18 μM [14C]-creatine (0.4 μCi) and 5 μM creatine was quickly added. After 10 min, uptake was terminated by the addition of 1 ml of NaCl-free ice-cold buffer C. Samples were immediately filtered, under vacuum, through a Whatman GF/C glass filter (1825-025) pre-wetted with buffer C. Filters were further washed with 10 ml of ice-cold buffer C, dissolved in scintillation fluid and the radioactivity determined by liquid scintillation spectrometry.

Creatine uptake into SVs

The uptake of 13C-creatine was assayed according to a conventional procedure 260 with slight modifications: the immunoisolated synaptic vesicles by 10 μg Syp antibody (101011, Synaptic Systems, Germany) were resuspended with the uptake buffer (150mM meglumine-tartrate, 4 mM KCl, 4 mM MgSO4, 10mM HEPES-KOH pH 7.4, and complete EDTA-free protease inhibitor cocktail) containing 4mM Mg-ATP or additional 4mM MgSO4, followed by preincubation for 30 min at 25L. The uptake reaction was started by addition of 1mM 13C-creatine dissolved in the uptake buffer with a final volume of 125 μL (pH at 6.8). After 10min at 25L, 1mL of ice-cold uptake buffer was added to the incubation to stop the reaction, followed by 5 more times washing. The SV contents were extracted using the protocol described in the determination of vesicular contents part. 100 nM Cr was used as the internal control. CE-MS and LC-MS was used to verify and quantify the creatine contents of samples. A Vanquish UHPLC system coupled to a Q Exactive HF-X mass spectrometer (both instrument Thermo Fisher Scientific, USA) were used for LC-MS analysis along with SeQuant ZIC-HILIC column (150 mm × 2.1 mm, 3.5 μm, Merck Millipore, 150442) in the positive mode and SeQuant ZIC-pHILIC column (150 mm × 2.1 mm, 5 μm, Merck Millipore, 150460) in the negative mode. For ZIC-HILIC column, the mobile phase A was 0.1% formic acid in water and the mobile phase B was 0.1% formic acid in acetonitrile. The linear gradient was as following: 0 min, 80% B; 6 min, 50% B; 13 min, 50% B; 14 min, 20% B; 18 min, 20% B; 18.5 min, 80% B; 30 min, 80% B. The flow rate used was 300 μl/min and the column temperature was maintained at 30°C. For ZIC-pHILIC column, the mobile phase A is 20 mM ammonium carbonate in water, adjusted to pH 9.0 with 0.1% ammonium hydroxide solution (25%) and the mobile phase B is 100% acetonitrile. The linear gradient was as follows: 0 min, 80% B; 2 min, 80% B; 19 min, 20% B; 20 min, 80% B; 30 min, 80% B. The flow rate used is 150 μl/min and the column temperature is 25°C. Samples were maintained at 4°C in Vanquish autosampler. 3 µl of extracted metabolites were injected for each run. IP samples were subjected to ZIC-HILIC column in positive mode for major metabolites detection, and then subject to ZIC-pHILIC column in negative mode for orthogonal detection.

Supporting information